### 2021-11-25 appendix/featurizerTable.tex
Updates from Overleaf

\begin{center}
\begin{longtable}{M{3.5cm} M{6.5cm} M{2.0cm}}
\caption{This work's chosen 39 featurizers from matminer. Descriptions are either found from Ref. \cite{Ward2018} or from the project's Github page. For entries lacking references, we refer to Ref.~\cite{Ward2018}.}
\label{table:featurizers} 
\hline 
Features & Description & Reference \\ 
\hline 
  \textbf{Composition features} & & \\ 
  AtomicOrbitals & Highest occupied molecular orbital (HOMO) & \cite{Kotochigova1997}  \\   
   & and lowest unoccupied molecular orbital (LUMO) &  \\   
  AtomicPacking-Efficiency & Packing efficiency & \cite{Laws2015}  \\   
  BandCenter & Estimate absolute position of band center  & \cite{Butler1978} \\   
   & using geometric mean of electronegativity &  \\  
  ElementFraction & Fraction of each element in a composition &    \\   
  ElementProperty & Statistics of various element properties & \cite{Ong2013,Ward2016, Deml2016}  \\   
  IonProperty & Maximum and average ionic character & \cite{Ward2016} \\   
  Miedema & Formation enthalpies of intermetallic compounds, solid solutions, & \cite{Weeber1987} \\   
   & and amorphous phases using semi-empirical Miedema model &  \\   
  Stoichiometry & $L^p$ norm-based stoichiometric attributes & \cite{Ward2016} \\   
  TMetalFraction & Fraction of magnetic transition metals & \cite{Deml2016}  \\   
  ValenceOrbital & Valence orbital attributes such as & \cite{Ward2016}  \\   
   &  the mean number of electrons in each shell &   \\   
  YangSolid-Solution & Mixing thermochemistry and size mismatch terms & \cite{Yang2012} \\
    \hline 
  \textbf{Oxide composition features} &  &  \\
  Electronegativity-Diff & Statistics on electronegativity difference & \cite{Deml2016} \\   
   &  between anions and cations & \\ 
  OxidationStates & Statistics of oxidation states & \cite{Deml2016}  \\   
\hline 
  \textbf{Structure features} & & \\   
  DensityFeatures & Calculate density, volume per atom and packing fraction & - \\   
  GlobalSymmetry-Features & Determines spacegroup number, crystal system  & - \\   
   & (1-7) and inversion symmetry & \\ 
  RadialDistribution-Function & Calculates the radial distribution  & - \\   
   & function of a crystal system & \\ 
  CoulombMatrix & Generate the Coulomb matrix for nuclear interactions  & \cite{Rupp2012}  \\    
  %CoulombMatrix & Generate the Coulomb matrix, which is a representation of the nuclear coulombic interaction of the input structure. & \cite{Rupp2012}  \\      
  PartialRadial-Distribution-Function & Compute the partial radial distribution  & \cite{Schuett2014}  \\   
   & function of a crystal structure & \\ 
  SineCoulomb-Matrix & Computes a variant of the coulomb matrix & \cite{Faber2015}  \\   
   & developed for periodic crystals & \\ 
  EwaldEnergy & Computes the energy from Coulombic interactions  & \cite{Ewald1921}  \\   
   & based on charge states of each site & \\ 
  BondFractions & Compute the fraction of each bond in a  & \cite{Hansen2015}  \\   
   & structure, based on nearest neighbors & \\ 
  Structural Heterogeneity & Calculates the variance in bond lengths and  & \cite{Ward2017}  \\   
   & atomic volumes in a structure & \\ 
  MaximumPacking-Efficiency & Calculates the maximum packing efficiency of a structure & \cite{Ward2017} \\   
  Chemical-Ordering & Computes how much the ordering of species  & \cite{Ward2017}  \\   
   & differs from random in a structure & \\ 
  XRDPowder-Pattern & 1D array representing normalized powder diffraction & \cite{Ong2013} \\ 
   &of a structure as calculated by pymatgen  & \\ 
    \hline 
  \textbf{Site features} & & \\
  AGNI-Fingerprints & Calculates the product integral of RDF & \cite{Botu2014}  \\    
   & and Gaussian window function  & \\ 
  AverageBond- Agle & Determines the average bond angle of a specific  & \cite{Jong2016}  \\
   & site with its nearest neighbors  & \\ 
  AverageBond-Length & Determines the average bond length between one specific site & \cite{Jong2016}  \\ 
   & and all its nearest neighbors  & \\ 
  BondOrientational-Parameter & Calculates the averages of spherical  & \cite{Seko2017, Steinhardt1983}  \\ 
   & harmonics of local neighbors & \\ 
  ChemEnvSite-Fingerprint & Calculates the resemblance of given sites to ideal & \cite{Waroquiers2017, Zimmermann2017}  \\   
   & environment using pymatgens ChemEnv package  & \\ 
  Coordination-Number & The number of first nearest neighbors of a site & \cite{Zimmermann2017}  \\
  CrystalNN-Fingerprint & A local order parameter fingerprint for periodic crystals & -  \\   
  GaussianSymm-Func & Calculates the gaussian radial and angular symmetry functions  & \cite{Behler2011,Khorshidi2016}  \\   
   & originally suggested for fitting machine learning potentials &  \\ 
  GeneralizedRadial-Distribution-Function & Computes the general radial distribution function for a site & \cite{Seko2017}  \\   
  LocalProperty-Difference & Computes the difference in elemental properties  & \cite{Ward2017, Jong2016} \\   
   & between a site and its neighboring sites & \\ 
  OPSite-Fingerprint & Computes the local structure order parameters & \cite{Zimmermann2017} \\   
   & from a site's neighbor environment  & \\ 
  Voronoi-Fingerprint & Calculates the Voronoi tessellation-based  & \cite{Peng2011,Wang2019} \\
   & features around a target site & \\ 
\hline     
  \textbf{Density of state features} & & \\
  DOSFeaturizer & Computes top contributors to the density of  & \cite{Dylla2020} \\ 
  & states at the VBM and CBM  & \\ 
  %DOSFeaturizer & Computes top contributors to the density of states at the valence and conduction band edges. Thus includes chemical species, orbital character, and orbital location information. & \cite{Dylla2020} \\
  \hline  
  \textbf{Band structure features} & & \\
  BandFeaturizer & Converts a complex electronic band  & - \\ 
   &structure into discrete quantities  & \\ 
%BandFeaturizer & Converts a complex electronic band structure into quantities such as band gap and the norm of k point coordinates at which the conduction band minimum and valence band maximum occur. & - \\
\hline 
\end{longtable}
\end{center}