### 2021-11-25 figures/tikz-plots/insightful-approach.tex

\begin{figure}[!ht]
  \centering
  \begin{forest}
    for tree={l sep=1em, s sep=1em, anchor=center, inner sep=0.3em, fill=red!50, circle}
    [$24544$ compounds, node box, alias=bagging, above=1em
    [Logistic regression,node box,alias=a1
      [$23702$]
      [$842$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    [Decision tree,node box,alias=a1
      [$23347$]
      [$1197$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    [Random forest,node box,alias=a1
      [$24001$]
      [$543$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    [Gradient boost,node box,alias=a1
      [$23948$]
      [$596$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    ]
  \end{forest}
\vspace*{-95mm}
\caption{Visualization of the predicted material candidates from the insightful approach. The green nodes display the number of predicted suitable candidates while unsuitable ones are marked in red.} 
\label{fig:03-predictions}
\end{figure}