### 2021-12-10 figures/tikz-plots/prediction-fig.tex
Updates from Overleaf

\begin{figure}[t]
    \centering
    (a)
    \\
    \begin{forest}
        for tree={l sep=1em, s sep=1em, anchor=center, inner sep=0.3em, fill=red!50, circle}
        [$23623$ compounds, node box, alias=bagging, above=1em
        [Logistic regression,node box,alias=a1
          [$11243$]
          [$12380$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
          ]
        ]
        [Decision trees,node box,alias=a1
          [$12308$]
          [$11315$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
          ]
        ]
        [Random forests,node box,alias=a1
          [$9345$]
          [$14278$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
          ]
        ]
        [Gradient boosting,node box,alias=a1
          [$11835$]
          [$11788$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
          ]
        ]
        ]
    \end{forest}
%\vspace*{-95mm}
    \\
    (b)
    \\
\begin{forest}
    for tree={l sep=1em, s sep=1em, anchor=center, inner sep=0.3em, fill=red!50, circle}
    [$22550$ compounds, node box, alias=bagging, above=1em
    [Logistic regression,node box,alias=a1
      [$7557$]
      [$14993$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    [Decision trees,node box,alias=a1
      [$8143$]
      [$14407$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    [Random forests,node box,alias=a1
      [$7199$]
      [$15351$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    [Gradient boosting,node box,alias=a1
      [$8762$]
      [$13788$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    ]
  \end{forest}
%\vspace*{20mm}
    \\
    (c)
    \\
 \begin{forest}
    for tree={l sep=1em, s sep=1em, anchor=center, inner sep=0.3em, fill=red!50, circle}
    [$24544$ compounds, node box, alias=bagging, above=1em
    [Logistic regression,node box,alias=a1
      [$23702$]
      [$842$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    [Decision trees,node box,alias=a1
      [$23347$]
      [$1197$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    [Random forests,node box,alias=a1
      [$24001$]
      [$543$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    [Gradient boosting,node box,alias=a1
      [$23948$]
      [$596$,fill=green!50,edge label={node[above=1ex,green arrow]{}}
      ]
    ]
    ]
  \end{forest}
\vspace*{-95mm}
\caption{Visualization of the predicted material candidates from the (a) Ferrenti, (b) augemented Ferrenti and (c) insightful approaches. The green nodes display the number of predicted suitable candidates while unsuitable ones are marked in red.}
\label{fig:predictions}
\end{figure}