### 2022-01-07 figures/feature-importance/01-ferrenti-approachDT-final.tex
Updates from Overleaf

% This file was created by tikzplotlib v0.9.8.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.8,0.4,0.466666666666667}

\begin{axis}[
height=1.4222438079424382in,
tick align=outside,
tick pos=left,
width=2.8444876158848764in,
x grid style={white!69.0196078431373!black},
xmajorgrids,
xmin=0.5, xmax=26,
xtick style={color=black},
xticklabels={,,},
y grid style={white!69.0196078431373!black},
ymajorgrids,
ymin=0, ymax=0.296093145959793,
ytick style={color=black},
legend style={font=\footnotesize},
]
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:1.4,0.0602126647764764);
\draw[draw=none,fill=color0] (axis cs:1.6,0) rectangle (axis cs:2.4,0.110963188395912);
\draw[draw=none,fill=color0] (axis cs:2.6,0) rectangle (axis cs:3.4,0.0240371193898026);
\draw[draw=none,fill=color0] (axis cs:3.6,0) rectangle (axis cs:4.4,0.0221547131912102);
\draw[draw=none,fill=color0] (axis cs:4.6,0) rectangle (axis cs:5.4,0.246093145959793);
\draw[draw=none,fill=color0] (axis cs:5.6,0) rectangle (axis cs:6.4,0.0188217883182872);
\draw[draw=none,fill=color0] (axis cs:6.6,0) rectangle (axis cs:7.4,0.0105015974384699);
\draw[draw=none,fill=color0] (axis cs:7.6,0) rectangle (axis cs:8.4,0.00699136294394965);
\draw[draw=none,fill=color0] (axis cs:8.6,0) rectangle (axis cs:9.4,0.00853041123767904);
\draw[draw=none,fill=color0] (axis cs:9.6,0) rectangle (axis cs:10.4,0.00934264373238672);
\draw[draw=none,fill=color0] (axis cs:10.6,0) rectangle (axis cs:11.4,0.0201245026439328);
\draw[draw=none,fill=color0] (axis cs:11.6,0) rectangle (axis cs:12.4,0.00803736953390133);
\draw[draw=none,fill=color0] (axis cs:12.6,0) rectangle (axis cs:13.4,0.0201541360551478);
\draw[draw=none,fill=color0] (axis cs:13.6,0) rectangle (axis cs:14.4,0.0180446021881651);
\draw[draw=none,fill=color0] (axis cs:14.6,0) rectangle (axis cs:15.4,0.0132118462234375);
\draw[draw=none,fill=color0] (axis cs:15.6,0) rectangle (axis cs:16.4,0.0181046069828721);
\draw[draw=none,fill=color0] (axis cs:16.6,0) rectangle (axis cs:17.4,0.0309174582001746);
\draw[draw=none,fill=color0] (axis cs:17.6,0) rectangle (axis cs:18.4,0.0359878148351983);
\draw[draw=none,fill=color0] (axis cs:18.6,0) rectangle (axis cs:19.4,0.0271294275084838);
\draw[draw=none,fill=color0] (axis cs:19.6,0) rectangle (axis cs:20.4,0.012494436624754);
\draw[draw=none,fill=color0] (axis cs:20.6,0) rectangle (axis cs:21.4,0.0167371542524648);
\draw[draw=none,fill=color0] (axis cs:21.6,0) rectangle (axis cs:22.4,0.0172798021408093);
\draw[draw=none,fill=color0] (axis cs:22.6,0) rectangle (axis cs:23.4,0.00542141841736881);
\draw[draw=none,fill=color0] (axis cs:23.6,0) rectangle (axis cs:24.4,0.0402418644282019);
\draw[draw=none,fill=color0] (axis cs:24.6,0) rectangle (axis cs:25.4,0.0227346258367308);
\path [draw=color0, semithick]
(axis cs:1,5.12811391574566e-05)
--(axis cs:1,0.120374048413795);

\path [draw=color0, semithick]
(axis cs:2,0.0688908846055959)
--(axis cs:2,0.153035492186228);

\path [draw=color0, semithick]
(axis cs:3,0.000443281753072355)
--(axis cs:3,0.0476309570265328);

\path [draw=color0, semithick]
(axis cs:4,0.000463808547678048)
--(axis cs:4,0.0438456178347424);

\path [draw=color0, semithick]
(axis cs:5,0.201175759797789)
--(axis cs:5,0.291010532121796);

\path [draw=color0, semithick]
(axis cs:6,-0.00207540963664902)
--(axis cs:6,0.0397189862732235);

\path [draw=color0, semithick]
(axis cs:7,-0.000261197656391885)
--(axis cs:7,0.0212643925333317);

\path [draw=color0, semithick]
(axis cs:8,-0.00523656669530865)
--(axis cs:8,0.0192192925832079);

\path [draw=color0, semithick]
(axis cs:9,-0.00108881158485408)
--(axis cs:9,0.0181496340602122);

\path [draw=color0, semithick]
(axis cs:10,-0.00317148697566561)
--(axis cs:10,0.0218567744404391);

\path [draw=color0, semithick]
(axis cs:11,0.00759809291557517)
--(axis cs:11,0.0326509123722904);

\path [draw=color0, semithick]
(axis cs:12,-0.00126897867798911)
--(axis cs:12,0.0173437177457918);

\path [draw=color0, semithick]
(axis cs:13,0.00435547873601669)
--(axis cs:13,0.035952793374279);

\path [draw=color0, semithick]
(axis cs:14,-0.000431492489608061)
--(axis cs:14,0.0365206968659383);

\path [draw=color0, semithick]
(axis cs:15,-0.00810189470153616)
--(axis cs:15,0.0345255871484112);

\path [draw=color0, semithick]
(axis cs:16,-0.00758733165253984)
--(axis cs:16,0.0437965456182841);

\path [draw=color0, semithick]
(axis cs:17,-0.000304997596235309)
--(axis cs:17,0.0621399139965845);

\path [draw=color0, semithick]
(axis cs:18,0.00670250518936169)
--(axis cs:18,0.0652731244810349);

\path [draw=color0, semithick]
(axis cs:19,0.00234124172436161)
--(axis cs:19,0.0519176132926061);

\path [draw=color0, semithick]
(axis cs:20,-0.0107608268316967)
--(axis cs:20,0.0357497000812046);

\path [draw=color0, semithick]
(axis cs:21,-0.00616044853556576)
--(axis cs:21,0.0396347570404953);

\path [draw=color0, semithick]
(axis cs:22,-0.00878787812034752)
--(axis cs:22,0.0433474824019661);

\path [draw=color0, semithick]
(axis cs:23,-0.00371385025428554)
--(axis cs:23,0.0145566870890232);

\path [draw=color0, semithick]
(axis cs:24,-0.00489403909324856)
--(axis cs:24,0.0853777679496524);

\path [draw=color0, semithick]
(axis cs:25,-0.0140804363436156)
--(axis cs:25,0.0595496880170773);
\addlegendimage{semithick, color=color0};
\addlegendentry{Decision tree};
\addplot [semithick, color0, mark=-, mark size=0, mark options={solid}, only marks]
table {%
1 5.12811391574566e-05
2 0.0688908846055959
3 0.000443281753072355
4 0.000463808547678048
5 0.201175759797789
6 -0.00207540963664902
7 -0.000261197656391885
8 -0.00523656669530865
9 -0.00108881158485408
10 -0.00317148697566561
11 0.00759809291557517
12 -0.00126897867798911
13 0.00435547873601669
14 -0.000431492489608061
15 -0.00810189470153616
16 -0.00758733165253984
17 -0.000304997596235309
18 0.00670250518936169
19 0.00234124172436161
20 -0.0107608268316967
21 -0.00616044853556576
22 -0.00878787812034752
23 -0.00371385025428554
24 -0.00489403909324856
25 -0.0140804363436156
};
\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 0.120374048413795
2 0.153035492186228
3 0.0476309570265328
4 0.0438456178347424
5 0.291010532121796
6 0.0397189862732235
7 0.0212643925333317
8 0.0192192925832079
9 0.0181496340602122
10 0.0218567744404391
11 0.0326509123722904
12 0.0173437177457918
13 0.035952793374279
14 0.0365206968659383
15 0.0345255871484112
16 0.0437965456182841
17 0.0621399139965845
18 0.0652731244810349
19 0.0519176132926061
20 0.0357497000812046
21 0.0396347570404953
22 0.0433474824019661
23 0.0145566870890232
24 0.0853777679496524
25 0.0595496880170773
};
\end{axis}

\end{tikzpicture}