### 2022-01-07 figures/feature-importance/01-ferrenti-approachRF-final.tex

% This file was created by tikzplotlib v0.9.8.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.866666666666667,0.8,0.466666666666667}

\begin{axis}[
height=1.4222438079424382in,
tick align=outside,
tick pos=left,
width=2.8444876158848764in,
x grid style={white!69.0196078431373!black},
xmajorgrids,
xmin=0.5, xmax=26,
xtick style={color=black},
xticklabels={,,},
y grid style={white!69.0196078431373!black},
ymajorgrids,
ymin=0, ymax=0.163961791193378,
ytick style={color=black},
yticklabels={,0,,0.1,,0.2},
legend style={font=\footnotesize},
]
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:1.4,0.0630557431105419);
\draw[draw=none,fill=color0] (axis cs:1.6,0) rectangle (axis cs:2.4,0.0831193744181552);
\draw[draw=none,fill=color0] (axis cs:2.6,0) rectangle (axis cs:3.4,0.022912143158908);
\draw[draw=none,fill=color0] (axis cs:3.6,0) rectangle (axis cs:4.4,0.0195487261065447);
\draw[draw=none,fill=color0] (axis cs:4.6,0) rectangle (axis cs:5.4,0.113961791193378);
\draw[draw=none,fill=color0] (axis cs:5.6,0) rectangle (axis cs:6.4,0.0166845058106126);
\draw[draw=none,fill=color0] (axis cs:6.6,0) rectangle (axis cs:7.4,0.0111880865301946);
\draw[draw=none,fill=color0] (axis cs:7.6,0) rectangle (axis cs:8.4,0.0109815412799707);
\draw[draw=none,fill=color0] (axis cs:8.6,0) rectangle (axis cs:9.4,0.013418791526676);
\draw[draw=none,fill=color0] (axis cs:9.6,0) rectangle (axis cs:10.4,0.0148028536779526);
\draw[draw=none,fill=color0] (axis cs:10.6,0) rectangle (axis cs:11.4,0.0131608319701371);
\draw[draw=none,fill=color0] (axis cs:11.6,0) rectangle (axis cs:12.4,0.00838039909198669);
\draw[draw=none,fill=color0] (axis cs:12.6,0) rectangle (axis cs:13.4,0.0148854194402229);
\draw[draw=none,fill=color0] (axis cs:13.6,0) rectangle (axis cs:14.4,0.0203640215741575);
\draw[draw=none,fill=color0] (axis cs:14.6,0) rectangle (axis cs:15.4,0.0147922332160194);
\draw[draw=none,fill=color0] (axis cs:15.6,0) rectangle (axis cs:16.4,0.0167058216702408);
\draw[draw=none,fill=color0] (axis cs:16.6,0) rectangle (axis cs:17.4,0.0285684173665608);
\draw[draw=none,fill=color0] (axis cs:17.6,0) rectangle (axis cs:18.4,0.0251819365387591);
\draw[draw=none,fill=color0] (axis cs:18.6,0) rectangle (axis cs:19.4,0.0342442655954786);
\draw[draw=none,fill=color0] (axis cs:19.6,0) rectangle (axis cs:20.4,0.0138513854588785);
\draw[draw=none,fill=color0] (axis cs:20.6,0) rectangle (axis cs:21.4,0.0135559278428533);
\draw[draw=none,fill=color0] (axis cs:21.6,0) rectangle (axis cs:22.4,0.0150237331161702);
\draw[draw=none,fill=color0] (axis cs:22.6,0) rectangle (axis cs:23.4,0.0096270211737359);
\draw[draw=none,fill=color0] (axis cs:23.6,0) rectangle (axis cs:24.4,0.0206136662393266);
\draw[draw=none,fill=color0] (axis cs:24.6,0) rectangle (axis cs:25.4,0.0335101355200213);
\path [draw=color0, semithick]
(axis cs:1,0.0573010303012627)
--(axis cs:1,0.0688104559198211);

\path [draw=color0, semithick]
(axis cs:2,0.0715458523831103)
--(axis cs:2,0.0946928964532001);

\path [draw=color0, semithick]
(axis cs:3,0.0162325617369429)
--(axis cs:3,0.0295917245808731);

\path [draw=color0, semithick]
(axis cs:4,0.0151929284504258)
--(axis cs:4,0.0239045237626635);

\path [draw=color0, semithick]
(axis cs:5,0.103650391870818)
--(axis cs:5,0.124273190515939);

\path [draw=color0, semithick]
(axis cs:6,0.0127466082846881)
--(axis cs:6,0.0206224033365371);

\path [draw=color0, semithick]
(axis cs:7,0.00938556140819314)
--(axis cs:7,0.012990611652196);

\path [draw=color0, semithick]
(axis cs:8,0.00854307516043412)
--(axis cs:8,0.0134200073995072);

\path [draw=color0, semithick]
(axis cs:9,0.0116751719811282)
--(axis cs:9,0.0151624110722239);

\path [draw=color0, semithick]
(axis cs:10,0.0103717999049377)
--(axis cs:10,0.0192339074509675);

\path [draw=color0, semithick]
(axis cs:11,0.00968301823796202)
--(axis cs:11,0.0166386457023121);

\path [draw=color0, semithick]
(axis cs:12,0.0064665824303152)
--(axis cs:12,0.0102942157536582);

\path [draw=color0, semithick]
(axis cs:13,0.00834089792296834)
--(axis cs:13,0.0214299409574775);

\path [draw=color0, semithick]
(axis cs:14,0.005070824077227)
--(axis cs:14,0.0356572190710881);

\path [draw=color0, semithick]
(axis cs:15,0.00706511442570956)
--(axis cs:15,0.0225193520063292);

\path [draw=color0, semithick]
(axis cs:16,0.00601575858602535)
--(axis cs:16,0.0273958847544562);

\path [draw=color0, semithick]
(axis cs:17,0.0113173620435266)
--(axis cs:17,0.045819472689595);

\path [draw=color0, semithick]
(axis cs:18,0.0103353119352383)
--(axis cs:18,0.0400285611422799);

\path [draw=color0, semithick]
(axis cs:19,0.011625254341459)
--(axis cs:19,0.0568632768494982);

\path [draw=color0, semithick]
(axis cs:20,0.00342679873482929)
--(axis cs:20,0.0242759721829277);

\path [draw=color0, semithick]
(axis cs:21,0.00673116906011602)
--(axis cs:21,0.0203806866255907);

\path [draw=color0, semithick]
(axis cs:22,0.00496948896783449)
--(axis cs:22,0.0250779772645059);

\path [draw=color0, semithick]
(axis cs:23,0.00660725108889622)
--(axis cs:23,0.0126467912585756);

\path [draw=color0, semithick]
(axis cs:24,0.00380908046748388)
--(axis cs:24,0.0374182520111692);

\path [draw=color0, semithick]
(axis cs:25,0.00587721747702246)
--(axis cs:25,0.0611430535630202);
\addlegendimage{semithick, color=color0};
\addlegendentry{Random forest};
\addplot [semithick, color0, mark=-, mark size=0, mark options={solid}, only marks]
table {%
1 0.0573010303012627
2 0.0715458523831103
3 0.0162325617369429
4 0.0151929284504258
5 0.103650391870818
6 0.0127466082846881
7 0.00938556140819314
8 0.00854307516043412
9 0.0116751719811282
10 0.0103717999049377
11 0.00968301823796202
12 0.0064665824303152
13 0.00834089792296834
14 0.005070824077227
15 0.00706511442570956
16 0.00601575858602535
17 0.0113173620435266
18 0.0103353119352383
19 0.011625254341459
20 0.00342679873482929
21 0.00673116906011602
22 0.00496948896783449
23 0.00660725108889622
24 0.00380908046748388
25 0.00587721747702246
};
\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 0.0688104559198211
2 0.0946928964532001
3 0.0295917245808731
4 0.0239045237626635
5 0.124273190515939
6 0.0206224033365371
7 0.012990611652196
8 0.0134200073995072
9 0.0151624110722239
10 0.0192339074509675
11 0.0166386457023121
12 0.0102942157536582
13 0.0214299409574775
14 0.0356572190710881
15 0.0225193520063292
16 0.0273958847544562
17 0.045819472689595
18 0.0400285611422799
19 0.0568632768494982
20 0.0242759721829277
21 0.0203806866255907
22 0.0250779772645059
23 0.0126467912585756
24 0.0374182520111692
25 0.0611430535630202
};
\end{axis}

\end{tikzpicture}