### 2022-01-07 figures/feature-importance/03-insightful-approachDT-final-2.tex

% This file was created by tikzplotlib v0.9.8.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.8,0.4,0.466666666666667}

\begin{axis}[
height=1.4222438079424382in,
tick align=outside,
tick pos=left,
width=2.8444876158848764in,
x grid style={white!69.0196078431373!black},
xmajorgrids,
xmin=0.5, xmax=15.5,
xtick style={color=black},
xticklabels={,,},
y grid style={white!69.0196078431373!black},
ymajorgrids,
ymin=0, ymax=0.84619547527771,
ytick style={color=black},
yticklabels={,0,, 0.4,,0.8},
legend style={font=\footnotesize}
]
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:1.4,0.731619547527771);
\draw[draw=none,fill=color0] (axis cs:1.6,0) rectangle (axis cs:2.4,0.18113983898672);
\draw[draw=none,fill=color0] (axis cs:2.6,0) rectangle (axis cs:3.4,0.00345402728374034);
\draw[draw=none,fill=color0] (axis cs:3.6,0) rectangle (axis cs:4.4,0.0208705310819208);
\draw[draw=none,fill=color0] (axis cs:4.6,0) rectangle (axis cs:5.4,0.0155739365467582);
\draw[draw=none,fill=color0] (axis cs:5.6,0) rectangle (axis cs:6.4,0.0237724911651494);
\draw[draw=none,fill=color0] (axis cs:6.6,0) rectangle (axis cs:7.4,0.003003922613769);
\draw[draw=none,fill=color0] (axis cs:7.6,0) rectangle (axis cs:8.4,0.00923425403456265);
\draw[draw=none,fill=color0] (axis cs:8.6,0) rectangle (axis cs:9.4,0.0113314507596083);
\path [draw=color0, semithick]
(axis cs:1,0.709690218797472)
--(axis cs:1,0.753548876258071);

\path [draw=color0, semithick]
(axis cs:2,0.159795728748012)
--(axis cs:2,0.202483949225428);

\path [draw=color0, semithick]
(axis cs:3,-0.00265218119467417)
--(axis cs:3,0.00956023576215485);

\path [draw=color0, semithick]
(axis cs:4,0.00724021662950192)
--(axis cs:4,0.0345008455343396);

\path [draw=color0, semithick]
(axis cs:5,-0.00216067805943182)
--(axis cs:5,0.0333085511529482);

\path [draw=color0, semithick]
(axis cs:6,0.00816087621620421)
--(axis cs:6,0.0393841061140945);

\path [draw=color0, semithick]
(axis cs:7,-0.00231674240297224)
--(axis cs:7,0.00832458763051024);

\path [draw=color0, semithick]
(axis cs:8,-0.002629967814537)
--(axis cs:8,0.0210984758836623);

\path [draw=color0, semithick]
(axis cs:9,-0.00405574281778054)
--(axis cs:9,0.0267186443369972);

\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 0.709690218797472
2 0.159795728748012
3 -0.00265218119467417
4 0.00724021662950192
5 -0.00216067805943182
6 0.00816087621620421
7 -0.00231674240297224
8 -0.002629967814537
9 -0.00405574281778054
};
\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 0.753548876258071
2 0.202483949225428
3 0.00956023576215485
4 0.0345008455343396
5 0.0333085511529482
6 0.0393841061140945
7 0.00832458763051024
8 0.0210984758836623
9 0.0267186443369972
};
\addlegendimage{semithick, color=color0};
\addlegendentry{Decision tree};
\end{axis}

\end{tikzpicture}