### 2022-01-07 figures/feature-importance/03-insightful-approachGB-final-2.tex

% This file was created by tikzplotlib v0.9.8.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.0666666666666667,0.466666666666667,0.2}

\begin{axis}[
height=1.4222438079424382in,
tick align=outside,
tick pos=left,
width=2.8444876158848764in,
x grid style={white!69.0196078431373!black},
xmajorgrids,
xmin=0.5, xmax=15.5,
xtick style={color=black},
xticklabels={,,},
y grid style={white!69.0196078431373!black},
ymajorgrids,
ymin=0, ymax=0.941178662009005,
yticklabels={,0,, 0.4,,0.8},
ytick style={color=black},
legend style={font=\footnotesize}
]
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:1.4,0.871178662009005);
\draw[draw=none,fill=color0] (axis cs:1.6,0) rectangle (axis cs:2.4,0.0217568511840037);
\draw[draw=none,fill=color0] (axis cs:2.6,0) rectangle (axis cs:3.4,0.00253604996390695);
\draw[draw=none,fill=color0] (axis cs:3.6,0) rectangle (axis cs:4.4,0.0192708196133984);
\draw[draw=none,fill=color0] (axis cs:4.6,0) rectangle (axis cs:5.4,0.0198922275506549);
\draw[draw=none,fill=color0] (axis cs:5.6,0) rectangle (axis cs:6.4,0.0260375424696281);
\draw[draw=none,fill=color0] (axis cs:6.6,0) rectangle (axis cs:7.4,0.0063356815006726);
\draw[draw=none,fill=color0] (axis cs:7.6,0) rectangle (axis cs:8.4,0.00319869517406271);
\draw[draw=none,fill=color0] (axis cs:8.6,0) rectangle (axis cs:9.4,0.00540195176007804);
\draw[draw=none,fill=color0] (axis cs:9.6,0) rectangle (axis cs:10.4,0.00695771640251456);
\draw[draw=none,fill=color0] (axis cs:10.6,0) rectangle (axis cs:11.4,0.00312270971606427);
\draw[draw=none,fill=color0] (axis cs:11.6,0) rectangle (axis cs:12.4,0.0030932299465922);
\draw[draw=none,fill=color0] (axis cs:12.6,0) rectangle (axis cs:13.4,0.0112178627094187);
\path [draw=color0, semithick]
(axis cs:1,0.854956014167526)
--(axis cs:1,0.887401309850484);

\path [draw=color0, semithick]
(axis cs:2,0.00907965200713417)
--(axis cs:2,0.0344340503608733);

\path [draw=color0, semithick]
(axis cs:3,-0.000182654889842383)
--(axis cs:3,0.00525475481765629);

\path [draw=color0, semithick]
(axis cs:4,0.00923661909617688)
--(axis cs:4,0.0293050201306199);

\path [draw=color0, semithick]
(axis cs:5,-0.00820788248180509)
--(axis cs:5,0.0479923375831149);

\path [draw=color0, semithick]
(axis cs:6,0.00143690786973603)
--(axis cs:6,0.0506381770695201);

\path [draw=color0, semithick]
(axis cs:7,-0.00149882818592621)
--(axis cs:7,0.0141701911872714);

\path [draw=color0, semithick]
(axis cs:8,-0.00213482608794717)
--(axis cs:8,0.00853221643607259);

\path [draw=color0, semithick]
(axis cs:9,-0.00260933890414728)
--(axis cs:9,0.0134132424243034);

\path [draw=color0, semithick]
(axis cs:10,-0.0038254976017747)
--(axis cs:10,0.0177409304068038);

\path [draw=color0, semithick]
(axis cs:11,-0.0044786262895489)
--(axis cs:11,0.0107240457216774);

\path [draw=color0, semithick]
(axis cs:12,-0.000395131606741384)
--(axis cs:12,0.00658159149992578);

\path [draw=color0, semithick]
(axis cs:13,-0.00413997684116445)
--(axis cs:13,0.0265757022600018);

\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 0.854956014167526
2 0.00907965200713417
3 -0.000182654889842383
4 0.00923661909617688
5 -0.00820788248180509
6 0.00143690786973603
7 -0.00149882818592621
8 -0.00213482608794717
9 -0.00260933890414728
10 -0.0038254976017747
11 -0.0044786262895489
12 -0.000395131606741384
13 -0.00413997684116445
};
\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 0.887401309850484
2 0.0344340503608733
3 0.00525475481765629
4 0.0293050201306199
5 0.0479923375831149
6 0.0506381770695201
7 0.0141701911872714
8 0.00853221643607259
9 0.0134132424243034
10 0.0177409304068038
11 0.0107240457216774
12 0.00658159149992578
13 0.0265757022600018
};
\addlegendimage{semithick, color=color0};
\addlegendentry{Gradient boost};
\end{axis}

\end{tikzpicture}