### 2022-01-07 figures/feature-importance/03-insightful-approachLOG-final-2.tex

% This file was created by tikzplotlib v0.9.8.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.533333333333333,0.8,0.933333333333333}

\begin{axis}[
height=1.4222438079424382in,
tick align=outside,
tick pos=left,
width=2.8444876158848764in,
x grid style={white!69.0196078431373!black},
xmajorgrids,
xmin=0.5, xmax=15.5,
xtick style={color=black},
xticklabels={,,},
y grid style={white!69.0196078431373!black},
ymajorgrids,
ymin=-0.623461141126395, ymax=0.530741648814479,
ytick style={color=black},
legend style={font=\footnotesize}
]
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:1.4,-0.253461141126395);
\draw[draw=none,fill=color0] (axis cs:1.6,0) rectangle (axis cs:2.4,0.0901171486976447);
\draw[draw=none,fill=color0] (axis cs:2.6,0) rectangle (axis cs:3.4,0.0863947483049856);
\draw[draw=none,fill=color0] (axis cs:3.6,0) rectangle (axis cs:4.4,-0.0772968684905453);
\draw[draw=none,fill=color0] (axis cs:4.6,0) rectangle (axis cs:5.4,0.0300554413526054);
\draw[draw=none,fill=color0] (axis cs:5.6,0) rectangle (axis cs:6.4,0.0644447825923199);
\draw[draw=none,fill=color0] (axis cs:6.6,0) rectangle (axis cs:7.4,0.000906315046421985);
\draw[draw=none,fill=color0] (axis cs:7.6,0) rectangle (axis cs:8.4,0.0527024964515279);
\draw[draw=none,fill=color0] (axis cs:8.6,0) rectangle (axis cs:9.4,-0.0642484568407123);
\draw[draw=none,fill=color0] (axis cs:9.6,0) rectangle (axis cs:10.4,-0.111007430848113);
\draw[draw=none,fill=color0] (axis cs:10.6,0) rectangle (axis cs:11.4,0.0148183210539744);
\draw[draw=none,fill=color0] (axis cs:11.6,0) rectangle (axis cs:12.4,-0.000553836590813862);
\draw[draw=none,fill=color0] (axis cs:12.6,0) rectangle (axis cs:13.4,0.100741648814479);
\draw[draw=none,fill=color0] (axis cs:13.6,0) rectangle (axis cs:14.4,0.0378766308558455);
\draw[draw=none,fill=color0] (axis cs:14.6,0) rectangle (axis cs:15.4,0.0181014244362775);
\path [draw=color0, semithick]
(axis cs:1,-0.587142604409583)
--(axis cs:1,0.0802203221567937);

\path [draw=color0, semithick]
(axis cs:2,0.0731623861401247)
--(axis cs:2,0.107071911255165);

\path [draw=color0, semithick]
(axis cs:3,0.0687951204204073)
--(axis cs:3,0.103994376189564);

\path [draw=color0, semithick]
(axis cs:4,-0.146499644473957)
--(axis cs:4,-0.0080940925071333);

\path [draw=color0, semithick]
(axis cs:5,-0.0313238410576597)
--(axis cs:5,0.0914347237628705);

\path [draw=color0, semithick]
(axis cs:6,0.0198594131562677)
--(axis cs:6,0.109030152028372);

\path [draw=color0, semithick]
(axis cs:7,-0.0499168743976032)
--(axis cs:7,0.0517295044904472);

\path [draw=color0, semithick]
(axis cs:8,0.00179966720486961)
--(axis cs:8,0.103605325698186);

\path [draw=color0, semithick]
(axis cs:9,-0.124803037939885)
--(axis cs:9,-0.00369387574154015);

\path [draw=color0, semithick]
(axis cs:10,-0.178052830807337)
--(axis cs:10,-0.0439620308888901);

\path [draw=color0, semithick]
(axis cs:11,-0.0315422676418746)
--(axis cs:11,0.0611789097498233);

\path [draw=color0, semithick]
(axis cs:12,-0.0592218306178204)
--(axis cs:12,0.0581141574361927);

\path [draw=color0, semithick]
(axis cs:13,0.0576301597775916)
--(axis cs:13,0.143853137851366);

\path [draw=color0, semithick]
(axis cs:14,-0.0196357412360911)
--(axis cs:14,0.0953890029477821);

\path [draw=color0, semithick]
(axis cs:15,-0.0217054681168758)
--(axis cs:15,0.0579083169894307);

\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 -0.587142604409583
2 0.0731623861401247
3 0.0687951204204073
4 -0.146499644473957
5 -0.0313238410576597
6 0.0198594131562677
7 -0.0499168743976032
8 0.00179966720486961
9 -0.124803037939885
10 -0.178052830807337
11 -0.0315422676418746
12 -0.0592218306178204
13 0.0576301597775916
14 -0.0196357412360911
15 -0.0217054681168758
};
\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 0.0802203221567937
2 0.107071911255165
3 0.103994376189564
4 -0.0080940925071333
5 0.0914347237628705
6 0.109030152028372
7 0.0517295044904472
8 0.103605325698186
9 -0.00369387574154015
10 -0.0439620308888901
11 0.0611789097498233
12 0.0581141574361927
13 0.143853137851366
14 0.0953890029477821
15 0.0579083169894307
};
\addlegendimage{semithick, color=color0};
\addlegendentry{Logistic regression};
\end{axis}

\end{tikzpicture}