### 2022-01-07 figures/feature-importance/03-insightful-approachPC-final-2.tex

% This file was created by tikzplotlib v0.9.8.
\begin{tikzpicture}

\begin{axis}[
height=1.4222438079424382in,
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
width=2.8444876158848764in,
x grid style={white!69.0196078431373!black},
xmajorgrids,
xmin=0.5, xmax=15.5,
xlabel={Principal components},
xtick style={color=black},
y grid style={white!69.0196078431373!black},
ymajorgrids,
ymin=0, ymax=0.227659111193136,
ytick style={color=black},
yticklabels={,0,, 0.2},
legend style={font=\footnotesize}
]
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:0.6,0) rectangle (axis cs:1.4,0.207659111193136);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:1.6,0) rectangle (axis cs:2.4,0.100931286761405);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:2.6,0) rectangle (axis cs:3.4,0.0552189869141694);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:3.6,0) rectangle (axis cs:4.4,0.0393144279067069);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:4.6,0) rectangle (axis cs:5.4,0.0359175173728584);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:5.6,0) rectangle (axis cs:6.4,0.0352100230844184);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:6.6,0) rectangle (axis cs:7.4,0.0281116402524651);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:7.6,0) rectangle (axis cs:8.4,0.0255858301182614);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:8.6,0) rectangle (axis cs:9.4,0.020895147833239);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:9.6,0) rectangle (axis cs:10.4,0.020208986578298);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:10.6,0) rectangle (axis cs:11.4,0.0173851755021279);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:11.6,0) rectangle (axis cs:12.4,0.0170402398481309);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:12.6,0) rectangle (axis cs:13.4,0.0150421767580223);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:13.6,0) rectangle (axis cs:14.4,0.0135123051552189);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:14.6,0) rectangle (axis cs:15.4,0.0123812036773514);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:15.6,0) rectangle (axis cs:16.4,0.0121461292946794);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:16.6,0) rectangle (axis cs:17.4,0.0113049221584044);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:17.6,0) rectangle (axis cs:18.4,0.0105679286164572);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:18.6,0) rectangle (axis cs:19.4,0.0099240637848415);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:19.6,0) rectangle (axis cs:20.4,0.00947939547970391);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:20.6,0) rectangle (axis cs:21.4,0.00880748003307345);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:21.6,0) rectangle (axis cs:22.4,0.00854299753930025);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:22.6,0) rectangle (axis cs:23.4,0.00819923337711622);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:23.6,0) rectangle (axis cs:24.4,0.00783441956089457);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:24.6,0) rectangle (axis cs:25.4,0.00725436756435554);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:25.6,0) rectangle (axis cs:26.4,0.00706748361834891);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:26.6,0) rectangle (axis cs:27.4,0.0067946334120975);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:27.6,0) rectangle (axis cs:28.4,0.00649269621923748);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:28.6,0) rectangle (axis cs:29.4,0.00631012021109268);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:29.6,0) rectangle (axis cs:30.4,0.0059935965924482);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:30.6,0) rectangle (axis cs:31.4,0.00591276082294336);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:31.6,0) rectangle (axis cs:32.4,0.00563442172532614);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:32.6,0) rectangle (axis cs:33.4,0.00539612814948874);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:33.6,0) rectangle (axis cs:34.4,0.00529456870611715);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:34.6,0) rectangle (axis cs:35.4,0.0052012922431412);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:35.6,0) rectangle (axis cs:36.4,0.00496741617032561);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:36.6,0) rectangle (axis cs:37.4,0.00483584771153227);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:37.6,0) rectangle (axis cs:38.4,0.00475069774706768);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:38.6,0) rectangle (axis cs:39.4,0.00453755427803973);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:39.6,0) rectangle (axis cs:40.4,0.00431289524873035);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:40.6,0) rectangle (axis cs:41.4,0.00413291864955501);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:41.6,0) rectangle (axis cs:42.4,0.00404009691245206);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:42.6,0) rectangle (axis cs:43.4,0.00384707175639724);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:43.6,0) rectangle (axis cs:44.4,0.00370794252319461);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:44.6,0) rectangle (axis cs:45.4,0.00360347785161737);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:45.6,0) rectangle (axis cs:46.4,0.00350293803678438);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:46.6,0) rectangle (axis cs:47.4,0.00335377406642866);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:47.6,0) rectangle (axis cs:48.4,0.00332782254644005);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:48.6,0) rectangle (axis cs:49.4,0.00325065460225723);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:49.6,0) rectangle (axis cs:50.4,0.00312401268859448);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:50.6,0) rectangle (axis cs:51.4,0.00307409451814868);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:51.6,0) rectangle (axis cs:52.4,0.00295493366839609);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:52.6,0) rectangle (axis cs:53.4,0.00283452933734321);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:53.6,0) rectangle (axis cs:54.4,0.00281415629292076);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:54.6,0) rectangle (axis cs:55.4,0.00277164449537654);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:55.6,0) rectangle (axis cs:56.4,0.0027301380386249);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:56.6,0) rectangle (axis cs:57.4,0.00269085035745271);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:57.6,0) rectangle (axis cs:58.4,0.00256884319342844);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:58.6,0) rectangle (axis cs:59.4,0.00252762617436217);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:59.6,0) rectangle (axis cs:60.4,0.00237101574205109);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:60.6,0) rectangle (axis cs:61.4,0.0023429271914207);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:61.6,0) rectangle (axis cs:62.4,0.00230359137012767);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:62.6,0) rectangle (axis cs:63.4,0.00221364734281885);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:63.6,0) rectangle (axis cs:64.4,0.0021462360088001);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:64.6,0) rectangle (axis cs:65.4,0.00211833683255068);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:65.6,0) rectangle (axis cs:66.4,0.00205592806762017);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:66.6,0) rectangle (axis cs:67.4,0.00203631661765791);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:67.6,0) rectangle (axis cs:68.4,0.00200913992503488);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:68.6,0) rectangle (axis cs:69.4,0.00198860732256562);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:69.6,0) rectangle (axis cs:70.4,0.00191877504746405);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:70.6,0) rectangle (axis cs:71.4,0.00183983787651402);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:71.6,0) rectangle (axis cs:72.4,0.00178967267072127);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:72.6,0) rectangle (axis cs:73.4,0.00176113054018146);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:73.6,0) rectangle (axis cs:74.4,0.00175587755437847);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:74.6,0) rectangle (axis cs:75.4,0.00166568926296783);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:75.6,0) rectangle (axis cs:76.4,0.00164865674220492);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:76.6,0) rectangle (axis cs:77.4,0.00163963642273892);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:77.6,0) rectangle (axis cs:78.4,0.00152246485297179);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:78.6,0) rectangle (axis cs:79.4,0.00150671318108692);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:79.6,0) rectangle (axis cs:80.4,0.00146762230969936);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:80.6,0) rectangle (axis cs:81.4,0.00144254604987905);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:81.6,0) rectangle (axis cs:82.4,0.00142852516970499);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:82.6,0) rectangle (axis cs:83.4,0.00140797817653715);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:83.6,0) rectangle (axis cs:84.4,0.00137907370730369);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:84.6,0) rectangle (axis cs:85.4,0.00137189411391102);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:85.6,0) rectangle (axis cs:86.4,0.00134894755604915);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:86.6,0) rectangle (axis cs:87.4,0.00131887269241828);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:87.6,0) rectangle (axis cs:88.4,0.00125161529911504);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:88.6,0) rectangle (axis cs:89.4,0.00123038981331776);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:89.6,0) rectangle (axis cs:90.4,0.00120293547721247);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:90.6,0) rectangle (axis cs:91.4,0.00118929296627723);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:91.6,0) rectangle (axis cs:92.4,0.00115421686183217);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:92.6,0) rectangle (axis cs:93.4,0.00113288923394035);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:93.6,0) rectangle (axis cs:94.4,0.00108899958168978);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:94.6,0) rectangle (axis cs:95.4,0.00107268732150737);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:95.6,0) rectangle (axis cs:96.4,0.00104142782755626);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:96.6,0) rectangle (axis cs:97.4,0.00101906762512476);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:97.6,0) rectangle (axis cs:98.4,0.00101077830297021);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:98.6,0) rectangle (axis cs:99.4,0.000989640667068959);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:99.6,0) rectangle (axis cs:100.4,0.000984439118691105);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:100.6,0) rectangle (axis cs:101.4,0.000960734989790538);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:101.6,0) rectangle (axis cs:102.4,0.000941461336335633);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:102.6,0) rectangle (axis cs:103.4,0.000928355012563315);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:103.6,0) rectangle (axis cs:104.4,0.000913315342059004);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:104.6,0) rectangle (axis cs:105.4,0.000897708116603778);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:105.6,0) rectangle (axis cs:106.4,0.000874661763686027);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:106.6,0) rectangle (axis cs:107.4,0.000854349918034593);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:107.6,0) rectangle (axis cs:108.4,0.000841201883286582);
\draw[draw=none,fill=white!53.3333333333333!black,fill opacity=0.5] (axis cs:108.6,0) rectangle (axis cs:109.4,0.000819094760963474);
\addlegendimage{semithick, color=white!53.3333333333333!black};
\addlegendentry{Explained variance};
\end{axis}

\end{tikzpicture}