### 2022-01-07 figures/feature-importance/03-insightful-approachRF-final-2.tex

% This file was created by tikzplotlib v0.9.8.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.866666666666667,0.8,0.466666666666667}

\begin{axis}[
height=1.4222438079424382in,
tick align=outside,
tick pos=left,
width=2.8444876158848764in,
x grid style={white!69.0196078431373!black},
xmajorgrids,
xmin=0.5, xmax=15.5,
xtick style={color=black},
xticklabels={,,},
y grid style={white!69.0196078431373!black},
ymajorgrids,
ymin=0, ymax=0.455941102264594,
ytick style={color=black},
legend style={font=\footnotesize}
]
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:1.4,0.405941102264594);
\draw[draw=none,fill=color0] (axis cs:1.6,0) rectangle (axis cs:2.4,0.0980875027579887);
\draw[draw=none,fill=color0] (axis cs:2.6,0) rectangle (axis cs:3.4,0.048776110224331);
\draw[draw=none,fill=color0] (axis cs:3.6,0) rectangle (axis cs:4.4,0.0307695146020032);
\draw[draw=none,fill=color0] (axis cs:4.6,0) rectangle (axis cs:5.4,0.0225284912793522);
\draw[draw=none,fill=color0] (axis cs:5.6,0) rectangle (axis cs:6.4,0.0376912406521808);
\draw[draw=none,fill=color0] (axis cs:6.6,0) rectangle (axis cs:7.4,0.0408639729682783);
\draw[draw=none,fill=color0] (axis cs:7.6,0) rectangle (axis cs:8.4,0.0453356735282065);
\draw[draw=none,fill=color0] (axis cs:8.6,0) rectangle (axis cs:9.4,0.0258685047848363);
\draw[draw=none,fill=color0] (axis cs:9.6,0) rectangle (axis cs:10.4,0.0341199103889856);
\draw[draw=none,fill=color0] (axis cs:10.6,0) rectangle (axis cs:11.4,0.0225669600689547);
\draw[draw=none,fill=color0] (axis cs:11.6,0) rectangle (axis cs:12.4,0.0197271830187896);
\draw[draw=none,fill=color0] (axis cs:12.6,0) rectangle (axis cs:13.4,0.0294260599312947);
\draw[draw=none,fill=color0] (axis cs:13.6,0) rectangle (axis cs:14.4,0.0100732082030255);
\draw[draw=none,fill=color0] (axis cs:14.6,0) rectangle (axis cs:15.4,0.00768012576296064);
\path [draw=color0, semithick]
(axis cs:1,0.38913349982482)
--(axis cs:1,0.422748704704367);

\path [draw=color0, semithick]
(axis cs:2,0.0842715544768568)
--(axis cs:2,0.111903451039121);

\path [draw=color0, semithick]
(axis cs:3,0.0397491095240302)
--(axis cs:3,0.0578031109246318);

\path [draw=color0, semithick]
(axis cs:4,0.0174838367514499)
--(axis cs:4,0.0440551924525565);

\path [draw=color0, semithick]
(axis cs:5,0.00917160744440918)
--(axis cs:5,0.0358853751142952);

\path [draw=color0, semithick]
(axis cs:6,0.0246759311379229)
--(axis cs:6,0.0507065501664387);

\path [draw=color0, semithick]
(axis cs:7,0.0296874637992993)
--(axis cs:7,0.0520404821372574);

\path [draw=color0, semithick]
(axis cs:8,0.0303547427614151)
--(axis cs:8,0.0603166042949978);

\path [draw=color0, semithick]
(axis cs:9,0.0147923986315111)
--(axis cs:9,0.0369446109381616);

\path [draw=color0, semithick]
(axis cs:10,0.0226833594028516)
--(axis cs:10,0.0455564613751197);

\path [draw=color0, semithick]
(axis cs:11,0.0134546981738569)
--(axis cs:11,0.0316792219640525);

\path [draw=color0, semithick]
(axis cs:12,0.0115712020878138)
--(axis cs:12,0.0278831639497654);

\path [draw=color0, semithick]
(axis cs:13,0.0188115961460041)
--(axis cs:13,0.0400405237165853);

\path [draw=color0, semithick]
(axis cs:14,0.00377725683628826)
--(axis cs:14,0.0163691595697627);

\path [draw=color0, semithick]
(axis cs:15,0.00393722295943178)
--(axis cs:15,0.0114230285664895);

\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 0.38913349982482
2 0.0842715544768568
3 0.0397491095240302
4 0.0174838367514499
5 0.00917160744440918
6 0.0246759311379229
7 0.0296874637992993
8 0.0303547427614151
9 0.0147923986315111
10 0.0226833594028516
11 0.0134546981738569
12 0.0115712020878138
13 0.0188115961460041
14 0.00377725683628826
15 0.00393722295943178
};
\addplot [semithick, color0, mark=-, mark size=4, mark options={solid}, only marks]
table {%
1 0.422748704704367
2 0.111903451039121
3 0.0578031109246318
4 0.0440551924525565
5 0.0358853751142952
6 0.0507065501664387
7 0.0520404821372574
8 0.0603166042949978
9 0.0369446109381616
10 0.0455564613751197
11 0.0316792219640525
12 0.0278831639497654
13 0.0400405237165853
14 0.0163691595697627
15 0.0114230285664895
};
\addlegendimage{semithick, color=color0};
\addlegendentry{Random forest};
\end{axis}

\end{tikzpicture}