### 2022-01-17 paper_v2.tex
Updates from Overleaf

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[superscriptaddress,unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{physics}
\usepackage{lipsum}
% \usepackage{subfig} %not compatible with subcaption
% \usepackage{braket}
\usepackage{siunitx}
\usepackage{color}
% \usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz,pgfplots}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}
\usepackage{makecell}
\captionsetup[table]{skip=10pt}
\pgfplotsset{compat=1.17}
\sisetup{separate-uncertainty}
\usepackage{hyperref} 
% \usepackage{longtable}

\usepackage{forest}
\forestset{
    .style={
        for tree={
            base=bottom,
            child anchor=north,
            align=center,
            s sep+=1cm,
    straight edge/.style={
        edge path={\noexpand\path[\forestoption{edge},thick,-{Latex}]
        (!u.parent anchor) -- (.child anchor);}
    },
    if n children={0}
        {tier=word, draw, thick, rectangle}
        {draw, diamond, thick, aspect=2},
    if n=1{%
        edge path={\noexpand\path[\forestoption{edge},thick,-{Latex}]
        (!u.parent anchor) -| (.child anchor) node[pos=.2, above] {Y};}
        }{
        edge path={\noexpand\path[\forestoption{edge},thick,-{Latex}]
        (!u.parent anchor) -| (.child anchor) node[pos=.2, above] {N};}
        }
        }
    }
}
\tikzset{
  green arrow/.style={
    midway,green,sloped,fill, minimum height=2cm, single arrow, single arrow head extend=.5cm, single arrow head indent=.25cm,xscale=0.3,yscale=0.15,
    allow upside down
  },
  yellow arrow/.style={
    midway,yellow,sloped,fill, minimum height=2cm, single arrow, single arrow head extend=.5cm, single arrow head indent=.25cm,xscale=0.3,yscale=0.15,
    allow upside down
  },
  black arrow/.style 2 args={-stealth, shorten >=#1, shorten <=#2},
  black arrow/.default={1mm}{1mm},
  tree box/.style={draw, rounded corners, inner sep=1em},
  node box/.style={white, draw=white, text=black, rectangle, rounded corners},
}

%\newcommand\abs[1]{\left|#1\right|}
%\newcommand\bra[1]{\left| #1 \right \rangle}
%\newcommand\ket[1]{\left \langle #1 \right |}

\newcommand{\oliver}[1]{\textcolor{violet}{#1}} 
\newcommand{\morten}[1]{\textcolor{green}{#1}}
\newcommand{\sebastian}[1]{\textcolor{cyan}{#1}}
\newcommand{\marianne}[1]{\textcolor{blue}{#1}}
\newcommand{\oyvind}[1]{\textcolor{maroon}{#1}}
\newcommand{\lasse}[1]{\textcolor{red}{#1}}

\begin{document}

%\title{Predicting Solid State Qubit Candidates}
\title{Predicting Solid State Material Platforms for Quantum Technologies}

\author{Oliver Lerst{\o}l Hebnes}
\affiliation{Your new job} 
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Marianne Etzelm\"uller Bathen}
\affiliation{Advanced Power Semiconductor Laboratory, ETH Z{\"u}rich, 8092  Z{\"u}rich,  Switzerland}
\affiliation{Department of Physics and Center for Materials Science and Nanotechnology, University of Oslo, N-0316 Oslo, Norway}

\author{{\O}yvind Sigmundson Sch{\o}yen}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Sebastian G. Winther-Larsen}
\affiliation{Menon Economics, N-0369 Oslo, Norway}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Lasse Vines}
\affiliation{Department of Physics and Center for Materials Science and Nanotechnology, University of Oslo, N-0316 Oslo, Norway}

\author{Morten Hjorth-Jensen}
\affiliation{Department of Physics and Astronomy and Facility for Rare Ion Beams, Michigan State University, East Lansing, MI 48824, USA}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\begin{abstract}
Semiconductor materials provide a compelling platform for quantum technologies (QT), and the properties of a vast amount of materials can be found in databases containing information from both experimental and theoretical explorations. However, scouring these databases to find promising candidate materials for quantum technology applications among tens of thousands of contenders is a major challenge. Therefore, we have developed a framework for the automated discovery of semiconductor host platforms for QT using material informatics and machine learning methods.

Data extraction tools have been developed for a number of materials science databases resulting in a data set consisting of over $25.000$ materials. The information from the databases was converted into nearly $5000$ physics-informed features. Furthermore, we have developed and implemented three different approaches for the data mining process: \textit{The Ferrenti approach}, \textit{the extended Ferrenti approach}, and \textit{the empirical approach}. 
The three approaches define distinct training sets for the supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting.

We find that the Ferrenti approach and the extended Ferrenti approach predict a broad range of materials as suitable for quantum technologies. The restrictions set in the empirical approach, which relies exclusively on findings from the literature, predicted substantially fewer candidates, and provided a clear distinction between suitable and unsuitable candidates when comparing the two largest eigenvalues in the covariance matrix. All four machine learning (ML) methods applied to the training set derived using the empirical approach agreed on a set of $214$ predicted candidate materials.
All three approaches and all four ML methods agreed on a subset of $47$ eligible candidates %(to a probability of $>50 \ \%$) 
of $8$ elemental, $29$ binary, and $10$ tertiary compounds, and provide a basis for futher material explorations towards quantum technology.

\end{abstract}

\pacs{02.70.Ss, 31.15.A-, 31.15.bw, 71.15.-m, 73.21.La}

\maketitle

\section*{Introduction}
Quantum technologies (QT) based on solid state platforms are rapidly gaining in both success and notoriety. 
Among the promising applications that are already available we find important breakthroughs  such as in vivo sensing of magnetic fields in cells \cite{Lesage_2013}, secure communication over large distances by separation of entangled photons \cite{Ursin2007} and, notably, various quantum information processing prototypes and architectures.  
Quantum computers are in high demand to meet the increasing need for computing power to solve complex many-body problems that are beyond the scope of classical machines.   
Recent advances have highlighted the potential of quantum information processors to outperform classical systems. Indeed, Google's $53$ qubit quantum computer based on superconducting electronics solved a computational problem that was beyond the capabilities of a $200.000$ core supercomputer \cite{Arute_2019}. Most recently, IBM announced its $127$ qubit quantum processor~\cite{IBM2021}. Simultaneously, the concepts of entanglement and teleportation may eventually facilitate advanced quantum communication protocols such as quantum cryptography and the quantum internet, spurring further investigations into technologies based on quantum mechanics.

%Figure 1 introduction schematic 
\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/ht-workflow-new-2.png}
    \caption{Schematic of an example workflow in material informatics. Results from theory, experiment and computation are fed into material databases (arrow pointing to the center). A cycle involving material screening, machine learning and predictions leads to knowledge gain and ultimately applications in fields such as clean energy and quantum technology. 
    }
    \label{fig:ht-workflow}
\end{figure}

Several platforms are available for the development of quantum technologies, but the materials and fabrication technologies are less mature than those for, e.g., classical computers and sensors. 
An important concern in this context is that of scalability. 
For example, the best performing quantum computer prototypes available today rely on superconducting electronics, requiring millikelvin temperatures to operate, with challenging fabrication and the stability of interactions between qubits is an issue. Instead, semiconductors are emerging as a promising alternative platform, offering competitive characteristics combined with the possibility of room temperature operation and mature and scalable material processing and fabrication.

Quantum technologies based on semiconductors rely on either defects or quantum dots where the latter kind can be of the self-assembled or fabricated type. 
Semiconductor defects can act as single-photon emitters or spin centers and are compatible with the three main QT categories of computing, communication and sensing.  
These characteristics are most often found for the case of defects that introduce deep energy levels into the semiconductor band gap. So-called deep level defects can trap charge carriers in localized states that are essentially isolated from the surroundings, making them highly suitable for QT due to, e.g., indistinguishable single-photon emission and long spin coherence times. 
The most well-known quantum compatible defect is the negatively charged nitrogen-vacancy (NV) center in diamond \cite{Doherty_2013}, but silicon carbide (SiC) and the various quantum emitters therein are strong contenders especially for quantum communication purposes due to the favorable emission wavelength region in the near infrared coupled with more mature material processing and fabrication \cite{Bathen2021}. 
However, defect based QT is still in the early stages, and the issues left to address include identification of suitable host materials and candidate defects, and scalable and reproducible quantum device fabrication. 
Furthermore, we lack a complete understanding of the requirements for a semiconductor material to manifest quantum compatible properties,  
and the selection of known quantum compatible host materials is slim.

The majority of discoveries of QT compatible characteristics in semiconductors has so far happened by serendipity, and there is an urgent need for a better and more systematic understanding of which material requirements must be met for QT compatible characteristics (e.g. single-photon emission and manipulable spin states) to manifest. In this context, a framework for dedicated materials search and analysis is needed.

The fourth science paradigm of big data driven science presents a significant improvement over the third paradigm focusing on computational science, for instance via density functional theory calculations of different materials properties. Instead, the fourth paradigm introduces the potential of targeted search for promising material systems in which to expect QT compatible properties. 
Rather than searching through a host of signals for those that match our criteria, we are now able to \textit{predict} which materials and signatures we should target for more detailed studies, following the framework illustrated in Figure~\ref{fig:ht-workflow}. 
This is made feasible by the availability of databases containing material properties for a wide range of different systems. In this work, the data in question are provided by bulk density functional theory (DFT) calculations to obtain the ground state properties of different elements and compounds. Combined with machine learning methods we provide a path towards precise classification of qubit candidates. The inclusion of machine learning methods follows recent trends in applications of statistical learning, data science and machine learning for scientific discoveries, see for example Refs.~\cite{deiana2021,Carleo2019}.

In this work we provide a framework for the data mining and automated discovery of promising solid-state semiconductor hosts for quantum emitters and spin centers using targeted database search and machine learning methods combined with knowledge from the field. 
In this context, we do not distinguish between the specific mechanism giving rise to properties such as single-photon emission and long spin coherence times (e.g., semiconductor defects or quantum dots); instead, we attempt to target all materials that may accommodate the desired characteristics.  
The methodology developed herein can be modified for other material types and application areas provided that high quality databases containing relevant theoretical and/or experimental data are available.

The approach we have developed relies on data extraction from the databases Materials Project \cite{Jain2013,Jain2018}, the Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} and AFLOW-ML \cite{Isayev2017}. 
The Matminer Python library for data mining \cite{Ward2018} was then used for material analysis to featurize the extracted data. An important aspect of our work is the database building and pertinent development of the training sets for machine learning methods. We have developed three different approaches to data mining: (i) \emph{the Ferrenti approach} which is similar to that proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, (ii) 
%Marianne I have now changed augmented Ferrenti to extended Ferrenti 
\emph{the extended Ferrenti approach} and (iii) \emph{the empirical approach}. We find that the former two methodologies base their data extraction protocols on broad material descriptors, leading to large sets of potentially suitable candidates \cite{Mehta2019,Hastie2009}. The third approach, on the other hand, relies on including materials with experimentally proven advantageous characteristics in the training set and therefore yields a narrower set of possible candidates. The three training sets where then analyzed with the four supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting, and yielded $214$ predicted candidates that were common between all approaches and ML methods. 
%(to a $0.5$ cut-off)  
Example materials include ZnGeP$_2$, MgSe, CdS, BP, BC$_2$N, BP, Ge, GeC, InP, and InAs. 
Focused theoretical and experimental studies, using methods such as, e.g., density functional theory, photoluminescence spectroscopy and optically detected magnetic resonance, are needed to verify the predictions that quantum properties may manifest as a result of defects or nanostructures in the above listed materials.

\section*{Results}

\subsection*{Information Flow}

The information stream in this work can be regarded as many modular parts 
%connected in logical pieces. 
%MHJ What do we mean with logical pieces? 
that are connected. 
The initial step for gathering material data and building features is visualized by the  flowchart in Figure~\ref{fig:flowchart}. 
Initially, we start by extracting all entries in the Materials Project (MP) database that match a specific query. 
The MP database contains ground state properties of different materials that are computed using density functional theory (DFT). The DFT calculations in the database were performed using the Vienna ab initio simulation package (VASP) \cite{Kresse1996} and the Perdew-Burke-Ernzerhof (PBE) \cite{Perdew1996} exchange-correlation functional to calculate the electronic structure of the materials. 
We note that despite being immensely successful in describing a number of material properties, PBE is widely known for underestimating the band gap of semiconductors \cite{Freysoldt2014}. Therefore, not all properties predicted using the PBE functional are reliable, and the band gaps in particular cannot be trusted in absolute terms. Nonetheless, the functional is in wide use due to the combination of reasonable accuracy and high computational throughput, and is usually considered to be reliable for large-scale trends in semiconductor properties.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flow_chart_2.png}
    \caption{The data flow of the main project, starting from an initial MP query, and ending with a featurized data set with entries from several other databases. To limit the memory and computational usage, the data are partitioned into smaller subsets where the respective Pymatgen objects  
    (Pymatgen is a robust, open-source Python library for materials analysis \cite{pymatgen}) 
    are obtained through a query to be used in the following featurization steps. This process is repeated iteratively until all the data has been featurized. Abbreviations used are Materials Project (MP), density of states (DOS) and electronic band structure (band).
    }
    \label{fig:flowchart}
\end{figure}

The conditions for the initial MP query are that the materials must derive from 
%experimental data 
the Inorganic Crystal Structure Database (ICSD) database and have a band gap wider than $0.1$~eV to exclude metallic compounds. The ICSD is the world's largest database for completely identified inorganic crystal structures \cite{Allen1987,Zagorac2019}. 
%OH: corrected icsd citation to german one (different from nist)
In a parallel step, entries that are deemed similar to the entries from the initial query are extracted from the databases OQMD \cite{Saal2013,Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015}, AFLOW-ML \cite{Isayev2017} and the Citrination platform \cite{OMaraJordan2016MDIA}. The results of these steps are combined into a data set for further analysis.

Thereafter, we apply Matminer’s \cite{Ward2018} tools to generate thousands of features from the data. 
Before applying any machine learning algorithm, the data was transformed into a numerical representation that reflects the relationship between the input and output data. This transformation is known as generating descriptors 
or features, however, we will in this work adopt the name featurization. 
The open-source library of Matminer provides many tools to featurize existing features extracted from the Materials Project. 
%MB I moved the below to methods 
%To apply Matminer's featurization tools, we extend an existing implementation by \citeauthor{Breuck2021} \cite{Breuck2021}, which were used to generate a supervised machine learning framework called the MODnet. Their implementation provides featurization for a material's composition, structure and atomic sites. However, Matminer also provides featurization tools for a material's density of states (DOS) and band structure, therefore, we extend their implementation to facilitate such featurizations. 
A schematic visualizing the featurization process in Matminer is contained in Fig.~\ref{fig:flowchart} 
and focuses on a material's composition,  structure, atomic sites, density of states and band structure. 
The $39$ featurizers selected as material descriptors in this work are summarized and described in the Supplementary Material at \cite{supplementary}. The selection of features was kept rather wide as we do not precisely know which features best describe a solid state material platform for quantum technology.

%It is important to note that 
The constructed data sets encompass compounds formed by a plethora of combinations of surfaces, interfaces, nanostructures, compositions and structures, but this complexity is not necessarily reflected in the material descriptors. 
Furthermore, we have utilized data obtained from high-throughput density functional theory calculations. Indeed, there are possible errors associated with every step, starting from an initial calculation, adding of data in the database, gathering of data, featurization of data, preprocessing of data, data mining, and finally training a model and making a prediction. Unfortunately, if an error has happened in the first part of the process, the error follows the entire process and will get increasingly harder to detect. Therefore, we are dependent on that the Materials Project database contains data with high quality, but we note that some errors are likely present in our data set.

\subsection*{Data Mining}

After compiling a data set from the procedure outlined above, we must define a training set for the machine learning algorithms. This is challenging for two reasons.  
Firstly, although QT compatible properties are becoming increasingly well studied for the case of, e.g., nitrogen-vacancy centers in diamond, single-photon emitters in silicon carbide and quantum dot structures, relatively few candidate materials are known to be suitable. An additional consideration is that the physical mechanisms promoting favorable properties are not fully understood, neither for the case of defect- or QD-based quantum technologies. Secondly, defining materials as unsuitable candidates for QT has proven to be surprisingly intricate. 
We follow three separate procedures for defining training sets consisting of materials that are labeled as both suitable and unsuitable candidates for QT.

\subsubsection*{The Ferrenti approach}
The first approach to defining a training set is based on the criteria proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}.
They suggest a data mining process consisting of four stages by systematically evaluating the suitability of host materials from the Materials Project. In this framework, we label \textbf{suitable} candidates by the following steps: 
\begin{enumerate}
    \item Include materials that;
    \begin{itemize}
        \item contain elements with more than $50 \ \%$ natural abundance of spin zero isotopes,
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database, 
        and
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Pragmatically remove toxic, radioactive and otherwise ``difficult'' materials;
    \begin{itemize}
        \item exclude Th, U, Cd and Hg because they are radioactive and/or toxic in the most stable forms,
        \item exclude any rare-earth metals (because of the difficulty of obtaining pure materials free of isotopes with nuclear spin) and noble gases (due to the lack of stable solid phases),
        \item exclude transition metal elements with unpaired electrons like Fe and Ni because of their paramagnetism, Ru and Os are also removed because they only exist in the data as complex cluster structures. 
    \end{itemize}
    \item Include only materials with a band gap larger than $0.5$~eV calculated using DFT and the PBE functional. The value of $0.5$~eV was chosen to match that typically predicted for silicon by PBE-level DFT calculations. 
    \item Ensure that the energy above hull is less than $0.2$~eV per atom.
\end{enumerate}

These inclusion criteria are based on the work of \citeauthor{Weber2010} \cite{Weber2010} and targeted primarily at semiconductors that can host deep level defects with spin qubit capabilities. In this context, long spin coherence times are needed, and the possibility of having an environment free of nuclear spins and permanent magnetism is therefore desirable. Additionally, to obtain defect-based single-photon emitters with sharp zero-phonon lines and indistinguishable single-photon photon emission, non-polar materials are assumed to be preferable. 
%The inclusion criteria listed above are based on the work of \citeauthor{Weber2010} \cite{Weber2010}.

Transition metal elements are eliminated in this approach if they have unpaired electrons because the presence of permanent electric dipole moments may have a detrimental impact on the optical coherence of defect emission. In this context, also the symmetry of the crystalline host may be important. Therefore, only phases which take on non-polar space groups are considered in this approach. 
The energy above hull requirement is included to ensure that the selected compounds are thermodynamically stable.

Next, \textbf{unsuitable} candidates are labeled according to the reverse requirements of the above; as materials in the ICSD database \cite{Allen1987,Zagorac2019} 
that crystallize in polar space groups, are calculated to be magnetic and have a band gap larger than $0.1$~eV in the MP database (to exclude metals but include lower-band gap semiconductors). 
The resulting complete training set contains $1581$ materials where $35 \ \%$
are labeled as unsuitable and 
$65 \ \%$ as suitable for QT applications.

%The training set from the Ferrenti approach reveal that carbon in a diamond-like structure is classified as a suitable candidate, in good agreement with experimental observations, (see Table~\ref{tab:qt-materials}) and hence also with the candidates that were manually defined as suitable in the empirical approach.  Interestingly, in the training set from the Ferrenti approach, we find that carbon in two-dimensional graphite-like structures are labeled as suitable as well. All structures of silicon are also labeled as suitable together with one entry of silicon carbide. Note that this is the 3C polytype of SiC, meaning that the most well-established quantum compatible SiC polytype, 4H-SiC, is missing from the training set. Among other potentially suitable candidates we find ZnS, ZnSe, ZnO and ZnTe present in the training data.

\subsubsection*{The extended Ferrenti approach}
In the second approach, we adjust the former one in order to expand the training set beyond practical considerations. This approach is therefore named \emph{the extended Ferrenti approach}.

In the first approach we included criteria that were 
%not physically motivated, but instead 
to some extent based on a practical perspective, such as excluding toxic and radioactive elements. Such criteria are not necessarily related to what makes a material suitable for quantum technology - eventually it is up to the experimentalist to decide the feasibility of the material to be employed. 
Consequently, in the extended Ferrenti approach, we remove stage two from the approach above. Moreover, we include a few additional elements that have shown promising properties but were initially excluded due to the lack of spin zero isotopes.

The following steps constitute the process of choosing \textbf{suitable} candidates in the \emph{extended Ferrenti approach}:
\begin{enumerate}
    \item Include materials that; 
    \begin{itemize}
        \item contain elements where more than half have a natural abundance of spin zero isotopes, including Al, P, Ga, As, B and N, 
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database,
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Only keep materials that have a band gap larger than $1.5$~eV in the MP database. The higher band gap requirement (as compared to the Ferrenti approach) is included here to avoid including an unfeasible number of candidate materials in the training set. 
    \item Ensure that the energy above hull is less than $0.2$~eV per atom. 
\end{enumerate}

For \textbf{unsuitable} candidates, we implement the same strategy as for the first Ferrenti approach. The result is a somewhat unbalanced training set, with up to $75 \ \%$ of the materials found in the suitable group. However, the training set is $78 \ \%$ larger than for the Ferrenti approach consisting of  $2141$ suitable and $684$ unsuitable materials.

%MB I moved this down to not confuse the order of information 
% 
% In the training set of the extended Ferrenti approach, we find a single entry for each of SiC, Si, GaN, ZnS, GaP, AlAs and AlP, carbon in both diamond- and graphite-like structures, and AlN in three different configurations. The training set includes a larger variety of materials that are known to be quantum compatible as compared to the Ferrenti approach due to admitting more elements in the initial selection process. However, since we also included a more stringent band gap restriction of $1.5$~eV (to restrict the amount of materials in the training set), we find a more sparse representation of each known chemical formula present in the training set.

\subsubsection*{The empirical approach}
The third approach differs substantially from the first two approaches in terms of material labeling, as we apply knowledge from the field (see for instance Refs.~\cite{Atatuere2018,Zhang2020,Son2020,Toth2019,Bathen2021} for an overview) to pick our promising material host candidates.  
In other words, we restrict ourselves to materials where quantum compatible properties have been either experimentally demonstrated or theoretically predicted. 
We therefore name this scheme \emph{the empirical approach}. Due to the concern of having a training set that is too narrow in scope, we will also include some  materials that are promising in the sense that they have suitable properties for accommodating deep level defects that can potentially exhibit quantum effects - even though such effects may not yet have been shown. These material choices were motivated by the criteria and predictions for QT compatible deep level defects defined by \citeauthor{Weber2010} \cite{Weber2010}. 
Materials where properties like single-photon emission and spin manipulation have been observed but attributed to excitonic effects or quantum dots (formed by self-assembly or lithographic structuring) rather than being defect related will also be included in this context.

Table~\ref{tab:qt-materials} contains an overview of known semiconductor materials with demonstrated quantum compatible characteristics (e.g., single-photon emission and coherent spin manipulation). The table forms the basis for picking suitable candidates for the empirical approach. The properties being studied arise from mechanisms related to, e.g., point defects, bound excitons, and both self-assembled and lithographically structured quantum dots and  nanostructures such as 2D materials. 
Quantum emission signatures have been assigned to specific defects in both diamond and SiC, but for most other materials, secure identification of the responsible defects or structure related mechanism is still lacking.

%In the empirical approach, we include materials that are known to be quantum compatible as outlined in Table~\ref{tab:qt-materials} in the training set. Additionally, we also include materials matching the formulas ZnSe, AlP, GaP, AlAs, ZnTe, CdS and SiGe as these materials have been predicted to behave as suitable quantum hosts based on favorable properties such as a wide band gap and low spin-orbit coupling \cite{Weber2010,Hardy2019}.

The strategy for picking \textbf{suitable} candidates in the empirical approach can be summed up as;  
\begin{enumerate}
    \item Select candidate materials that match the formulas in  Table~\ref{tab:qt-materials}, or the formulas ZnSe, AlP, GaP, AlAs, ZnTe, CdS \cite{Weber2010} and SiGe \cite{Hardy2019}, as these materials have been predicted to behave as suitable quantum hosts based on favorable properties such as a wide band gap and low spin-orbit coupling.  
    \item Ensure that the candidates are present in the ICSD database.  
    \item Perform a manual screening for appropriate crystallographic structures. 
\end{enumerate}

\begin{table}[b]
    \centering 
    \caption{Overview of materials and defects that have been demonstrated to exhibit QT compatible characteristics such as single-photon emission and coherent spin manipulation.}
    \begin{tabular}{c|c|c|c}
    Material & Band gap (eV) & Defect candidates & References \\
    \hline
    Diamond  & $5.5$  & N$_\mathrm{C}V_\mathrm{C}$, Si$_\mathrm{C}V_\mathrm{C}$, Ge$_\mathrm{C}V_\mathrm{C}$ & \cite{Taylor2008,Balasubramanian_2009,Barclay2011,Gordon2013,Rogers_2014,Bhaskar_2018} \\ 
    SiC & $2.2$-$3.3$ & $V_\mathrm{Si}$, $V_\mathrm{Si}V_\mathrm{C}$, C$_\mathrm{Si}V_\mathrm{C}$, N$_\mathrm{C}V_\mathrm{Si}$ & \cite{Widmann2014,Christle_2015,Castelletto_2014,Zargaleh_2018}  \cite{Weber2010, Son2020, Falk2013} \\ 
    Si & $1.1$ & P, G, unidentified & \cite{Muhonen_2014,Durand_2020,Redjem2020} \\ 
    (2D) \textit{h}-BN & $6.0$ & Unidentified defects & \cite{Tran_2016,Tran_2016b,Hayee_2020} \\ 
    (2D) MoS$_2$, WSe$_2$, WS$_2$ & $<2.5$~eV & Bound excitons & \cite{Toth2019} \\
    ZnO & $3.4$ & Unidentified defects & \cite{Morfa2012} \\ 
    ZnS & $3.6$ (zincblende) & Unidentified defects & \cite{Stewart2019} \\ 
    GaAs & $1.4$ & Quantum dots & \cite{Bluhm2010} \\ 
    GaN & $3.4$ & Quantum dots, unidentified defects & \cite{Roux2017,Berhane2018} \\
    AlN & $6.0$ & Unidentified defects & \cite{Xue2020}\\
    \end{tabular}
    \label{tab:qt-materials}
\end{table}

After the first stage of picking candidates we are left with a list of $202$ matching formulas which includes $12$ entries that have a band gap of less than $0.4$~eV. These $12$ entries are calculated to be thermodynamically unstable in terms of the energy above hull, and will decompose into other materials in the list --- incidentally, the resulting structures all have calculated band gaps that are substantially above $0.5$~eV. We include all of the $202$ structures except for C (mp-$568410$) which has a PBE-computed band gap of $0.12$~eV; i.e., a metal according to AFLOW-ML and it is therefore labeled as an unsuitable candidate.

%% OH: Dirty details about suitable materials:
Entries matching the formulas C, SiC, BN, MoS$_2$, WSe$_2$ and WS$_2$ (quantum compatible characteristics have been demonstrated) were manually screened to see whether they have a matching structure to the respective candidates discussed earlier and summarized in Table~\ref{tab:qt-materials}. 
For carbon, we admit three-dimensional diamond-like structures as explicitly stated in the column tags from the Materials Project. Additionally, we find several two-dimensional structures of carbon with a large band gap ($>1.5$~eV) among the data. We add these as suitable candidates. Complex structures (e.g., C$_{28}$, C$_{48}$ and C$_{60}$) were moved to the test set in our machine learning studies, see discussion below. For SiC we admitted all entries which included the $2$H, $3$C, $4$H, $6$H and $15$R polytypes. Concerning BN, MoS$_2$, WSe$_2$ and WS$_2$, we only admit two-dimensional structures. For non-matching structures not mentioned so far, we move them to the test set to see if they will be predicted as suitable or not by the machine learning methods that are applied in a later stage.

The materials AlP, GaP, AlAs, ZnTe and CdS were manually screened for tetrahedrally coordinated structures, and have been included since \citeauthor{Weber2010} \cite{Weber2010} identified them as potentially promising candidates due to suitable material properties. 
We note that only tetrahedrally coordinated structures of the given formulas were included in the training set after imposing a band gap restriction of $0.5$~eV.

After the three screening steps in the empirical approach, a total of $187$ entries were labeled as suitable candidates for the training set. 
Notably, the candidates labeled as suitable in the empirical approach contain only compounds that are either elementary (unary) or binary. 
We want to avoid discriminating based on the number of elements in a compound, and therefore we remove the feature that describes the number of elements in a compound. 
Since the training set constituting suitable candidates in the third approach contains rather few entries, we add $400$ candidates that are labeled as \emph{unsuitable}. These are picked at random from the pool of unsuitable candidates from the two previous approaches, in addition to those that were marked as unsuitable during the manual screening process.

\subsubsection*{Comparing the approaches}
The three different approaches to data mining place a particular emphasis on each of their specific goals. The Ferrenti approach relies on choosing only elements with spin zero isotopes and compounds that crystallize in nonpolar spacegroups together with practical filters, while the extended Ferrenti approach allows a larger variety of elements and removes the practical reasons for excluding candidate materials. Thus, the first approach targets a more narrow prediction space than the second approach does. However, perhaps the most restricted approach is the empirical one. Since the variety of known suitable materials is substantially smaller than for the two other approaches, we would expect the empirical approach to provide an even narrower prediction space. %Here, one should keep in mind that ...

%MB copied from above  
Considering the materials present in the training sets of the first two approaches more closely, we find that  carbon in a diamond-like structure is classified as a suitable candidate in the Ferrenti approach, in good agreement with experimental observations (see Table~\ref{tab:qt-materials}) and hence also with the candidates that were manually defined as suitable in the empirical approach.   Interestingly, in the training set from the Ferrenti approach, we find that carbon in two-dimensional graphite-like structures are labeled as suitable as well. All structures of silicon are also labeled as suitable together with one entry of silicon carbide. Note that this is the $3$C polytype of SiC, meaning that the most well-established quantum compatible SiC polytype, $4$H-SiC, is missing from the training set. Among other potentially suitable candidates we find ZnS, ZnSe, ZnO and ZnTe present in the training data.  
%Extended Ferrenti 
In the training set of the extended Ferrenti approach, we find a single entry for each of SiC, Si, GaN, ZnS, GaP, AlAs and AlP, carbon in both diamond- and graphite-like structures, and AlN in three different configurations. The training set includes a larger variety of materials that are known to be quantum compatible as compared to the Ferrenti approach due to admitting more elements in the initial selection process. However, since we also included a more stringent band gap restriction of $1.5$~eV (to reduce the amount of materials in the training set), we find a more sparse representation of each known chemical formula present in the training set.

The training data for each approach is visualized in Figure~\ref{fig:parallel-coordinates-approaches} as parallel coordinate plots for a few selected features informed by the criteria proposed by \citeauthor{Weber2010} \cite{Weber2010}. Parallel coordinate schemes  \cite{Inselberga1990, Inselberg1985} represent a multi-dimensional data tuple as one polyline crossing a parallel axis. The selected features are found on the x-axis, while the y-axis shows the value of the data. Thus, parallel coordinate plots can turn complex many-dimensional data into a compact  two-dimensional representation. Due to possible data cluttering, the figure visualizes a random sample of each class (suitable or unsuitable) with an upper limit of $250$ per class with transparent lines. The green and red polylines represent suitable and unsuitable candidates, respectively.

The Ferrenti approach and the extended Ferrenti approach share several similarities. 
Interestingly, we observe that the entries in  Fig.~\ref{fig:parallel-coordinates-approaches} span roughly the same dimensions for the Ferrenti and extended Ferrenti approaches, despite the more restricted nature of the former.  
The two approaches' upper limits for ionic character and covalent range are very similar in the case of candidates labeled as both suitable and unsuitable.   
Moreover, all candidates that were labeled as unsuitable exhibit polar space groups (follows from the definition of the training sets), and the two approaches also share that suitable candidates can consist of up to five different elements.

The most striking difference is found for the empirical approach. The entries labeled as suitable candidates do not exhibit any magnetization, even though the materials crystallize in both polar and non-polar space groups --- contrarily to the two Ferrenti-based approaches. Moreover, we observe that the range of covalent radius and maximum ionic character span a substantially smaller parameter space in the case of the empirical approach than for the two former ones.  
Here, we can begin to deduce some trends, if the findings from the empirical approach are taken as guidance. Where the polarity of the material's space group may be less important than assumed in the Ferrenti based approaches, ionic character and covalent range are distinguished from the other features visualized in Fig.~\ref{fig:parallel-coordinates-approaches}. 
Identification of the features that promote quantum compatible  characteristics could be crucial for targeted quantum device design based on careful material selection, nanostructuring and defect engineering.

\begin{figure}[t] %!htp
    \centering
    \begin{subfigure}{0.7\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/01-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.7\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/02-augmented-ferrenti-approach-v2.pgf}
    \end{subfigure}
    \begin{subfigure}{0.7\textwidth}
        \centering
          \input{figures/parallel-coordinates-plots/03-insightful-approach-v2.pgf}
    \end{subfigure}
    %\vspace*{10mm}
    \caption{Parallel coordinate plots for the different approaches. To limit the data cluttering, we have randomly collected up to $250$ entries for each class and made the lines transparent. For the empirical approach, we have used all the $187$ suitable candidates. The axes show total magnetization (mag) taken from the Materials Project (MP) database along with the space group (SG), ionic character (ionic char), covalent range (cov range) as calculated from elemental properties, number of elements (num elements) and energy gap (Eg) extracted from MP.} 
    \label{fig:parallel-coordinates-approaches}
\end{figure}

\begin{figure}%[h] %!tbp
    \centering
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/01-ferrenti-approach-v3.pdf}
    \end{subfigure}%
    \begin{subfigure}{0.3\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/02-augmented-ferrenti-approach-v2.pdf}
    \end{subfigure}

    \begin{subfigure}{0.4\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/03-insightful-approach-v3.pdf}
    \end{subfigure}
    %\vspace*{-95mm}
    \caption{Two-dimensional scatter plots for the three different approaches. We have identified the two eigenvectors corresponding to the two largest eigenvalues of the covariance-matrix, that is, the two most important principal components (PC1 and PC2) of the initial data from the Materials Project query. Then, we have transformed the three training sets resulting from the three approaches and visualized them as scatter plots. Green squares display suitable candidates, along with the black (SiC), blue (diamond) and yellow (Si) symbols in the bottom panel, and red triangles represent unsuitable candidates.}
    \label{fig:2dscatterplotpca}
\end{figure}

%MB !! Keep this here or move below??  !!
To further illustrate the complexity of the training sets, we have identified the two largest eigenvalues of the covariance matrix \cite{Hastie2009} of the initial data set from the Materials Project, and transformed the training sets according to the corresponding two eigenvectors. The resulting scatter plots are depicted in Figure~\ref{fig:2dscatterplotpca}. 
Green squares and red triangles represent suitable and unsuitable candidates, respectively. Due to the complexity of reducing the large amount of features down to only two, both suitable and unsuitable candidates for the Ferrenti and the extended Ferrenti approaches are largely overlapping. 
The logic behind categorizing materials in  two classes (suitable and unsuitable) appears to not translate into a distinct separation, at least not in the representation of Fig.~\ref{fig:2dscatterplotpca} for the Ferrenti and extended Ferrenti approaches.  
Hence, using these approaches for predicting QT material hosts could prove challenging for any model that would try to glean a clear-cut boundary between materials that are and are not suitable for QT. 
%Nonetheless, we note that the unsuitable candidates appear to be restricted to a certain parameter space while the suitable candidates extend much further.  
We expect therefore that the two Ferrenti approaches could need supplementary dimensions for further distinguishing between the materials in the two categories (suitable and unsuitable), or that it may prove challenging to find a generalized model based on the training sets that were defined according to the above discussion.

In the case of the empirical approach, on the other hand, we can discern a trend where the upper left part of the figure is dominated by suitable candidates (green, blue, yellow and black symbols). 
Interestingly, we observe that different configurations of the famously quantum compatible materials of diamond (blue symbols), silicon carbide (black symbols) and silicon (yellow symbols) are grouped together but each material class is separated in its own region. 
The unsuitable candidates (red squares) are similarly restricted to the lower right corner and there appears to be a clear barrier between the two (suitable and unsuitable), albeit with some exceptions. This indicates that the empirical approach may be capable of separating materials based on their underlying properties. Thus, devising a logical framework for the data mining process is likely important in order to obtain a suitable training set for the machine learning methods.

%This paragraph is very diffuse and do not really say anything. I propsoe we remove it.
%It seems probable that the differences in Fig.~\ref{fig:2dscatterplotpca} can be attributed to the specific choices made when extracting data in each of the three approaches. Where the empirical approach takes advantage of the experimental and theoretical knowledge that is available in the literature, the Ferrenti and extended Ferrenti schemes rely on assumptions for which material properties are responsible for the desired quantum effects. An additional factor to consider is the use of practical considerations (e.g., toxic materials or the availability of spin zero isotopes) when restricting the parameter space for material choice. 
%Intriguingly, a much clearer separation is observed for the empirical approach, where no restrictions were imposed in terms of specific characteristics for the materials, as compared to the two others. 
%These observations constitute an important finding of our work, as we show that it is possible to distinguish between suitable and unsuitable material candidates for QT based on a principal components analysis. The manual approach for selecting suitable candidates based on experimental findings yields strong indications that a trend does exist and thus is possible to identify. 
%This will be an important goal of a future work. 
%Importantly, the methodology for selecting candidates as implemented in the empirical approach can be adapted also to other material systems where a specific set of properties is desired (and observed or predicted for a sufficient number of materials) but the specific mechanisms behind them are unknown.

%summary - do we need this? 
In summary, using various data extraction tools for a number of materials science databases, we have obtained a data set consisting of over $25.000$ materials. The information from this data set has then been converted into nearly $5000$ physics-informed features.  
After these steps with fine-tuning and selection of entries for our training sets we are now ready to apply the machine learning methods on the test set in order to classify materials as potential candidates for quantum technology applications.

\subsection*{Machine learning and principal component analysis}

Four well-tested machine learning algorithms were employed in this work: logistic regression, decision trees, and ensemble methods like random forests and gradient boosting \cite{Mehta2019,Hastie2009}. 
These four machine learning methods were applied to the training sets that were optimized using each of the three approaches (Ferrenti, extended Ferrenti and empirical) outlined above. This yielded twelve sets of materials that were predicted as suitable material hosts for QT applications. 
In the evaluation of the machine learning models for the three different approaches we apply a $5\times 5$ stratified cross-validation when searching for the optimal hyperparameter  combinations. 
\textbf{Principal components were found by ...} 
The extracted principal components after applying the gradient boosting model for the training sets from the Ferrenti and empirical approaches are shown in Figure~\ref{fig:PComponents}. 
Further details on the evaluation of the machine learning models and the results from the principal component analyses for all four ML models are shown in the Supplementary Material at \cite{supplementary}.

Fig.~\ref{fig:PComponents} displays the principal components as they are sorted in descending order by the explained variance. For the Ferrenti approach (Fig.~\ref{fig:PComponents}a) to reach the $95 \ \%$ accumulated explained variance, a total of $144$ principal components must be included. Importantly, we observe that the single most important feature for all models in the Ferrenti approach is the fifth principal component. Moreover, by selecting the highest values in this eigenvector, we find that the corresponding features originate from the DFT computed band gap of the elemental solids among the elements in the compound as calculated by Materials Agnostic Platform for Informatics and Exploration (MagPie). The second component's largest values correspond to the 
%electronegativity, ionic property, and covalence radius 
covalent radius, ionic character and packing efficiency 
among the elements in the composition. 
The data originates from elemental calculations from MagPie and are aggregated as either minimum, mean, standard deviation, or maximum. 
While the first principal component encompasses the largest explained variance, it does not provide any specific information on which features it represents.

\begin{figure}[t]
    \centering
    %\includegraphics[width=0.9\textwidth]{figures/PC_Ferrenti_intuitive.png}
    \includegraphics[width=0.9\textwidth]{figures/PCcomponents.pdf}
    \caption{Visualizing different parameters for the $25$ most principal components ranked in descending order by the explained variance for (a) the Ferrenti approach and (b) the empirical approach. Upper panels show the gradient boosting coefficients and the lower panels show the explained variance that is retained by choosing each of the eigenvectors. %\textit{Placeholder figure.} 
    }
    \label{fig:PComponents}
\end{figure}

%extended Ferrenti 
For the extended Ferrenti approach, we find that gradient boosting performs optimally at $93$ principal components with a balanced accuracy score of $0.85$. The interpretation of feature importance for the extended Ferrenti approach is substantially more difficult than in the Ferrenti approach. We find, for logistic regression and decision trees, that no feature is different than any other in the cross-validation due to a large variety of accuracy. However, we find that random forests and gradient boosting experience the fifth principal component as important. Similar to the Ferrenti approach, the corresponding features with the highest value for the fifth principal component originate from the DFT computed band gap of the elemental solids among the elements in the composition.

%empirical approach
In the case of the empirical approach, in contrast to the Ferrenti based ones, the decision trees and random forests models exhibit the best performance for just a few principal components, and experience a considerable degree of overfitting for larger PC values. 
Gradient boosting, in contrast to the case for the two Ferrenti based approaches, also experiences the best performance for just a few principal components.

The gradient boosting model performs optimally at $13$ principal components with a mean F$1$-score of $0.96$. The empirical approach differs in many aspects from the Ferrenti and extended Ferrenti approaches. Firstly, we find that the number of principal components necessary to obtain $95 \ \%$ variance is reduced to $103$ components, which is $41$ and $56$ less than the Ferrenti and extended Ferrenti approaches, respectively. Thus, the variance of the training set is found to be described with fewer principal components, indicating a simpler model.   
Secondly, we find that the first principal component is by far the most important feature for all models in the empirical approach, as visualized in Fig.~\ref{fig:PComponents}(b).  This is part of the reason why we experience a large accuracy for only a single feature (see also the Supplementary Material at \cite{supplementary} fur further details). The first principal component's corresponding features are challenging to explain due to small variations of values, but we find that it includes bond orientational parameters, coordination numbers, and the radial distribution function of a compound's crystal system. 
The standard deviation of the radial distribution function appears multiple times in the list of features and appears to be of particular importance. 
Thirdly, the empirical approach differs in how much explained variation is retained by the first component, which is $21 \ \%$, while it is $14 \ \%$ for the Ferrenti  and $11 \ \%$ for the extended Ferrenti approach. We find the difference striking considering that the approaches share the same ultimate goal, but where the training sets apparently exhibit large and significant variations.

Considering the principal components that are deemed important by the three approaches, we find several intriguing differences. The Ferrenti and extended Ferrenti approaches place particular value on the band gap, of both the predicted compounds and the elemental solids, as well as features such as the ionic or covalent character of the bonding to matter. This is interesting as these features were used as guidance in the initial selection process when forming the training sets. The machine learning models appear to recognize at least some of the expected features as important. Intriguingly, for the empirical approach, other properties dominate. While, e.g., the band gap may still be important, the ML models appear to be informed by other characteristics more related to symmetry and crystal structure, based on the repeated appearance of bond length, orientation and radial distribution in the first principal component. In this case, we did not guide the selection process in terms of specific properties, as was done in the Ferrenti based approaches. Instead, the ML models appear to be inferring that the common trend among the materials that are empirically known to be suitable for QT is related to more complex mechanisms in the crystal structure and bonding.

\subsection*{Predicting Suitable Material Hosts for Quantum Technology} 
%In this section, 
After training the machine learning models using the three data mining approaches the models were employed on the test sets. An overview of the number of predicted suitable candidates as a function of ML model confidence when the four models in an approach agree is summarized in Table~\ref{tab:probabilites}.

\begin{table}[b]
    \centering 
    \caption{Overview of the number of predicted suitable candidates as a function of ML model confidence when the four models in an approach agree. A cut-off value of $0.5$ represents the confidence of a coin flip, while $1.0$ is a fully confident prediction.}
    \begin{tabular}{c|c|c|c}
      & $0.50$ & $0.75$ & $0.85$ \\
     \hline
     Ferrenti approach &  $6804$ & $1784$ & $258$  \\
     Extended Ferrenti approach &  $9227$ & $4735$  & $2001$  \\ 
     Empirical approach & $214$ & $66$ & $9$ \\
     \hline
     All approaches together & $47$ & $6$ & 0 \\
    \end{tabular}
    \label{tab:probabilites}
\end{table}

\subsubsection*{The Ferrenti approach}

We first consider the machine learning classification of the test set derived in the Ferrenti approach. The number of candidates predicted by each of the four machine learning methods logistic regression, decision trees, random forests and gradient boosting is contained in Figure~6 of the Supplementary Material at \cite{supplementary} for each of the three approaches. 
%MEB We can take this out, I suppose, if we need more space, and put it in Supplementary (I put fig 5 there) 
From the training set of $23.623$ materials, the four methods consistently predict at least $11.000$ materials as promising candidates in the case of the Ferrenti approach. All four ML methods agree on a total of $6804$ suitable candidates, however, many of these materials are predicted with an incidence similar to that of a coin flip. If we were to raise the minimum confidence cut-off of a prediction to (e.g.) $0.7$ instead of $0.5$, the above ML algorithms would only agree on $1784$ suitable candidates. 
We find that all four ML methods admit almost all materials with a chemical formula matching the known suitable candidates (see  Table~\ref{tab:qt-materials}) that were present in the training set. 
This can allow materials with unfortunate structures to be labeled as suitable candidates by all of our ML methods. 
For example, the ML methods do not perpetuate the band gap restriction from the training set definition. 
Based on the restrictions we defined initially, we would expect that all materials with band gaps lower than $0.5$~eV would be eliminated 
%for a cut-off set to above $0.5$ 
after applying the ML algorithms.  
Although this is expected behavior based on the training set, the trend is not carried over to the predicted data. We find that many entries with band gaps lower than $0.5$~eV are predicted as both suitable and unsuitable candidates by all four ML methods employed herein, 
indicating that the exact cut-off value we set for the band gap ($0.5$~eV for the Ferrenti approach) is not recognized by the machine learning algorithms --- despite the PC analysis revealing the band gap as an important feature for the ML classification in the Ferrenti approach.  
%This cut-off value was imposed on the training data because small band gaps could make it challenging to facilitate deep level defects \cite{Weber2010}. 
On the other hand, due to the known underestimation of band gaps by the PBE functional, the band gaps could in reality be larger for many of the materials.

\subsubsection*{The extended Ferrenti approach}
Next, we turn to the more liberal extended Ferrenti approach containing a $78 \ \%$ larger 
%amount of materials 
data set 
than the Ferrenti approach.  
The four ML methods predict at least $13.000$ materials as suitable candidates out of $22.550$ materials in the training set, where they all agree on a total of $9227$ predicted suitable candidates. Comparing to the training set reveals that all of the unlabeled candidates that are known to be suitable (see Table~\ref{tab:qt-materials}) are, in fact, predicted as suitable candidates. Unfortunately, the ML methods also yield predictions that do not match with what one would expect from physics informed  considerations. 
All four ML algorithms confidently (to confidences of $0.83$ and $0.60$ for two different configurations) predict NaCl as a suitable candidate, which we believe is unlikely due to the strong electrostatic interactions between Na and Cl and the ionic character of their bonding (see, e.g., Ref.~\cite{Weber2010}).  
Indeed, single-photon emitters with sharp and bright zero-phonon lines are uncommon in materials with a strong ionic character. 
Moreover, despite enforcing an even more conservative band gap restriction of $1.5$~eV for the training set as compared to the Ferrenti approach, we still find that all four implemented ML methods predict suitable candidates that exhibit band gaps substantially lower than $0.5$~eV.
Clearly, the ML methods are not adhering rigidly to our imposed band gap restriction when classifying the materials. 
As we based this band gap restriction on assumptions for important material parameters in the QT context, the ML methods might be picking up on something we are not aware of.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.50\textwidth]{figures/vesta-images/cubic-5-structures.png}
    \includegraphics[width=0.23\textwidth]{figures/vesta-images/hex-CdSe-CuI.png}
    \includegraphics[width=0.16\textwidth]{figures/vesta-images/BC2N.png}
    \caption{Illustrations of the crystal structures of the $9$ materials predicted by the empirical approach to $0.85$ confidence. Three different symmetry categories are clearly observable: (a) cubic, (b) hexagonal and (c) \textit{weird} (orthorombic top, chalcopyrite bottom).    }
    \label{fig:vesta-empirical-0p85}
\end{figure}

\subsubsection*{The empirical approach}
The ML methods that were trained on the data extracted in the empirical approach predict substantially fewer candidates as compared to the other two approaches. 
A total of $842$, $1197$, $543$ and $596$ materials are classified as suitable candidates by logistic regression, decision trees, random forests and gradient boosting, respectively. All the four ML methods agree on a total of $214$ suitable candidates to $0.5$ confidence.   
Although this is comparable to the probability of a coin toss, the agreement of the four ML algorithms lends support to the suitability of the $214$ materials. 
Note that $51$ of these have a band gap of $0.5$~eV or smaller as reported in Materials Project from PBE-level DFT calculations.

Consider now the $9$ materials that were classified by the empirical approach as suitable to a confidence of $0.85$ or higher by all four ML models; BN, CdSe, BC$_2$N, InAs, CuI, and ZnCd$_3$Se$_4$. 
%include this? 
The crystal structures of these materials are schematically drawn in Figure~\ref{fig:vesta-empirical-0p85} and grouped into three categories according to whether they are (a) cubic, (b) hexagonal or (c) other. 
%  
The compound BN (mp-$1639$) was in fact present in the training data as a suitable candidate, and we believe this a partial explanation for why the models recognize BN  as a suitable candidate with high probability. Furthermore, two compositions of CdSe (the cubic mp-$2691$ and the hexagonal mp-$1070$) have been predicted as suitable, possibly as a consequence of the similar compound CdS being labeled as a suitable candidate in the training set. 
%The element S resides in the same group as Se, and 
The two compounds of CdSe share similar properties with calculated band gaps in the MP database of $0.5$~eV and $0.6$~eV, respectively.  
Comparing to experiment, ... 
%high quality nanocrystals of CdS have been synthesized and studied for use as quantum dots since the nanocrystals emit in the visible range with a direct band gap of $2.5$-$2.6$~eV \cite{BanerjeeR2000Eots,CelebiSerdar2007SaCo,Qutub2012}. 
%Bulk CdS has a slightly lower reported band gap of $2.4$~eV \cite{Qutub2012}.

In the case of BC$_2$N we find two compositions with the same chemical formula; the orthorhombically coordinated (mp-$629458$) and the chalcopyrite structured (mp-$1008523$) BC$_2$N. The former takes on a polar space group while the latter does not. The band gaps are listed as $1.9$~eV and $1.7$~eV, respectively, in the MP database. 
%BC$_2$N is known as heterodiamond and is a hybrid of diamond and BN. 
Both structures exhibit, as expected, strong covalent character and have been studied for applications as nanostructures for electronic devices \cite{Gao2017}, hydrogen storage \cite{Cai2017} and superhard materials \cite{Li2017, Jiang2020}. Importantly, the diamond-like structure of BC$_3$N was recently predicted as a prominent spin qubit material host \cite{Wang2020SpinQB}. By creating a boron (B) vacancy one can in theory obtain a defect center with similar properties to those found for the NV center in diamond. Whether this is also possible for BC$_2$N remains to be seen. We note that BC$_3$N was not represented in the MP database at the time of downloading and is therefore not included in our data set.

The compounds InAs (cubic, mp-$20305$), CuI (cubic, mp-$22895$ and hexagonal, mp-$569346$) and ZnCd$_3$Se$_4$ (cubic, mp-$1078597$) are listed in the MP database with band gaps of $0.3$, $1.2$, $1.2$ and $1.7$~eV, respectively. 
%, and exhibit comparable properties. 
To the best of our knowledge, ZnCd$_3$Se$_4$ has yet to be synthesized. 
Self-assembled InAs quantum dots are an exciting possible material to use in quantum technology \cite{Liu2018}. 
CuI has recently been synthesized as single-crystal epitaxial films and was shown to exhibit remarkable optoelectronic properties \cite{Ahn2020}.  Interestingly, the material exhibits a large ionic character with more similarities to some of the oxide compounds in the data set than, e.g., the materials that are more covalent in character like Si, SiC and  diamond.  
It seems that ionic character alone is not an obstacle for a material to be quantum compatible if the prediction of CuI proves to be merited. It is unknown at this time whether any potential quantum compatible properties of CuI would originate from deep level defects or nanostructuring (e.g., quantum dots). Note that NaCl was not predicted as suitable for QT in the empirical approach.

Lowering the cut-off requirement from $0.85$ to $0.75$ for all models results in $66$ candidate materials for the empirical approach.  
The full list of these $66$ listings is contained in the Supplementary Material at \cite{supplementary}. 
The list of predicted suitable candidates now also includes ternary compounds of the formula ABC$_2$, in addition to elemental and binary semiconductors, and the nine materials listed above. For the ABC$_2$ structures, the elements Ga, Cd, In and Zn can occupy the A-site, Cu, Sn, Ag and Ge take the B-site, while S, Se, Te, P or As may reside at the C site. Most of the predicted compounds include at least one toxic element with one exception: ZnGeP$_2$ (mp-$4524$) in a chalcopyrite-like structure with an indirect band gap of $1.2$~eV \cite{Zhang2015} as reported in the MP database. In comparison, the experimentally reported band gap is somewhat larger at $2.0$~eV \cite{Xing1989}. 
ZnGeP$_2$ crystallizes in a non-polar space group, possesses no magnetic moment, exhibits strongly covalent bonds, and has been reported as an excellent mid-IR transparent crystal material that is suitable for nonlinear optical applications \cite{Zhang2015}. Importantly, it is possible to integrate sources of photon quantum states based on nonlinear optics with ZnGeP$_2$ \cite{Caspani2017}. 
We therefore identify ZnGeP$_2$ as an eligible candidate material for QT, but it remains to be seen whether the candidate can facilitate, e.g., the isolated deep energy levels often associated with defects exhibiting quantum compatible properties, or instead be a candidate for nanostructure based QT.

Among the $66$ materials that all models in the empirical approach agreed on to a $0.75$ confidence level, we emphasize the presence of interesting materials like Ge (in three configurations mp-$137$, mp-$1067619$ and mp-$1198022$), GeC, BP and InP. Ge in the cubic structure (mp-$1198022$) shares many properties with Si and C in addition to the periodic column number. 
Ge has the highest hole mobility of all semiconductors at room temperature, and is therefore considered a key material for the process of extending the chip performance in classical computers beyond the limits imposed by miniaturization \cite{Scappucci2020}. Furthermore, like SiC, device design based on Ge can take advantage of the mature large-scale fabrication of silicon due to the material's comparable properties.  
Similar considerations could be made for the case of GeC. 
Data from the Materials Project database suggests that the cubic compound GeC (mp-$1002164$) is a covalently bonded semiconductor having a band gap of $1.8$~eV. 
Consequently, with SiC being widely known as a highly suitable host material for QT compatible defects, we encourage further research on GeC due to its similarities to SiC.

The compound BP (mp-$1479$ and mp-$1008559$) is present in our predictions in both the cubic and hexagonal structures. The indirect band gaps of the different structures are calculated at $1.5$~eV and $1.1$~eV, respectively, as reported in the MP database. Both configurations of BP are nonmagnetic and share several properties with the entries mentioned above; including ... BP is non-toxic and has been synthesized with a potential for large-scale production \cite{MukhanovVladimirA2016Umso}.

Lastly, we will discuss the prediction of InP (mp-$966800$) as a suitable candidate. The compound inhabits the hexagonal structure with corner-sharing InP$_4$ tetrahedra. InP is reported in the Materials Project to have a direct band gap of $0.5$~eV and is considered as one of the most promising candidates to compete with Cd- or Pb- based QDs for, e.g., display and lighting applications  \cite{Zhang2020a, Won2019}. 
We underline the possibility of using InP-based quantum dots for QT applications as well.

It is worth mentioning that all models agree on several oxides being potential candidates when the prediction cut-off is set low enough (but still above $0.5$). However, we find that almost all the oxide compounds fall between the decision boundary defining suitable and unsuitable candidates, and none are present in the list containing the $66$ suitable candidates ($0.75$ cut-off) in the Supplementary Material. Due to the labeling of ZnO as a suitable candidate in the training set, we believe that the boundary was shifted sufficiently to admit several oxides as suitable candidates. 
Removing ZnO from the training set for the empirical approach might result in fewer or even no oxide compounds being predicted above a suitability limit of $0.5$ and could be an interesting future pathway for investigation.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.85\textwidth]{figures/vesta-images/all-together-vesta-6-mat-all.png}
    \caption{Illustrations of the crystal structure of the $6$ materials predicted by all approaches to $0.75$ confidence. Three different symmetry categories are clearly observable: (a) cubic, (b) hexagonal and (c) tetragonal.   }
    \label{fig:vesta-all-0p75}
\end{figure}

Comparing to the work of \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, they suggests a list of $541$ viable hosts after the data mining procedure.  
Among these, we find only a single material present in our list of $66$ candidates predicted by the four ML models after training on the data set resulting from the empirical approach. This material is the nontoxic compound MgSe (mp-$10760$) which crystallizes in the rock-salt structure. Its band gap is $2.0$~eV in the MP database with the experimental value being reported from $2.45$~eV (optical band gap) \cite{Ubale2014} up to as large as $5.6$~eV \cite{SaumGeorge1959}. MgSe is notable for its available spin-zero isotopes in accordance with the criteria set by  \citeauthor{Weber2010} \cite{Weber2010} and \citeauthor{Ferrenti2020} Ref.~\cite{Ferrenti2020}. We believe that these criteria may favor defects acting as spin centers with qubit potential and MgSe is thus identified as a promising host material in this respect.   
%\oliver{Maybe also see if other compounds that were present in training set are present in their 541 hosts. }

\subsubsection*{Comparing the approaches}
Comparing the three approaches (see Table~\ref{tab:probabilites}), we observe  that the extended Ferrenti approach is the least restricted and admits the largest number of entries through to the predictions regardless of the confidence cut-off value. The Ferrenti approach also admits a large amount of materials through from the training set as compared to the empirical approach, but the resulting data are otherwise not very different from the extended version, whereas the empirical approach predicts fewer suitable candidates where manually verification is achievable. Manual verification through a literature survey will often not be possible, however, and perfecting automatic data mining and analysis is therefore an important goal of material informatics \cite{rickman2019}.

Despite notable differences, as mentioned above, there is substantial overlap of predicted materials between the three approaches. 
We find that $119$ of the $214$ candidates predicted as suitable by the empirical approach (when the cut-off is set to $0.5$) were also selected by all ML methods in the extended Ferrenti approach. 
Similarly, $78$ of the $214$ are also predicted as suitable by all models in the Ferrenti approach. All approaches and their corresponding ML methods agree on a total of $47$ potential candidates to $0.5$ confidence, where eight are elementary (unary), $29$ binary, and $10$ tertiary 
(see the Supplementary Material at \cite{supplementary} for the full list). 
Several interesting materials that were also discussed above for the predictions by the empirical approach are present among these $47$, including BP, CdSe, BP, GeC, InP and Ge. However, we also find certain materials that can be classified neither as semiconductors nor solid-state, such as P, I, N$_2$ and H$_2$. We note that none of these were admitted through for the empirical approach when the confidence cut-off was set to $0.75$ or higher.

Importantly, there are six materials that all approaches and methods agree on to a $0.75$ confidence level. These are ZnGeAs$_2$ (tetragonal mp-$4008$), CdSnP$_2$ (tetragonal mp-$5213$), GeSe (cubic mp-$10759$), InP (cubic mp-$20351$), InP (hexagonal mp-$966800$) and SiSn (cubic mp-$1009813$), and their crystal structures are visualized in Figure~\ref{fig:vesta-all-0p75}. 
Again, we can distinguish three groupings in crystal structure; cubic, hexagonal and tetragonal. 
We highlight these six compounds, along with the nine predicted by the empirical approach to $0.85$ confidence, as particularly interesting for future in-depth theoretical and experimental study.

\subsection*{Important features for QT}

%histograms 
\begin{figure}[t]
    \includegraphics[width=0.45\textwidth]{figures/histograms/new/MP_Eg_count_075.pdf}
    \includegraphics[width=0.45\textwidth]{figures/histograms/new/IonProperty_max ionic char_count_075.pdf}
    \includegraphics[width=0.45\textwidth]{figures/histograms/new/GeneralizedRDF_std_dev Gaussian center=1.0 width=1.0_count_075.pdf}
    \includegraphics[width=0.45\textwidth]{figures/histograms/new/AverageBondLength_std_dev Average bond length_count_075.pdf}
    \caption{Histograms of predicted suitable materials as a function of the (a) PBE-calculated band gap from Materials Project, (b) maximum ionic character, (c) standard deviation of the radial distribution function and (d) standard deviation of the average bond length. The total number of predicted materials is  $6804$ for the Ferrenti approach, $9227$ for the extended Ferrenti approach, and $214$ for the empirical approach. The Ferrenti and extended Ferrenti approaches refer to the left y-axis and the empirical approach to the right.  }
    \label{fig:histogram_new}
\end{figure}

%Turning to why the specific materials were chosen ... 
The analysis of the principal components extracted from the ML models revealed that principal component number five was dominant for both the Ferrenti and extended Ferrenti approach. This PC includes features related to the band gap and chemical environment, thus, the band gap criterion imposed in the training set selection is at least somewhat satisfied. On the other hand, the band gap related features were not important in the dominant PC in the empirical approach, i.e., PC number 1. However, Figure~\ref{fig:histogram_new}(a) contains histograms of the number of predicted suitable materials for all approaches as a function of MP band gap, and reveals that all approaches predicted a substantial amount of materials with a low band gap. Moreover, the distribution of materials that was identified as suitable is broader with respect to the band gap for the Ferrenti approaches as compared to that of the empirical approach; the empirical approach exhibits a sharper peak around $2.5$~eV.  
Coincidentally, the band gap of, e.g., 4$H$-SiC is usually computed in the range around $2.5$~eV at the PBE level.

%Took this out, not correct
%As mentioned above, the PBE functional employed by the Materials Project calculations is known to underestimate the band gap of many semiconductors \cite{Freysoldt2014}. 
%It is therefore possible that the band gap is erroneously not considered an important parameter by the ML methods - or, we may be overestimating the importance of this criterion. 
%Indeed, we did not find that the band gap was prominent in the principal components, although the band gap could be correlated with other more important features. 
%Thus, one can argue that the ML methods are finding other patterns that represent a better distinction between suitable and unsuitable candidates in the training sets, resulting in the band gap being redundant.
%In this context, we also note that the patterns recognized by the ML algorithms are challenging to discern. Thus, while we intend for the methods to predict materials for QT, we cannot be certain that this is happening. 
%However, the large degree admittance of materials present in the training set of all three approaches (Ferrenti, extended Ferrenti and empirical) through to the predictions indicates that at least some of the intended parameters are being recognized as important.

The secondmost important principal component in the Ferrenti approach, PC nr. 2, encompasses properties such as the covalent radius and maximum packing efficiency (see the Supplementary Material at \cite{supplementary} for the predicted material distributions of these features).  
The maximum ionic character is also deemed an important part of PC2. Intriguingly, as shown in Figure~\ref{fig:histogram_new}(b), the predicted candidates distribute over a broad range of ionicities for all three approaches - even peaking at a relatively high ionic character of $0.8$.

The machine learning models in the empirical approach consistently identified the first PC as the predominant one. Identifying the single most important feature in this PC proved challenging as it is the combined impact of several features that is important in PC1. 
Here, we find that the standard deviation of the radial distribution function (RDF) has a  particularly strong impact, as it appears four times in different forms in the top ten list over dominating features. One configuration of the standard deviation of the RDF is demonstrated in  Figure~\ref{fig:histogram_new}(c), with two others being included in the Supplementary Material at \cite{supplementary}.  Additionally, the standard deviation of the average bond length (see Figure~\ref{fig:histogram_new}d), site fingerprint of the chemical environment and bond orientation (see the Supplementary Material) are influential in guiding the ML algorithms upon classifying suitable and unsuitable materials for QT in the empirical approach.

The standard deviation (std) of the RDF and average bond length for the three approaches reveals important information about the criteria followed by the ML methods when selecting suitable QT candidates. We interpret this feature such that zero std in the RDF means that there is no variation in the radial symmetry throughout the material. Similarly, zero std in the average bond length would mean that all bonds are identical throughout the crystal. Intriguingly, the peaks in the distributions of the predicted materials are not found for perfectly symmetric materials with identical bond lengths; instead, some variation in the bond and wavefunction distributions is found to be optimal for a material to be suitable for QT.  
We note that the peaks seem to appear for moderate standard deviations in the RDF and average bond length, indicating that a certain degree of symmetry is likely necessary for a material to act as a good QT host. The exact degree and type of symmetry is still open to debate and should be studied further.

We note that out of all the feature distributions visualized herein, the largest difference between the Ferrenti approaches and the empirical approach is found for the standard deviation in the radial distribution function (see Fig.~\ref{fig:histogram_new}c). 
This reflects the observations from the principal component analysis where PC1 (most important for the empirical approach) is dominated by the RDF and similarly symmetry related features. For the empirical approach, there is a clear peak in the preferred std value for the RDF, while the Ferrenti and augmented Ferrenti approaches experience a more even distribution across a broader range.

Returning to the predicted materials above, an interesting observation can be made. None of the nine materials predicted by the empirical approach with $0.85$ confidence, nor of the six materials predicted by all approaches to $0.75$ confidence, are elemental. At most three different elements are present, but the emphasis is clearly on binary compounds. This resonates strongly with the observations from the principal component analysis: an optimal degree of crystalline order likely exists, where some variations throughout the crystal in (e.g.) bond length and symmetry, for a material to manifest QT compatible properties. Indeed, this is in contrast to the expectations that guided the formulation of the training sets in the Ferrenti and extended Ferrenti approaches. Where most previous works have highlighted features such as band gap, polarity and ionic character as vital for a semiconductor to be a good quantum host, our results instead reveal that local variations in the crystal structure related to symmetry and bond angles, length and orientation should be considered instead.

Considering our findings in light of the specific defect centers we know to be quantum compatible, we find a fascinating agreement. None of the known quantum emitters or spin centers appear in completely uniform systems. For example, in the symmetric materials of diamond and silicon, QT compatible defects are not intrinsic; neither the silicon vacancy in silicon nor the carbon vacancy in diamond have exhibited single-photon emission or controllable spin coherence. Instead, we find that QT effects often appear after impurities are introduced, as evidenced by the phosphorous impurity in silicon [REF] and the nitrogen-vacancy, germanium-vacancy, tin-vacancy an lead-vacancy centers in diamond [REF]. 
For the binary system of silicon carbide in different polytypes, on the other hand, intrinsic defects like the silicon vacancy and the divacancy are appropriate for our goals. 
While these trends have yet to be verified on a grander scale, our findings provide strong indications that local variations in the crystal structure are paramount for QT  compatible properties to manifest in a semiconductor host.

%NB

\section*{Discussion} 
We have developed data extraction tools for numerous material science databases, and constructed over $4800$ physics-informed features for a data set consisting of more than $25.000$ materials. Furthermore, three approaches to data mining for QT applications were developed; (i) the Ferrenti approach, (ii) the extended Ferrenti approach, and (iii) the empirical approach.  
Three distinct training sets were defined for four different machine learning algorithms (logistic regression, decision trees, random forests and gradient boosting).

We find that the formulation of the training sets for the Ferrenti and extended Ferrenti approaches was broad, and hence multiple dimensions would be required for distinguishing between materials labeled as suitable and unsuitable for QT in the two training sets. The stringent restrictions set for the empirical approach, on the other hand, proved to give a distinct difference between suitable and unsuitable candidates by visualizing the two main principal components of the covariance matrix. 
Moreover, the smaller number of materials predicted for the empirical approach permits manual verification. 
%Our results show that including impractical materials (e.g., toxic, radioactive and without spin-zero isotopes) in the training set is vital for defining a suitable training set. 
% MHJ paragraph not concluded \textbf{This manifests in the ML predictions as ...} 
%Such materials can be eliminated based on practical considerations at a later stage but should be included initially in order to provide the ML methods with optimal conditions to divine physics based trends.

In short, the findings presented herein firmly establish that material informatics (from data mining via featurization to machine learning predictions) is a viable and important route to new discoveries in important fields.  Herein, we focus on predicting new candidate materials to host single-photon emitters and spin centers for quantum technology applications, but the framework we have developed is suitable for other fields as well. 
Two main routes can be followed to exploit the results. One aspect is the pursuit of experimental verification of QT compatible effect in the materials predicted as suitable by the four ML methods. Another, perhaps even more important, is to use the features and trends identified during the data mining and prediction process to understand the distinct material characteristics that enable quantum effects to manifest. Thus can new discoveries in the field of quantum technologies be made.

\section*{Methods}

\subsection*{Databases}
The Materials Project \cite{Jain2013, Jain2018} is an open-source project based on the Vienna Ab Initio Package (VASP) \cite{Kresse1996}. The Perdew-Burke-Ernzerhof \cite{Perdew1996} (PBE) functional is used to calculate band structures, while for transition metals, a $+U$ correction is applied to correct for correlation effects \cite{Wang2006}. The project is known as the initiator of materials genomics and offers a variety of calculated properties for over one hundred thousand inorganic crystalline materials, with frequent updates and extensions. Data extraction from Materials Project was performed in december of $2020$ for the Ferrenti and extended Ferrenti approach, while in march $2021$ for the empirical approach. Therefore, the two former approaches includes $77$ more materials due to deprecated entries in the Materials Project database.

The Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015} contains thermodynamic and structural properties of more than $600.000$  materials. The calculations are performed with the VASP software whereas the electron exchange and correlation are described with the PBE functional. The $+U$ extension is included for several calculations considering specific elements \cite{Stevanovic2012}. Data extraction from OQMD was done in February of $2021$.

JARVIS-DFT \cite{Choudhary2020} is an open database based on the VASP software and consists of roughly $40.000$ three-dimensional materials using the vdW-DF-OptB88 (OPT) functional \cite{Thonhauser2007, Klimes2011}. Structures included in the database are originally taken from the Materials
%mhj: ref to materials project is missing
Project, and then re-optimized using the OPT functional. Finally, the combination of the OPT and modified Becke-Johnson (mBJ) functionals is used to obtain a representative band gap for each structure \cite{Choudhary2018a}. Data extraction from JARVIS-DFT was done in january of $2021$, were we utilized the version made available 30.04.2021 at \url{figshare.com}.

The AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} repository is an automatic software framework for the calculations of a wide range of inorganic material properties. They utilize the PBE functional within VASP
%with projector-extended wave (PAW) potentials 
to relax twice and optimize the ICSD-structure. Data-extraction was done in the period of January-February of $2021$.

AFLOW-ML \cite{Isayev2017} is an application programming interface (API) that uses machine learning to predict thermo-mechanical and electronic properties based on the chemical composition and atomic structure alone, which are denoted as \textit{fragment descriptors}. Initially, the API decides whether a given material is a metal or an insulator, where the latter is confirmed with an additional regression method to predict the band gap. The accuracy is validated by a five-fold cross-validation process for each ML method, where they report a $93 \ \%$ prediction success of their initial binary classification method. This work utilized the Property Labeled Material Fragments (PLMF) openly available at their website \url{aflowlib.org}. We extract the crystallographic information file (CIF) of the crystals from Materials Project, and then use this as input to AFLOW-ML, which then returns an anticipated band gap. The process was done during January of $2021$.

\subsection*{Material Informatics}  
Matminer \cite{Ward2018} is an open source toolkit for material analysis written in Python. Matminer provides modules to extract information from a wide variety of databases. Additionally, they provide the tools to construct possibly thousands of features from calculations based on a material's composition, structure and electronic properties from DFT calculations, and have frameworks for visualization and automatic machine learning. 
To apply Matminer's featurization tools, we extend an existing implementation by \citeauthor{Breuck2021} \cite{Breuck2021}, which was used to generate a supervised machine learning framework called the MODnet. The implementation by \citeauthor{Breuck2021} provides featurization for a material's composition, structure and atomic sites. However, Matminer also provides featurization tools for a material's density of states (DOS) and band structure. Therefore, we extend their implementation to facilitate such featurizations. 
The features selected for featurization herein are summarized in the Supplementary Material at \cite{supplementary}.

Pymatgen, a robust, open-source Python library for materials analysis \cite{pymatgen}, was also employed.

\subsection*{Machine Learning}

Machine learning represents the science of giving computers the ability to learn without being explicitly programmed. The idea is that there exist generic algorithms which can be used to find patterns in a broad class of data sets without having to write code specifically for each problem. The algorithm builds its own logic based on the data.

The approaches to machine learning are many, but are often split into two main categories, supervised and unsupervised. In supervised learning we know the answer to a problem, and let the computer deduce the logic behind it. On the other hand, unsupervised learning is a method for finding patterns and relationship in data sets without any prior knowledge of the system. Many researchers also operate with a third category, namely reinforcement learning. This is a paradigm of learning inspired by behavioral psychology, where learning is achieved by trial-and-error, solely from rewards and punishment. In this work our focus is on supervised learning only with labeled data for classification problems.

In this work we have applied four well-known and tested machine learning methods for classification problems, these are (see for example \cite{Hastie2009,Mehta2019} for discussions and applications):
\begin{enumerate}
    \item Logistic regression,
    \item Decision trees,
    \item Random forests,
    \item Gradient boosting.
\end{enumerate}
Logistic regression \cite{Hastie2009} is a simple and frequently used method for binary and multi-category classification problems. In addition to logistic regression, we have also applied and tested the predictions made by decision trees and ensemble methods like random forests and gradient boosting, the latter through the application of the computationally efficient  XGBoost library \cite{xgboost2016}. Gradient boosting and random forests use decision trees as weak learners and improve their predictability. For random forests this is implemented through a collection of randomized decision trees where a  subset of the features in the data sets are selected randomly when building a decision tree. Boosting methods like gradient boosting use decision trees as  weak learners and improve upon these by an iterative process that involves the estimation of the gradients of the cost/loss function \cite{Hastie2009}. Pure decision trees can easily lead to overfitting of the data under study, leading to a machine learning method that exhibits a high variance. Ensemble methods like random forests and gradient boosting on the other hand tend to soften the overfitting problem, resulting in both a small bias and a reduced variance of the employed method, see for example Refs.~\cite{Hastie2009,Mehta2019} for an in-depth discussion of the bias-variance trade-off in machine learning. Gradient boosting implemented through the  XGBoost library \cite{xgboost2016} is widely used by data scientists to achieve state-of-the-art results on many machine learning challenges.

\section*{Data availability} 
The data sets that support the findings of this study are available online at \cite{Ohebbi2021}.

\section*{Code availability} 
The codes employed to develop the machine learning results are available online at 
%https://github.com/ohebbi/predicting-solid-state-qubit-material-hosts 
\cite{Ohebbi2021}.
%Update with new release later).

\bibliography{apssamp}

\begin{acknowledgments}

The work of LV and MEB was supported by the Research Council of Norway and the University of Oslo through the frontier research project FUNDAMeNT (no. 251131, FriPro ToppForsk-program). 
%Some of the computations were performed on resources provided by UNINETT Sigma2 - the National Infrastructure for High Performance Computing and Data Storage in Norway.  
The work of MEB was supported by an ETH Zurich Postdoctoral Fellowship. 
The work of MHJ was supported by the U.S. Department of Energy,
Office of Science, office of Nuclear Physics under grant
No. DE-SC0021152 and U.S. National Science Foundation Grants
No. PHY-1404159 and PHY-2013047.

\end{acknowledgments}

\section*{Author contributions}
Add! 
%MEB, LV and MHJ conceived the main theme of the work. OH developed the programs and performed the bulk of the work. All authors have contributed to the writing of the paper and the discussion and analysis of the data.

\section*{Competing interests}
The authors declare no competing interests.

\section*{Additional information}
Correspondence and requests for materials should be addressed to ?

\end{document}