### 2022-02-27 paper_v3_cut.tex
Updates from Overleaf

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[superscriptaddress,unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{physics}
\usepackage{lipsum}
% \usepackage{subfig} %not compatible with subcaption
% \usepackage{braket}
\usepackage{siunitx}
\usepackage{color}
% \usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz,pgfplots}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}
\usepackage{makecell}
\captionsetup[table]{skip=10pt}
\pgfplotsset{compat=1.17}
\sisetup{separate-uncertainty}
\usepackage{hyperref} 
% \usepackage{longtable}
\usetikzlibrary{patterns,shapes.arrows, arrows.meta,bending, shapes,calc,fadings,decorations.pathreplacing,positioning,arrows.meta}
\pgfplotsset{compat=newest}

\usepackage{sansmath}
\tikzset{>=stealth,
OptimumStyle/.style={align=center,anchor=east,rotate=90,font=\scriptsize}
}
\pgfplotsset{%samples=101,
axis lines = left,
every axis plot/.append style={line width=2pt},
}
\usepackage{forest}
\forestset{
    .style={
        for tree={
            base=bottom,
            child anchor=north,
            align=center,
            s sep+=1cm,
    straight edge/.style={
        edge path={\noexpand\path[\forestoption{edge},thick,-{Latex}]
        (!u.parent anchor) -- (.child anchor);}
    },
    if n children={0}
        {tier=word, draw, thick, rectangle}
        {draw, diamond, thick, aspect=2},
    if n=1{%
        edge path={\noexpand\path[\forestoption{edge},thick,-{Latex}]
        (!u.parent anchor) -| (.child anchor) node[pos=.2, above] {Y};}
        }{
        edge path={\noexpand\path[\forestoption{edge},thick,-{Latex}]
        (!u.parent anchor) -| (.child anchor) node[pos=.2, above] {N};}
        }
        }
    }
}
\tikzset{
  green arrow/.style={
    midway,green,sloped,fill, minimum height=2cm, single arrow, single arrow head extend=.5cm, single arrow head indent=.25cm,xscale=0.3,yscale=0.15,
    allow upside down
  },
  yellow arrow/.style={
    midway,yellow,sloped,fill, minimum height=2cm, single arrow, single arrow head extend=.5cm, single arrow head indent=.25cm,xscale=0.3,yscale=0.15,
    allow upside down
  },
  black arrow/.style 2 args={-stealth, shorten >=#1, shorten <=#2},
  black arrow/.default={1mm}{1mm},
  tree box/.style={draw, rounded corners, inner sep=1em},
  node box/.style={white, draw=white, text=black, rectangle, rounded corners},
}

%\newcommand\abs[1]{\left|#1\right|}
%\newcommand\bra[1]{\left| #1 \right \rangle}
%\newcommand\ket[1]{\left \langle #1 \right |}

\newcommand{\oliver}[1]{\textcolor{violet}{#1}} 
\newcommand{\morten}[1]{\textcolor{green}{#1}}
\newcommand{\sebastian}[1]{\textcolor{cyan}{#1}}
\newcommand{\marianne}[1]{\textcolor{blue}{#1}}
\newcommand{\oyvind}[1]{\textcolor{maroon}{#1}}
\newcommand{\lasse}[1]{\textcolor{red}{#1}}

\begin{document}

\title{Predicting Solid State Material Platforms for Quantum Technologies}

\author{Oliver Lerst{\o}l Hebnes}
\affiliation{Sopra Steria, Information Technology and Services, N-4020 Stavanger, Norway}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Marianne Etzelm\"uller Bathen}
\affiliation{Advanced Power Semiconductor Laboratory, ETH Z{\"u}rich, 8092  Z{\"u}rich,  Switzerland}

\author{{\O}yvind Sigmundson Sch{\o}yen}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Sebastian G. Winther-Larsen}
\affiliation{Menon Economics, N-0369 Oslo, Norway}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\author{Lasse Vines}
\affiliation{Department of Physics and Center for Materials Science and Nanotechnology, University of Oslo, N-0316 Oslo, Norway}

\author{Morten Hjorth-Jensen}
\affiliation{Department of Physics and Astronomy and Facility for Rare Ion Beams, Michigan State University, East Lansing, MI 48824, USA}
\affiliation{Department of Physics and Center for Computing in Science Education, University of Oslo, N-0316 Oslo, Norway}

\begin{abstract}
%MB 
%Original: 337 words
%New: 256 words 
%Wanted: 150 words

%Semiconductor materials provide a compelling platform for quantum technologies (QT), and the properties of a vast amount of materials can be found in databases containing information from both experimental and theoretical explorations. However, searching these databases to find promising candidate materials for quantum technology applications among tens of thousands of contenders is a major challenge. Therefore, we have developed a framework for the automated discovery of semiconductor host platforms for QT using material informatics and machine learning methods.

%Data extraction tools have been developed for a number of materials science databases resulting in a dataset consisting of over $25.000$ materials. The information from the databases was converted into nearly $5000$ physics-informed features. Furthermore, we have developed and implemented three different approaches for the data labeling process: \textit{The Ferrenti approach}, \textit{the extended Ferrenti approach}, and \textit{the empirical approach}. The three approaches define distinct sets of labeled data for the supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting.

%We find that the Ferrenti and extended Ferrenti approaches predict a broad range of materials as suitable for quantum technologies. The restrictions set in the empirical approach, which relies exclusively on findings from the literature, gave substantially fewer candidates with a clear distinction between suitable and unsuitable candidates when comparing the two largest eigenvalues in the covariance matrix. In contrast to the machine learning (ML) methods based on the Ferrenti and extended Ferrenti approaches that focused on band gap and ionic character, the ML methods from the empirical approach highlighted features related to symmetry and crystal structure, including bond length, orientation and radial distribution, as influential when labeling a material as suitable for QT. All four machine learning (ML) methods applied to the dataset derived using the empirical approach agreed on a set of $214$ predicted candidate materials. All three approaches and all four machine learning methods agreed on a subset of $47$ eligible candidates 
%(to a probability of $>50 \ \%$) 
%of $8$ elemental, $29$ binary, and $10$ tertiary compounds, and provide a basis for further material explorations towards quantum technology.

Semiconductor materials provide a compelling platform for quantum technologies (QT), and the properties of a vast amount of materials can be found in databases containing information from both experimental and theoretical explorations. However, searching these databases to find promising candidate materials for quantum technology applications among tens of thousands of contenders is a major challenge. Therefore, we have developed a framework for the automated discovery of semiconductor host platforms for QT using material informatics and machine learning methods, resulting in a dataset consisting of over $25.000$ materials and nearly $5000$ physcis-informed features. Three approaches were devised to label data for the supervised machine learning (ML) methods logistic regression, decision trees, random forests and gradient boosting. 
We find that of the three, the empirical approach relying exclusively on findings from the literature predicted substantially fewer candidates than the other two approaches with a clear distinction between suitable and unsuitable candidates when comparing the two largest eigenvalues in the covariance matrix. 
In contrast to expectations from the literature and that found for the the Ferrenti and extended Ferrenti approaches focusing on band gap and ionic character, the ML methods from the empirical approach highlighted features related to symmetry and crystal structure, including bond length, orientation and radial distribution, as influential when labeling a material as suitable for QT. 
All three approaches and all four machine learning methods agreed on a subset of $47$ eligible candidates 
%(to a probability of $>50 \ \%$) 
of $8$ elemental, $29$ binary, and $10$ tertiary compounds, and provide a basis for further material explorations towards quantum technology.

\end{abstract}

\pacs{02.70.Ss, 31.15.A-, 31.15.bw, 71.15.-m, 73.21.La}

\maketitle

\section*{Introduction}
Quantum technologies (QT) based on solid state platforms are rapidly gaining in both success and notoriety. 
Among the promising applications that are already available we find important breakthroughs  such as in vivo sensing of magnetic fields in cells \cite{Lesage_2013}, secure communication over large distances by separation of entangled photons \cite{Ursin2007} and, notably, various quantum information processing prototypes and architectures.  
Quantum computers are in high demand to meet the increasing need for computing power to solve complex many-body problems that are beyond the scope of classical machines.   
Recent advances have highlighted the potential of quantum information processors to outperform classical systems. Indeed, Google's $53$ qubit quantum computer based on superconducting electronics solved a computational problem that was beyond the capabilities of a $200.000$ core supercomputer \cite{Arute_2019}. Most recently, IBM announced its $127$ qubit quantum processor~\cite{IBM2021}. Simultaneously, the concepts of entanglement and teleportation may eventually facilitate advanced quantum communication protocols such as quantum cryptography and the quantum internet, spurring further investigations into technologies based on quantum mechanics.

%Figure 1 introduction schematic 
\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/ht-workflow-new-2.png}
    \caption{Schematic of an example workflow in material informatics. Results from theory, experiment and computation are fed into material databases (arrow pointing to the center). A cycle involving material screening, machine learning and predictions leads to knowledge gain and ultimately applications in fields such as clean energy and quantum technology. 
    }
    \label{fig:ht-workflow}
\end{figure}

Several platforms are available for the development of quantum technologies, but the materials and fabrication technologies are less mature than those for, e.g., classical computers and sensors. 
An important concern in this context is that of scalability. 
For example, the best performing quantum computer prototypes available today rely on superconducting electronics that require millikelvin temperatures to operate, with the stability of interactions between qubits being an important issue. Instead, semiconductors are emerging as a promising alternative platform, offering competitive characteristics combined with the possibility of room temperature operation and mature and scalable material processing and fabrication.

Quantum technologies based on semiconductors rely on either defects or quantum dots where the latter can be of the self-assembled or nanostructured type \cite{Aharonovich_2016}. 
Semiconductor defects can act as single-photon emitters or spin centers and are compatible with the three main QT categories of computing, communication and sensing \cite{Awschalom_2018}. 
These characteristics are most often found for the case of defects that introduce deep energy levels into the semiconductor band gap \cite{Weber2010}. So-called deep level defects can trap charge carriers in localized states that are essentially isolated from the surroundings, making them highly suitable for QT due to, e.g., indistinguishable single-photon emission and long spin coherence times. 
The most well-known quantum compatible defect is the negatively charged nitrogen-vacancy (NV) center in diamond \cite{Doherty_2013}, but silicon carbide (SiC) and the various quantum emitters therein are strong contenders for quantum communication purposes especially due to the favorable emission wavelength region in the near infrared coupled with more mature material processing and fabrication (see, e.g., Refs.~\cite{Castelletto_2015,Son2020,Bathen2021}). 
However, semiconductor based QT is still in the early stages, and the issues left to address include identification of suitable host materials and candidate defects, and scalable and reproducible quantum device fabrication. 
Furthermore, we lack a complete understanding of the requirements for a semiconductor material to manifest quantum compatible properties,  
and the selection of known quantum compatible host materials is slim \cite{Atatuere2018,Zhang2020}.

The majority of discoveries of QT compatible characteristics in semiconductors has so far happened by serendipity, and there is an urgent need for a better and more systematic understanding of which material requirements must be met for QT compatible characteristics like single-photon emission and single spin control to manifest. In this context, a framework for dedicated materials search and analysis is needed.

The fourth science paradigm of big data driven science reveals the potential of targeted search for promising material systems in which to expect QT compatible properties. 
%, see for example Ref.~\cite{Acin2018}.
Rather than searching through a host of signals for those that match our criteria, we aim to \textit{predict} which materials and signatures we should target for more detailed studies, following the framework illustrated in Figure~\ref{fig:ht-workflow}. 
This is made feasible by the availability of databases containing material properties for a wide range of different systems. In this work, the data in question are provided by bulk density functional theory (DFT) calculations to obtain the ground state properties of different elements and compounds. Combined with machine learning methods we provide a path towards precise classification of candidate materials. The inclusion of machine learning methods follows recent trends in applications of statistical learning, data science and machine learning for scientific discoveries, see for example Refs.~\cite{deiana2021,Carleo2019}.

Here we provide a framework for the data mining and automated discovery of promising solid-state semiconductor hosts for quantum emitters and spin centers using targeted database search and machine learning methods combined with knowledge from the field. 
Analyzing the output of the machine learning methods reveals that, given a suitable initial set of labeled materials for training, testing and validation, it is possible to discern the physical mechanisms that govern a material's suitability for quantum applications.  
In this framework we do not distinguish between the specific mechanism giving rise to properties such as single-photon emission and long spin coherence times (e.g., semiconductor defects or quantum dots); instead, we attempt to target all materials that may accommodate the desired characteristics.  
The methodology developed herein can be modified for other material types and application areas provided that high quality databases containing relevant theoretical and/or experimental data are available.

The developed approach relies on data extraction from the databases Materials Project \cite{Jain2013,Jain2018}, the Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} and AFLOW-ML \cite{Isayev2017}. 
The Matminer Python library for data mining \cite{Ward2018} was then used for material analysis to featurize the extracted data. An important aspect of our work is the database building and pertinent development of the datasets for machine learning methods. We have developed three different approaches to data mining: (i) \emph{the Ferrenti approach} which is similar to that proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, (ii) \emph{the extended Ferrenti approach} and (iii) \emph{the empirical approach}. We find that the two first methodologies base their data extraction protocols on broad material descriptors, leading to large sets of potentially suitable candidates \cite{Mehta2019,Hastie2009}. The empirical approach, on the other hand, relies on including materials with experimentally proven advantageous characteristics in the training set and therefore yields a narrower set of possible candidates. The three resulting sets of labeled data were then analyzed with the four supervised machine learning methods logistic regression, decision trees, random forests and gradient boosting \cite{Hastie2009,Murphy2012}, and yielded $214$ predicted candidates that were common between all approaches and machine learning methods.  
Example materials include ZnGeP$_2$, CdS, BP, BC$_2$N, GeC and InP. Focused theoretical and experimental studies, using methods such as, e.g., density functional theory, photoluminescence spectroscopy and optically detected magnetic resonance, are needed to verify the predictions that quantum properties may manifest as a result of defects or nanostructures in the above listed materials. 
%Write better! 
Importantly, our findings also reveal which material properties are weighted by the ML methods upon predicting a material as suitable for QT applications, thereby opening up for new  discoveries in the field of quantum technologies.

\section*{Results}

\subsection*{Information Flow}

The information stream in this work can be regarded as many inter-connected modular parts. 
The initial step for gathering material data and building features is visualized by the outer flowchart in Figure~\ref{fig:flowchart} (Jupyter notebooks containing the full
workflow can be found at \cite{Ohebbi2021}).
We start by extracting all entries in the MP database that match a specific query. 
The MP database contains ground state properties of different materials that are computed using density functional theory. The DFT calculations in the database were performed using the Vienna {\em ab initio} simulation package (VASP) \cite{Kresse1996} and the Perdew-Burke-Ernzerhof (PBE) \cite{Perdew1996} exchange-correlation functional to calculate the electronic structure of the materials. 
We note that despite being immensely successful in describing a number of material properties, PBE is widely known for underestimating the band gap of semiconductors \cite{Freysoldt2014}. Therefore, not all properties predicted using the PBE functional are reliable, and the band gaps in particular cannot be trusted in absolute terms. Nonetheless, the functional is in wide use due to the combination of reasonable accuracy and high computational throughput, and is usually considered to be reliable for large-scale trends in semiconductor properties.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flow_chart_2.png}
    \caption{The project workflow starting from an initial MP query, and ending with a featurized dataset with entries from several other databases. 
    To limit the memory and computational usage, the data are partitioned into smaller subsets where the respective Pymatgen objects (Pymatgen is a robust, open-source Python library for materials analysis \cite{pymatgen}) are obtained through a query to be used in the following featurization steps. This process is repeated iteratively until all the data has been featurized. Abbreviations used are Materials Project (MP), density of states (DOS) and electronic band structure (band).
    }
    \label{fig:flowchart}
\end{figure}

The conditions for the initial MP query are that the materials must derive from the Inorganic Crystal Structure Database (ICSD) and have a band gap wider than $0.1$~eV to exclude metallic compounds. The ICSD is the world's largest database for completely identified inorganic crystal structures \cite{Allen1987,Zagorac2019}. In a parallel step, entries that are deemed similar to the entries from the initial query are extracted from the databases OQMD  \cite{Saal2013,Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015}, AFLOW-ML \cite{Isayev2017} and the Citrination platform \cite{OMaraJordan2016MDIA}. The results of these steps are combined into a dataset for further analysis.

After material extraction we apply tools from the open-source library Matminer \cite{Ward2018} to generate thousands of features from the data. We will refer to this process as featurization herein. A schematic visualization of the featurization process in Matminer is shown in  Fig.~\ref{fig:flowchart} 
and focuses on a material's composition,  structure, atomic sites, density of states and band structure. 
The $39$ featurizers (each generates several features) selected as material descriptors in this work are summarized and described in the Supplementary Material at \cite{supplementary}. The selection of features was kept rather wide as we do not want to make {\em ex ante} assumptions on which features best describe a solid state material platform for quantum technology.

The constructed dataset encompasses compounds formed by a plethora of combinations of surfaces, interfaces, nanostructures, compositions and structures, but this complexity is not necessarily reflected in the material descriptors. 
Furthermore, we have utilized data obtained from high-throughput density functional theory calculations. Indeed, there are possible errors associated with every step, starting from an initial calculation, adding of data in the database, gathering of data, featurization of data, preprocessing of data, data mining, and finally training a model and making a prediction. Unfortunately, if an error has happened in the first part of the process, the error follows the entire process and will get increasingly harder to detect. Thus, our results depend on the quality of data in the databases we employ.

\subsection*{Data Mining}
The complete dataset consists of $25.000$ materials. A subset of these materials will be labeled into either suitable or unsuitable candidates for QT while the remainder will stay unlabeled. The labeled data are then grouped into training, test and validation sets for the machine learning methods.

Several challenges accompany the labeling process of labeling materials according to their suitability for QT. 
Firstly, although QT compatible properties are becoming increasingly well studied for the case of, e.g., nitrogen-vacancy centers in diamond, single-photon emitters in silicon carbide and quantum dot structures \cite{Doherty_2013,Bathen2021,Aharonovich_2016}, relatively few candidate materials are known to be suitable \cite{Atatuere2018,Zhang2020}. An additional consideration is that the physical mechanisms promoting favorable properties are not fully understood, neither for the case of defect- or QD-based quantum technologies. Thirdly, defining materials as unsuitable candidates for QT has proven to be surprisingly intricate. 
We follow three separate procedures for labeling materials as suitable or unsuitable candidates for QT.

\subsubsection*{The Ferrenti approach}
The first approach to labeling a selection of materials in the dataset is based on the criteria proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}. 
They suggest a data mining process consisting of four stages by systematically evaluating the suitability of host materials taken from the Materials Project. In this framework, we label \textbf{suitable} candidates by the following steps: 
\begin{enumerate}
    \item Include materials that;
    \begin{itemize}
        \item contain elements with more than $50 \ \%$ natural abundance of spin zero isotopes,
        \item crystallize in non-polar space groups,
        \item are present in the ICSD database, 
        and
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Pragmatically remove toxic, radioactive and otherwise ``difficult'' materials;
    \begin{itemize}
        \item exclude Th, U, Cd and Hg because they are radioactive and/or toxic in the most stable forms,
        \item exclude any rare-earth metals (because of the difficulty of obtaining pure materials free of isotopes with nuclear spin) and noble gases (due to the lack of stable solid phases),
        \item exclude transition metal elements with unpaired electrons like Fe and Ni because of their paramagnetism; Ru and Os are also excluded because they only exist in the dataset as complex cluster structures. 
    \end{itemize}
    \item Include only materials with a band gap larger than $0.5$~eV calculated using DFT and the PBE functional. The value of $0.5$~eV was chosen to match that typically predicted for silicon by PBE-level DFT calculations. 
    \item Ensure that the energy above hull is less than $0.2$~eV per atom.
\end{enumerate}

These inclusion criteria are based on the work of \citeauthor{Weber2010} \cite{Weber2010} and targeted primarily at semiconductors that can host deep level defects with spin qubit capabilities. In this context, long spin coherence times are needed, and the possibility of having an environment free of nuclear spins and permanent magnetism is therefore desirable. 
Moreover, non-polar materials are assumed to be preferable to obtain defect based single-photon emitters with sharp zero-phonon lines and indistinguishable single-photon emission.  
Transition metal elements are eliminated if they have unpaired electrons because the presence of permanent electric dipole moments may have a detrimental impact on the optical coherence of defect emission. 
Finally, the energy above hull requirement is included to ensure that the selected compounds are thermodynamically stable.

Next, \textbf{unsuitable} candidates are labeled according to the reverse requirements of the above; as materials in the ICSD database \cite{Allen1987,Zagorac2019} 
that crystallize in polar space groups, are calculated to be magnetic and have a band gap larger than $0.1$~eV in the MP database (to exclude metals but include lower-band gap semiconductors). 
The resulting set of labeled materials contains $1581$ materials where $35 \ \%$
are labeled as unsuitable and 
$65 \ \%$ as suitable for QT applications.

\subsubsection*{The extended Ferrenti approach}
Here, we adjust the Ferrenti approach to expand the data labeling process beyond practical  considerations. The second approach is therefore named \emph{the extended Ferrenti approach} and involves removing stage two from the approach above. Moreover, we include a few additional elements that have shown promising properties but were initially excluded due to the lack of spin zero isotopes.

The following steps constitute the process of labeling \textbf{suitable} candidates in the \emph{extended Ferrenti approach}:
\begin{enumerate}
    \item Include materials that; 
    \begin{itemize}
        \item contain elements where more than half have a natural abundance of spin zero isotopes, including Al, P, Ga, As, B and N, 
        \item crystallize in non-polar space groups,
        \item are present in the ICSD,
        \item are calculated to be nonmagnetic. 
    \end{itemize}
    \item Only keep materials that have a band gap larger than $1.5$~eV in the MP database. The higher band gap requirement (as compared to the Ferrenti approach) is included here to avoid labeling an unfeasibly large number of materials. 
    \item Ensure that the energy above hull is less than $0.2$~eV per atom. 
\end{enumerate}

For \textbf{unsuitable} candidates, we implement the same strategy as for the first Ferrenti approach. The result is a somewhat unbalanced set of labeled materials, with up to $75 \ \%$ of the materials found in the suitable group. However, the labeled data encompasses $78 \ \%$ more materials than for the Ferrenti approach, consisting of $2141$ suitable and $684$ unsuitable materials.

\subsubsection*{The empirical approach}
The third approach differs substantially from the others in terms of material labeling, as we apply knowledge from the field (see for instance Refs.~\cite{Atatuere2018,Toth2019,Zhang2020,Son2020} for an overview) to pick our promising material host candidates.  
In other words, we restrict ourselves to materials where quantum compatible properties have been either experimentally demonstrated or theoretically predicted. 
We therefore name this scheme \emph{the empirical approach}. Due to the concern of ending up with too small datasets for training, testing and validation of the machine learning methods, we will also include some  materials that are promising in the sense that they have suitable properties for accommodating deep level defects that can potentially exhibit quantum effects --- even though such effects may not yet have been shown. These material choices were motivated by the criteria and predictions for QT compatible deep level defects defined by \citeauthor{Weber2010} \cite{Weber2010}. 
We have also included materials where properties like single-photon emission and spin manipulation have been observed but attributed to excitonic effects or quantum dots (formed by self-assembly or lithographic structuring) rather than being defect related.

Table~\ref{tab:qt-materials} contains an overview of known semiconductor materials with demonstrated quantum compatible characteristics. The table forms the basis for picking suitable candidates for the empirical approach. The properties being studied arise from mechanisms related to, e.g., point defects, bound excitons, and both self-assembled and lithographically structured quantum dots and nanostructures such as 2D materials. 
Quantum emission signatures have been assigned to specific defects in both diamond and SiC, but for most other materials, secure identification of the responsible defects or structure related mechanism is still lacking.

The strategy for picking \textbf{suitable} candidates in the empirical approach can be summed up as;  
\begin{enumerate}
    \item Select candidate materials that match the formulas in  Table~\ref{tab:qt-materials}, or the formulas ZnSe, AlP, GaP, AlAs, ZnTe, CdS \cite{Weber2010} and SiGe \cite{Hardy2019}, as these materials have been predicted to behave as suitable quantum hosts based on favorable properties such as a wide band gap and low spin-orbit coupling.  
    \item Ensure that the candidates are present in the ICSD.  
    \item Perform a manual screening for appropriate crystallographic structures. 
\end{enumerate}

\begin{table}[b]
    \centering 
    \caption{Overview of materials and defects that have been demonstrated to exhibit quantum  compatible characteristics such as single-photon emission and coherent spin manipulation.}
    \begin{tabular}{c|c|c|c}
    Material & Band gap (eV) & Defect candidates & References \\
    \hline
    Diamond  & $5.5$  & N$_\mathrm{C}V_\mathrm{C}$, Si$_\mathrm{C}V_\mathrm{C}$, Ge$_\mathrm{C}V_\mathrm{C}$ & \cite{Taylor2008,Balasubramanian_2009,Barclay2011,Gordon2013,Rogers_2014,Bhaskar_2018} \\ 
    SiC & $2.2$-$3.3$ & $V_\mathrm{Si}$, $V_\mathrm{Si}V_\mathrm{C}$, C$_\mathrm{Si}V_\mathrm{C}$, N$_\mathrm{C}V_\mathrm{Si}$ & \cite{Widmann2014,Christle_2015,Castelletto_2014,Zargaleh_2018}  \cite{Weber2010, Son2020, Falk2013} \\ 
    Si & $1.1$ & P, G, unidentified & \cite{Muhonen_2014,Durand_2020,Redjem2020} \\ 
    (2D) \textit{h}-BN & $6.0$ & Unidentified defects & \cite{Tran_2016,Tran_2016b,Hayee_2020} \\ 
    (2D) MoS$_2$, WSe$_2$, WS$_2$ & $<2.5$~eV & Bound excitons & \cite{Toth2019} \\
    ZnO & $3.4$ & Unidentified defects & \cite{Morfa2012} \\ 
    ZnS & $3.6$ (zincblende) & Unidentified defects & \cite{Stewart2019} \\ 
    GaAs & $1.4$ & Quantum dots & \cite{Bluhm2010} \\ 
    GaN & $3.4$ & Quantum dots, unidentified defects & \cite{Roux2017,Berhane2018} \\
    AlN & $6.0$ & Unidentified defects & \cite{Xue2020}\\
    \end{tabular}
    \label{tab:qt-materials}
\end{table}

After the first stage of picking candidates we are left with a list of $202$ matching formulas which includes $12$ entries that have a band gap of less than $0.4$~eV. These $12$ entries are calculated to be thermodynamically unstable in terms of the energy above hull, and will decompose into other materials in the list --- incidentally, the resulting structures all have calculated band gaps that are substantially larger than $0.5$~eV. We include all of the $202$ structures apart from C (mp-$568410$) which is a metal according to AFLOW-ML.

%% OH: Dirty details about suitable materials:
Entries matching the formulas C, SiC, BN, MoS$_2$, WSe$_2$ and WS$_2$ (quantum compatible characteristics have been demonstrated) were manually screened to see whether they have a matching structure to the respective candidates discussed earlier and summarized in Table~\ref{tab:qt-materials}. 
For carbon, we admit three-dimensional diamond-like structures as explicitly stated in the column tags from the Materials Project. Additionally, we find several two-dimensional structures of carbon with a large band gap ($>1.5$~eV) among the data. We add these as suitable candidates. Complex structures (e.g., C$_{28}$, C$_{48}$ and C$_{60}$) were moved to the test set in our machine learning studies. For SiC we admitted all entries which included the $2$H, $3$C, $4$H, $6$H and $15$R polytypes. Concerning BN, MoS$_2$, WSe$_2$ and WS$_2$, we only admit two-dimensional structures. For non-matching structures not mentioned so far, we move them to the test set to see if they will be predicted as suitable or not by the machine learning methods that are applied in a later stage.

The materials AlP, GaP, AlAs, ZnTe and CdS were manually screened for tetrahedrally coordinated structures, and have been included since \citeauthor{Weber2010} \cite{Weber2010} identified them as potentially promising candidates due to suitable material properties. 
We note that only tetrahedrally coordinated structures of the given formulas were labeled as suitable after imposing a band gap restriction of $0.5$~eV.

Following the three screening steps in the empirical approach, a total of $187$ entries were labeled as suitable candidates for the training set. 
Notably, the candidates labeled as suitable in the empirical approach contain only compounds that are either elementary (unary) or binary. 
Since the set of labeled data constituting suitable candidates in the empirical approach contains rather few entries, we add $400$ candidates that are labeled as \textbf{unsuitable}. These are picked at random from the pool of unsuitable candidates from the two previous approaches, in addition to those that were marked as unsuitable during the manual screening process.

\subsubsection*{Comparing the approaches}
The three different approaches to data mining place a particular emphasis on each of their specific goals. The Ferrenti approach relies on choosing only elements with spin zero isotopes and compounds that crystallize in nonpolar spacegroups together with practical filters, while the extended Ferrenti approach allows a larger variety of elements and removes the practical reasons for excluding candidate materials. Thus, the first approach targets a more narrow prediction space than the second approach does. However, the most restricted approach is the empirical one, relying exclusively on previous findings rather than assuming which material properties are important.

%MB copied from above  
%Ferrenti 
Considering the materials present in the labeled datasets of the first two approaches more closely, we find that carbon in a diamond-like structure is classified as a suitable candidate in the Ferrenti approach, in good agreement with experimental observations (see Table~\ref{tab:qt-materials}) and hence also with the candidates that were manually defined as suitable in the empirical approach.   Interestingly, in the labeled data from the Ferrenti approach, we find that carbon in two-dimensional graphite-like structures are marked as suitable as well. All structures of silicon are also labeled as suitable together with one entry of silicon carbide. Note that this is the $3$C polytype of SiC, meaning that the most well-established quantum compatible SiC polytype, $4$H-SiC, is missing. 
% from the training set. 
Among other potentially suitable candidates we find ZnS, ZnSe, ZnO and ZnTe present in the training data.  
%Extended Ferrenti 
In the labeled dataset of the extended Ferrenti approach, we find a single entry for each of SiC, Si, GaN, ZnS, GaP, AlAs and AlP, carbon in both diamond- and graphite-like structures, and AlN in three different configurations. The labeled dataset includes a larger variety of materials that are known to be quantum compatible as compared to the Ferrenti approach due to admitting more elements in the initial selection process. However, since we also included a more stringent band gap restriction of $1.5$~eV (to reduce the amount of materials in the training set), we find a more sparse representation of each known chemical formula present in the labeled dataset.

The labeled data for each approach is visualized in Figure~\ref{fig:parallel-coordinates-approaches} as parallel coordinate plots for a few selected features informed by the criteria proposed by \citeauthor{Weber2010} \cite{Weber2010}. Parallel coordinate schemes  \cite{Inselberga1990, Inselberg1985} represent a multi-dimensional data tuple as one polyline crossing a parallel axis. The selected features are found on the x-axis, while the y-axis shows the value of the data. Thus, parallel coordinate plots can turn complex many-dimensional data into a compact  two-dimensional representation. Due to possible data cluttering, the figure visualizes a random sample of each class (suitable or unsuitable) with an upper limit of $250$ per class with transparent lines. The green and red polylines represent suitable and unsuitable candidates, respectively.

\begin{figure}[t] %!htp
    \centering
    \begin{subfigure}{1\textwidth}
        \centering
        \scalebox{0.85}{\input{figures/parallel-coordinates-plots/01-ferrenti-approach-v2.pgf}}
    \end{subfigure}
    \begin{subfigure}{1\textwidth}
        \centering
          \scalebox{0.85}{\input{figures/parallel-coordinates-plots/02-extended-ferrenti-approach-v3.pgf}}
    \end{subfigure}
    \begin{subfigure}{1\textwidth}
        \centering
          \scalebox{0.85}{\input{figures/parallel-coordinates-plots/03-empirical-approach-v3.pgf}}
    \end{subfigure}
    \caption{Parallel coordinate plots for the different approaches. To limit the data cluttering, we have randomly collected up to $250$ entries for each class and made the lines transparent. For the empirical approach, we have used all the $187$ suitable candidates. The axes show total magnetization (mag) taken from the Materials Project (MP) database along with the space group (SG), ionic character (ionic char), covalent range (cov range) as calculated from elemental properties, number of elements (num elements) and energy gap (Eg) extracted from MP.} 
    \label{fig:parallel-coordinates-approaches}
\end{figure}

\begin{figure}%[h] %!tbp
    \centering
    \begin{subfigure}{0.35\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/01-ferrenti-approach.pdf}
    \end{subfigure}%
    \begin{subfigure}{0.35\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/02-augmented-ferrenti-approach.pdf}
    \end{subfigure}
    \begin{subfigure}{0.4\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/pca-2d-plots/03-insightful-approach.pdf}
    \end{subfigure}
    \caption{Two-dimensional scatter plots for the three different approaches. We have identified the two eigenvectors corresponding to the two largest eigenvalues of the covariance-matrix, that is, the two most important principal components of the initial data from the Materials Project query. Then, we have transformed the three datasets resulting from the three approaches and visualized them as scatter plots. Green squares display suitable candidates, along with the black (SiC), blue (diamond) and yellow (Si) symbols in the bottom panel, and red triangles represent unsuitable candidates.}
    \label{fig:2dscatterplotpca}
\end{figure}

The Ferrenti approach and the extended Ferrenti approach share several similarities. For example, the entries in  Fig.~\ref{fig:parallel-coordinates-approaches} span roughly the same dimensions for the Ferrenti and extended Ferrenti approaches, despite the more restricted nature of the former.  
The two approaches' upper limits for ionic character and covalent range are very similar in the case of candidates labeled as both suitable and unsuitable.   
Moreover, all candidates that were labeled as unsuitable exhibit polar space groups (follows from the definition of the training sets), and the two approaches also share that suitable candidates can consist of up to five different elements.

The most striking difference is found for the empirical approach. The entries labeled as suitable candidates do not exhibit any magnetization, and the materials crystallize in both polar and non-polar space groups --- contrarily to the two Ferrenti-based approaches. Moreover, we observe that the range of covalent radius and maximum ionic character span a substantially smaller parameter space in the case of the empirical approach than for the two former ones.  
Here, we can begin to deduce some trends, if the findings from the empirical approach are taken as guidance. Where the polarity of the material's space group may be less important than assumed in the Ferrenti based approaches, ionic character and covalent range are distinguished from the other features visualized in Fig.~\ref{fig:parallel-coordinates-approaches}. 
%Identification of the features that promote quantum compatible  characteristics could be crucial for targeted quantum device design based on careful material selection, nanostructuring and defect engineering.

To illustrate the composition of the labeled data we perform a principal component analysis (PCA) \cite{Jolliffe2002}. Principal component analysis falls under the category of unsupervised learning within machine learning methods. Here we apply PCA to the labeled datasets in order to figure out if are specific features that stand out. In its standard  form, PCA relates the variance of the features with the eigenvalues of the covariance matrix \cite{Jolliffe2002,Murphy2012,Hastie2009}. We identify the two largest eigenvalues of the covariance matrix \cite{Hastie2009} of the complete initial dataset from the Materials Project, and transform the three labeled datasets according to the corresponding two eigenvectors. In essence, we reduce the dimensionality of the labeled data in order to visualize them.
The result of this procedure is displayed in the scatter plots of Figure~\ref{fig:2dscatterplotpca}. Note that some minor differences between the three panels in Fig.~\ref{fig:2dscatterplotpca} may occur due to the process of 
removing the mean and scaling to unit variance.
%centering the data. 
Red triangles represent unsuitable candidates while colored symbols (green, blue, black and yellow) represent suitable candidates. 
Due to the complexity of reducing the large amount of features down to only two, both suitable and unsuitable candidates for the Ferrenti and the extended Ferrenti approaches are largely overlapping. 
The logic behind categorizing materials in two classes (suitable and unsuitable) appears to not translate into a distinct separation, at least not in the representation of  Fig.~\ref{fig:2dscatterplotpca} for the Ferrenti and extended Ferrenti approaches.  
Hence, using these approaches for predicting QT material hosts could prove challenging for any model that would try to glean a clear-cut boundary between materials that are and are not suitable for QT. 
We expect therefore that the two Ferrenti approaches could need supplementary dimensions for further distinguishing between the materials in the two categories (suitable and unsuitable), or that it may prove challenging to find a generalized model based on the training sets that were defined according to the above discussion.

In the case of the empirical approach, on the other hand, we can discern a trend where the upper left part of the figure is dominated by suitable candidates (green, blue, yellow and black symbols), while the unsuitable candidates (red triangles) are similarly restricted to the lower right corner, albeit with some exceptions. 
Interestingly, we observe that different configurations of the famously quantum compatible materials of diamond (blue), silicon carbide (black) and silicon (yellow) are grouped together but each material class is separated in its own region. 
This indicates that the empirical approach may be capable of separating materials based on their underlying properties. Thus, devising a logical framework for the data mining process is likely important in order to obtain suitable datasets for training, testing and validation of the machine learning methods.

%This paragraph is very diffuse and do not really say anything. I propsoe we remove it.
%It seems probable that the differences in Fig.~\ref{fig:2dscatterplotpca} can be attributed to the specific choices made when extracting data in each of the three approaches. Where the empirical approach takes advantage of the experimental and theoretical knowledge that is available in the literature, the Ferrenti and extended Ferrenti schemes rely on assumptions for which material properties are responsible for the desired quantum effects. An additional factor to consider is the use of practical considerations (e.g., toxic materials or the availability of spin zero isotopes) when restricting the parameter space for material choice. 
%Intriguingly, a much clearer separation is observed for the empirical approach, where no restrictions were imposed in terms of specific characteristics for the materials, as compared to the two others. 
%These observations constitute an important finding of our work, as we show that it is possible to distinguish between suitable and unsuitable material candidates for QT based on a principal components analysis. The manual approach for selecting suitable candidates based on experimental findings yields strong indications that a trend does exist and thus is possible to identify. 
%This will be an important goal of a future work. 
%Importantly, the methodology for selecting candidates as implemented in the empirical approach can be adapted also to other material systems where a specific set of properties is desired (and observed or predicted for a sufficient number of materials) but the specific mechanisms behind them are unknown.

%summary - do we need this? 
%In summary, using various data extraction tools for a number of materials science databases, we have obtained a dataset consisting of over $25.000$ materials. The information from this dataset has then been converted into nearly $5000$ physics-informed features.  
After selecting and fine-tuning the entries for our labeled datasets we are now ready to apply the various ML methods discussed earlier. We split our labeled datasets into a so-called training set, a validation set and finally a test set which is not included in the training and validation process. This process will allow us to classify specific materials as potential candidates for quantum technology applications.

%Include a discussion on the difficulty of labeling unsuitable candidates. 
As mentioned above, a

\subsection*{Machine learning and principal component analysis}
In this work we have applied
four well-tested machine learning algorithms in order to classify specific materials as candidate systems for QT. These are logistic regression, decision trees, and ensemble methods like random forests and gradient boosting \cite{Mehta2019,Hastie2009,Murphy2012}. 
These four machine learning methods were applied to the training sets that were optimized using each of the three approaches (Ferrenti, extended Ferrenti and empirical) outlined above (recall that the labeled datasets are split into training, validation and test sets). This yielded twelve sets of materials that were predicted as suitable material hosts for QT applications.

In order to reduce the dimensionality while preserving most of the variance in the data,
%In order to deduce how the different material features are weighted by the four machine learning models when classifying materials into suitable and unsuitable candidates for QT,
we have performed a principal component analysis (PCA) \cite{Jolliffe2002} for each of the three approaches.
First, the eigenvalues of the covariance matrices of the three training sets are identified in order of decreasing value,  resulting in three distinct PCA models (as opposed to the case of Fig.~\ref{fig:2dscatterplotpca} where all panels are based on the same model).

 In the evaluation of the machine learning methods for the three different approaches we apply a $5\times 5$ stratified cross-validation \cite{Hastie2009} when searching for the optimal hyperparameter combinations. We note that all four machine learning methods have high evaluation metrics for the optimal hyperparameters. Further details on the evaluation of the machine learning methods and the results from the principal component analyses for all four machine learning methods are shown in the Supplementary Material at \cite{supplementary}.

%OH: Downsides. Could also be moved to methods?
%We observe that there are several potential pitfalls of using the chosen dimensionality reduction and machine learning techniques in this paper. PCA is a trivial algorithm that only accounts for linear correlation between features, where other algorithms, such as Kernel PCA, can facilitate non-linear dimensionality reduction but with higher computational demand. Additionally, by using a trivial PCA, we make an assumption of orthogonality, since the eigenvectors are orthogonal to each other. However, there might exist an non-orthogonal basis which captures the non-orthogonal data in the dataset better.

%Beneficial to use simple ML-methods for small datasets, since more advanced methods need (ALOT) more data to perform good
The three approaches define labeled data of varying size, where the smallest is from the empirical approach. For small datasets, it is proven beneficial to repeat the cross-validation, as this is a method that allows us to measure the stability of the predictions given against pertubations (i.e. few different entries) of the training data \cite{Beleites2008}.

Figure~\ref{fig:PComponents} visualizes different parameters for the most important principal components ranked in descending order by (a) the explained variance for the Ferrenti (upper panel) and empirical (lower panel) approach, and (b) the gradient boosting coefficients for the corresponding approaches. This differs from Fig.~\ref{fig:2dscatterplotpca}, where we visualize the two most important eigenvectors for all approaches that originates from the same covariance matrix. Note also that in Fig.~\ref{fig:2dscatterplotpca} we performed a PCA analysis on all data. In the ML analysis here the model development is done on the training and validation set. 
%Fig.~\ref{fig:PComponents}(a) displays the principal components as they are sorted in descending order by the explained variance. 
For the Ferrenti approach (Fig.~\ref{fig:PComponents}a) to reach the $95 \ \%$ accumulated explained variance, a total of $144$ principal components must be included. For the extended Ferrenti approach, we find that gradient boosting performs optimally at $93$.

In the case of the empirical approach and in contrast to the Ferrenti based approaches, decision trees and random forests exhibit the best performance for just a few principal components, and experience a considerable degree of overfitting 
%for larger principal component values. 
when involving more principal components. Gradient boosting, in contrast to the case for the two Ferrenti based approaches, also experiences the best performance for just a few principal components. 
\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \input{figures/feature-importance/01-ferrenti-approach-PC-v2.tex}
        \label{fig:01-fi-e}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \input{figures/feature-importance/01-ferrenti-approach-GB-v2.tex}
        \label{fig:01-fi-d}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \input{figures/feature-importance/03-empirical-approach-PC-v2.tex}
        \label{fig:03-fi-e}
        \subcaption{}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \input{figures/feature-importance/03-empirical-approach-GB-v2.tex}
        \label{fig:03-fi-d}
        \subcaption{}
    \end{subfigure}
    \caption{Visualizing different parameters for the $25$ most principal components ranked in descending order by (a) the explained variance for the Ferrenti (upper panel) and empirical (lower panel) approach, and (b) the gradient boosting coefficients for the corresponding approaches. Note well that in (a) we visualize only the results of the PCA analysis while in (b) there are results of the PCA reduced training and validation sets using gradient boosting \cite{Hastie2009,xgboost2016}. The latter shows that for the empirical approach most of the physics is represented by few features (see text for an in depth discussion). The results in (b) are similar to those obtained with logistic regression, decision trees and random forests.  
    }
    \label{fig:PComponents}
\end{figure}

Fig.~\ref{fig:PComponents}(b) reveals that the single most important feature for all ML methods in the Ferrenti approach is the fifth principal component (fifth largest eigenvalue). Moreover, by selecting the highest values in this eigenvector, we find that the corresponding features originate from the DFT computed band gap of the elemental solids among the elements in the compound as calculated by Materials Agnostic Platform for Informatics and Exploration (MagPie)~\cite{magpie}. 
The second most important principal component for the Ferrenti approach is the second one. 
The second component's largest values correspond to the 
%electronegativity, ionic property, and covalence radius 
covalent radius, ionic character and packing efficiency  
among the elements in the composition. 
The data originates from elemental calculations from MagPie and are aggregated as either minimum, mean, standard deviation, or maximum. 
While the first principal component encompasses the largest explained variance, it does not provide any specific information on which features it represents.

%extended Ferrenti 
The interpretation of feature importance for the extended Ferrenti approach is substantially more difficult than in the Ferrenti approach. We find, for logistic regression and decision trees, that no feature is different than any other in the cross-validation due to a large variety of accuracy. However, we find that random forests and gradient boosting experience the fifth principal component as important. Similar to the Ferrenti approach, the corresponding features with the highest value for the fifth principal component originate from the DFT computed band gap of the elemental solids among the elements in the composition.

%empirical approach
% OH: This paragraph is a duplicate from supplementary
The empirical approach differs in many aspects from the Ferrenti and extended Ferrenti approaches. 
%Firstly, for the empirical approach, the gradient boosting model performs optimally at $13$ principal components with a mean F$1$-score of $0.96$. 
Firstly, We find that the number of principal components necessary to obtain $95 \ \%$ variance is reduced to $103$ components, which is $41$ and $56$ less than the Ferrenti and extended Ferrenti approaches, respectively. Thus, the variance of the training set is found to be described with fewer principal components, indicating a simpler model.   
Secondly, the first principal component is by far the most important feature for all models in the empirical approach, as visualized in the lower panel of Fig.~\ref{fig:PComponents}(b). Similar conclusions are reached using logistic regression, decisions trees and random forests as classification methods, lending thus support to our main conclusions.

This is part of the reason why we experience a large accuracy for only a single feature (see also the Supplementary Material at \cite{supplementary} for further details). The first principal component's corresponding features is a complex combination of several material properties, but we find that it includes bond orientational parameters, coordination numbers, and the radial distribution function of a compound's crystal system. 
The standard deviation of the radial distribution function appears multiple times in the list of features and is of particular importance. 
Thirdly, the empirical approach differs in how much explained variation is retained by the first component, which is $21 \ \%$, while it is $14 \ \%$ for the Ferrenti  and $11 \ \%$ for the extended Ferrenti approach. We find the difference striking considering that the approaches share the same ultimate goal, but where the training sets apparently exhibit large and significant variations.

The principal components that are deemed important by the three approaches are distinct on multiple levels. The Ferrenti and extended Ferrenti approaches place particular value on the band gap, of both the predicted compounds and the elemental solids, as well as features such as the ionic or covalent character of the bonding. Notably, these features were used as guidance in the initial selection process when forming the training sets. Thus, the machine learning methods appear to preserve the criteria imposed in the Ferrenti approaches. For the empirical approach, we did not guide the selection process in terms of specific properties, thus, other properties dominate. Rather than focusing on, e.g., the band gap, the machine learning methods appear to be informed by other characteristics more related to symmetry and crystal structure, based on the repeated appearance of bond length, orientation and radial distribution in the first principal component. In this approach, the machine learning methods appear to be inferring that the common trend among the materials that are suitable for QT is related to more complex mechanisms in the crystal structure and bonding.

\subsection*{Predicting Suitable Material Hosts for Quantum Technology} 
%In this section, 
After training the machine learning approaches using parts of the data labeled during the three data mining approaches, the ML methods were employed on the test sets. 
The number of candidates predicted by each of the four machine learning methods logistic regression, decision trees, random forests and gradient boosting is visualized in the Supplementary Material at \cite{supplementary} for each of the three approaches. 
An overview of the number of predicted suitable candidates as a function of a given machine learning approach's confidence is summarized in Table~\ref{tab:probabilites}.

\begin{table}[t]
    \centering 
    \caption{Overview of the number of predicted suitable candidates as a function of a given machine learning approach's confidence when the four methods in an approach agree. A cut-off value of $0.5$ represents the confidence of a coin flip, while $1.0$ is a fully confident prediction.}
    \begin{tabular}{c|c|c|c}
      & $0.50$ & $0.75$ & $0.85$ \\
     \hline
     Ferrenti approach &  $6804$ & $1784$ & $258$  \\
     Extended Ferrenti approach &  $9227$ & $4735$  & $2001$  \\ 
     Empirical approach & $214$ & $66$ & $9$ \\
     \hline
     All approaches together & $47$ & $6$ & 0 \\
    \end{tabular}
    \label{tab:probabilites}
\end{table}

\subsubsection*{The Ferrenti approach}

We first consider the machine learning classification of the test set derived in the Ferrenti approach. 
From the labeled dataset of $23.623$ materials, the four methods consistently predict at least $11.000$ materials as promising candidates in the case of the Ferrenti approach. All four machine learning methods agree on a total of $6804$ suitable candidates, however, many of these materials are predicted with an incidence similar to that of a coin flip. If we were to raise the minimum confidence cut-off for a prediction to e.g. $0.75$ instead of $0.5$, the above machine learning algorithms would only agree on $1784$ suitable candidates. 
We find that all four machine learning methods admit almost all materials with a chemical formula matching the known suitable candidates (see Table~\ref{tab:qt-materials}) that were present in the labeled data. 
This can allow materials with unfortunate structures to be labeled as suitable candidates. %by all of our machine learning methods. 
For example, the machine learning methods do not perpetuate the band gap restriction from the training set definition. 
Based on the restrictions we defined initially, we would expect that all materials with band gaps lower than $0.5$~eV would be eliminated 
%for a cut-off set to above $0.5$ 
after applying the machine learning algorithms.  
Although this is expected behavior based on the data labeling process, the trend is not carried over to the predicted data. We find that many entries with band gaps lower than $0.5$~eV are predicted as both suitable and unsuitable candidates by all four machine learning methods employed herein, 
indicating that the exact cut-off value we set for the band gap ($0.5$~eV for the Ferrenti approach) is not recognized by the machine learning algorithms --- despite the principal component analysis revealing that the band gap is an important feature for the machine learning classification in the Ferrenti approach.  
On the other hand, due to the known underestimation of band gaps by the PBE functional, the band gaps could in reality be larger for many of the materials.

\subsubsection*{The extended Ferrenti approach}
Next, we turn to the more liberal extended Ferrenti approach containing a $78 \ \%$ larger 
amount of labeled data  
than the Ferrenti approach.  
The four machine learning methods predict at least $13.000$ materials as suitable candidates out of $22.550$ materials in the labeled dataset, where they all agree on a total of $9227$ predicted suitable candidates. Comparing to the labeled data reveals that all of the unlabeled candidates that are known to be suitable (see Table~\ref{tab:qt-materials}) are, in fact, predicted as suitable candidates. 
% rewrite?? 
However, the machine learning methods also yield predictions that do not match with the expectation that we are looking for a covalently bonded semiconductor to host a sharp and bright quantum emitter. 
% 
All four machine learning algorithms predict NaCl as a suitable candidate to confidences of $0.83$ and $0.60$ for two different configurations,  which was unexpected due to the strong electrostatic interactions between Na and Cl and the ionic character of their bonding (see, e.g., Ref.~\cite{Weber2010}).  
Indeed, single-photon emitters with sharp and bright zero-phonon lines are uncommon in materials with a strong ionic character. 
Moreover, despite enforcing an even more conservative band gap restriction of $1.5$~eV for the training set as compared to the Ferrenti approach, we still find that all four implemented machine learning methods predict suitable candidates that exhibit band gaps substantially lower than $0.5$~eV. 
Clearly, the machine learning methods are not adhering rigidly to our imposed band gap restriction when classifying the materials. 
As we based this band gap restriction on assumptions for important material parameters in the QT context, the machine learning methods might be picking up on something we did not anticipate.

\subsubsection*{The empirical approach}
The machine learning methods that were trained on the data extracted in the empirical approach predict substantially fewer candidates as compared to the other two approaches. 
A total of $842$, $1197$, $543$ and $596$ materials are classified as suitable candidates by logistic regression, decision trees, random forests and gradient boosting, respectively. All the four machine learning methods agree on a total of $214$ suitable candidates to $0.5$ confidence.   
Although this is comparable to the probability of a coin toss, the agreement of the four machine learning algorithms lends support to the suitability of the $214$ materials. 
Note that $51$ of these have a band gap of $0.5$~eV or smaller as reported in Materials Project from PBE-level DFT calculations.  Increasing the confidence cut-off to $0.75$ or $0.85$ yields $66$ or $9$ predicted suitable candidate materials in the empirical approach, respectively (see~Table~\ref{tab:probabilites}).

\begin{figure}[t]
    \centering
    \includegraphics[width=0.85\textwidth]{figures/vesta-images/fig6-all.png}
    \caption{Illustrations of the crystal systems that the $9$ materials predicted by the empirical approach to $0.85$ confidence, and the $6$ materials predicted by all approaches to $0.75$ confidence, may belong to. Four different symmetry categories are observed; cubic, hexagonal, tetragonal and orthorombic. The viewpoints for all materials are down along the $c$ axis.   }
    \label{fig:crystalsystems}
\end{figure}

%\begin{table}[b]
 %   \centering 
 %   \caption{Grouping of predicted materials (above $0.75$ confidence) into the four appearing crystal systems. The $9$ materials predicted by the empirical approach ($>0.85$) and the $6$ materials predicted by all approaches ($>0.75$) are considered. }
 %   \begin{tabular}{c|c c c c}
 %     & Cubic & Hexagonal & Tetragonal & Orthorombic \\
 %   \hline 
 %   Empirical approach & CuI (mp-$22895$) & CuI (mp-$569346$) & BC$_2$N (mp-$1008523$) & BC$_2$N (mp-$629458$) \\
 %   to $0.85$ confidence & CdSe (mp-$2691$) & CdSe (mp-$1070$) & & \\
 %    & BN (mp-$1639$) & & & \\
 %    & InAs (mp-$20305$) & & & \\
 %    & ZnCd$_3$Se$_4$ (mp-$1078597$) & & & \\
 %   \hline 
 %   All approaches & InP (mp-$20351$) & InP (mp-$966800$) & ZnGeAs$_2$ (mp-$4008$) & \\ 
 %   to $0.75$ confidence & GeSe (mp-$10759$) & & CdSnP$_2$ (mp-$5213$) & \\ 
 %    & SiSn (mp-$1009813$) & & & \\
 %   \end{tabular}
 %   \label{tab:crystalsystems}
%\end{table}

\begin{table}[b]
    \centering 
    \caption{Overview of material properties as listed in the MP database for the $9$ materials predicted by the empirical approach ($>0.85$) and the $6$ materials predicted by all approaches ($>0.75$). Compounds marked by an asterix are marked as unstable upon tranformation into other phases in the Materials Project database. }
    \begin{tabular}{c|c c c c c c c }
      & Material & Crystal structure & MP code & Density (g/cm$^3$) & Band gap (eV) & a, b, c (\AA) & $\alpha,\beta,\gamma$ ($^\circ$) \\
    \hline 
    Empirical & CdSe & Hexagonal & mp-$1070$ & $5.3$ & $0.6$ & $4.4,4.4,7.2$ & $90,90,120$ \\
    approach  & CuI & Hexagonal & mp-$569346$ & $5.8$ & $1.2$ & $4.3,4.3,7.0$ & $90,90,120$ \\ 
    to $0.85$  & CuI & Cubic & mp-$22895$ & $5.8$ & $1.2$ & $4.3,4.3,4.3$ & $60,60,60$  \\ 
    confidence & CdSe & Cubic & mp-$2691$ & $5.3$ & $0.5$ & $4.4,4.4,4.4$ & $60,60,60$  \\
     & BN & Cubic & mp-$1639$ & $3.5$ & $4.6$ & $2.6,2.6,2.6$ & $60,60,60$  \\
     & InAs & Cubic & mp-$20305$ & $5.3$ & $0.3$ & $4.4,4.4,4.4$ & $60,60,60$ \\
     & ZnCd$_3$Se$_4$ & Cubic & mp-$1078597$ & $5.3$ & $1.7$ & $6.1,6.1,6.1$ & $90,90,90$ \\
     & BC$_2$N & Tetragonal & mp-$1008523$ & $3.3$ & $1.6$ & $2.6,2.6,3.7$ & $90,90,90$ \\
     & BC$_2$N & Orthorombic & mp-$629458$ & $3.4$ & $1.8$ & $2.5,2.6,3.6$ & $90,90,90$ \\
    \hline 
    All & CdSnP$_2$ & Tetragonal & mp-$5213$ & $4.6$ & $0.7$ & $7.2,7.2,7.2$ & $131.1,131.1,71.7$ \\
    approaches & GeSe & Cubic & mp-$10759$ & $5.5$ & $0.4$ & $4.0,4.0,4.0$ & $60,60,60$ \\
    to $0.75$  & InP & Cubic & mp-$20351$ & $4.6$ & $0.5$ & $4.2,4.2,4.2$ &  $60,60,60$\\ 
    confidence & InP & Hexagonal & mp-$966800$ & $4.6$ & $0.5$ & $4.2,4.2,6.9$ & $90,90,120$ \\ 
     & SiSn & Cubic & mp-$1009813$ & $4.4$ & $0.4$ & $4.3,4.3,4.3$ & $60,60,60$ \\
     & ZnGeAs$_2$ & Tetragonal & mp-$4008$ & $5.2$ & $0.6$ & $7.0,7.0,7.0$ & $131.4,131.4,71.2$ \\
    \end{tabular}
    \label{tab:materialproperties}
\end{table}

Consider now the $9$ materials that were classified by the empirical approach as suitable to a confidence of $0.85$ or higher by all four machine learning methods; BN, CdSe (2 structures), BC$_2$N (2 structures), InAs, CuI (2 structures), and ZnCd$_3$Se$_4$. 
The nine materials (considering different crystal structures, see Table~\ref{tab:materialproperties}) each belong to one of the four crystal systems cubic, hexagonal, tetragonal and orthorombic. Figure~\ref{fig:crystalsystems} visualizes the four different crystal systems using examples and  Table~\ref{tab:materialproperties} lists important material properties of the relevant materials as listed in the MP database.  Interestingly, we find that all nine materials appear to be four-fold coordinated, and that the first two lattice vectors ($a$ and $b$) are identical in all nine cases whereas $c$ may differ.  
There appears to be a high degree of symmetry present for all materials, but we find no elemental and perfectly symmetric semiconductors in the list.

Turning to each of the individual material predictions, the compound BN (mp-$1639$) was in fact present in the training data as a suitable candidate, and we believe this a partial explanation for why the models recognize BN as a suitable candidate with high probability. Furthermore, two compositions of CdSe (the cubic mp-$2691$ and the hexagonal mp-$1070$) have been predicted as suitable, possibly as a consequence of the similar compound CdS being labeled as a suitable candidate in the training set. 
%The element S resides in the same group as Se, and 
The two compounds of CdSe share similar properties with calculated band gaps in the MP database of $0.5$~eV and $0.6$~eV, respectively.

%high quality nanocrystals of CdS have been synthesized and studied for use as quantum dots since the nanocrystals emit in the visible range with a direct band gap of $2.5$-$2.6$~eV \cite{BanerjeeR2000Eots,CelebiSerdar2007SaCo,Qutub2012}. 
%Bulk CdS has a slightly lower reported band gap of $2.4$~eV \cite{Qutub2012}.

In the case of BC$_2$N we find two compositions with the same chemical formula; the orthorhombically coordinated (mp-$629458$) and the tetragonally structured (mp-$1008523$) BC$_2$N. The former takes on a polar space group while the latter does not. The band gaps are listed as $1.9$~eV and $1.7$~eV, respectively, in the MP database. 
%BC$_2$N is known as heterodiamond and is a hybrid of diamond and BN. 
Both structures exhibit, as expected, strong covalent character and have been studied for applications as nanostructures for electronic devices \cite{Gao2017}, hydrogen storage \cite{Cai2017} and superhard materials \cite{Li2017, Jiang2020}. Importantly, the diamond-like structure of BC$_3$N was recently predicted as a prominent spin qubit material host \cite{Wang2020SpinQB}. By creating a boron (B) vacancy one can in theory obtain a defect center with similar properties to those found for the NV center in diamond. Whether this is also possible for BC$_2$N remains to be seen. We note that BC$_3$N was not represented in the MP database at the time of data extraction and is therefore not included in our dataset.

The compounds InAs (cubic, mp-$20305$), CuI (cubic, mp-$22895$ and hexagonal, mp-$569346$) and ZnCd$_3$Se$_4$ (cubic, mp-$1078597$) are listed in the MP database with band gaps of $0.3$, $1.2$, $1.2$ and $1.7$~eV, respectively. 
%, and exhibit comparable properties. 
To the best of our knowledge, ZnCd$_3$Se$_4$ has yet to be synthesized. 
%, and the MP database lists it as unstable upon decomposition into ZnSe and CdSe (in the list with two entries). 
Self-assembled InAs quantum dots are an exciting possible material to use in quantum technology \cite{Liu2018}. 
CuI has recently been synthesized as single-crystal epitaxial films and was shown to exhibit remarkable optoelectronic properties \cite{Ahn2020}. Interestingly, the material exhibits a large ionic character with more similarities to some of the oxide compounds in the dataset than, e.g., the materials that are more covalent in character like Si, SiC and  diamond.  
The presence of CuI in two configurations  proposes that ionic character alone is not an obstacle for a material to be quantum compatible if the prediction of CuI proves to be merited. It is unknown at this time whether any potential quantum compatible properties of CuI would originate from deep level defects or  nanostructuring (e.g., quantum dots). NaCl was not predicted as suitable for QT in the empirical approach.

Out of these nine materials we note that both instances of CuI are stated in the MP database to decompose into trigonal CuI, which was not present in our predictions. Hexagonal CdSe decomposes into cubic CdSe (in the list of $9$), the cubic BN decomposes to hexagonal BN (mp-$984$, not in the predictions), and ZnCd$_3$Se$_4$ decomposes to ZnSe $+$ CdSe (in the list). Finally, both compounds of the BC$_2$N structure are listed to decompose into hexagonal BN $+$ C.

Lowering the cut-off requirement from $0.85$ to $0.75$ for all models results in $66$ candidate materials for the empirical approach.  
The full list of these $66$ is contained in the Supplementary Material at \cite{supplementary}. 
The list of predicted suitable candidates now also includes ternary compounds of the formula ABC$_2$, in addition to elemental and binary semiconductors, and the nine materials listed above. For the ABC$_2$ structures, the elements Ga, Cd, In and Zn can occupy the A-site, Cu, Sn, Ag and Ge take the B-site, while S, Se, Te, P or As may reside at the C site. Most of the predicted compounds include at least one toxic element with one exception: ZnGeP$_2$ (mp-$4524$) in a chalcopyrite-like tetragonal crystal structure with an indirect band gap of $1.2$~eV \cite{Zhang2015} as reported in the MP database. In comparison, the experimentally reported band gap is somewhat larger at $2.0$~eV \cite{Xing1989}. 
ZnGeP$_2$ crystallizes in a non-polar space group, possesses no magnetic moment, exhibits strongly covalent bonds, and has been reported as an excellent mid-IR transparent crystal material that is suitable for nonlinear optical applications \cite{Zhang2015}. Importantly, it is possible to integrate sources of photon quantum states based on nonlinear optics with ZnGeP$_2$ \cite{Caspani2017}. 
We therefore identify ZnGeP$_2$ as an eligible candidate material for QT, but it remains to be seen whether the candidate can facilitate, e.g., the isolated deep energy levels often associated with defects exhibiting quantum compatible properties, or instead be a candidate for nanostructure based QT.

Among the $66$ materials that all models in the empirical approach agreed on to a $0.75$ confidence level, we also emphasize the presence of interesting materials like Ge (in three configurations mp-$137$, mp-$1067619$ and mp-$1198022$), GeC, BP and InP. The element Ge in the cubic structure (mp-$1198022$) shares many properties with Si and C in addition to the periodic column number. 
% MB Do we want to emphasize Ge, as we are downplaying the importance of elemental semiconductors? Or, it could contain an impurity
Germanium has the highest hole mobility of all semiconductors at room temperature, and is therefore considered a key material for the process of extending the chip performance in classical computers beyond the limits imposed by miniaturization \cite{Scappucci2020}. Furthermore, like SiC, device design based on Ge can take advantage of the mature large-scale fabrication of silicon due to the material's comparable properties.  
%
Similar considerations could be made for the case of GeC. 
Data from the Materials Project database suggests that the cubic compound GeC (mp-$1002164$) is a covalently bonded semiconductor having a band gap of $1.8$~eV. 
Consequently, with SiC being widely known as a highly suitable host material for QT compatible defects, we encourage further research on GeC due to its similarities to SiC.

The compound BP (mp-$1479$ and mp-$1008559$) is present in our predictions in both the cubic and hexagonal structures. The indirect band gaps of the different structures are calculated at $1.5$~eV and $1.1$~eV, respectively, as reported in the MP database. Both configurations of BP are nonmagnetic and share several properties with the entries mentioned above; including ... BP is non-toxic and has been synthesized with a potential for large-scale production \cite{MukhanovVladimirA2016Umso}. 
Lastly, InP (mp-$966800$) was also predicted as a suitable candidate to $0.75$ confidence. The compound inhabits the hexagonal structure with corner-sharing InP$_4$ tetrahedra. InP is reported in the Materials Project to have a direct band gap of $0.5$~eV and is considered as one of the most promising candidates to compete with Cd- or Pb- based QDs for, e.g., display and lighting applications  \cite{Zhang2020a, Won2019}. 
We underline the possibility of using InP-based quantum dots for QT applications as well.

We note that all ML methods in the empirical approach agree on several oxides being potential candidates when the prediction cut-off is set low enough (but still above $0.5$). However, we find that almost all the oxide compounds fall between the decision boundary defining suitable and unsuitable candidates, and none are present in the list containing the $66$ suitable candidates using a $0.75$ cut-off. Due to the labeling of ZnO as a suitable candidate in the training set, we believe that the boundary was shifted sufficiently to admit several oxides as suitable candidates. 
Removing ZnO from the training set for the empirical approach might result in fewer or even no oxide compounds being predicted above a suitability limit of $0.5$ and could be an interesting future pathway for investigation.

Comparing to the work of \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, they suggests a list of $541$ viable hosts after the data mining procedure.  
Among these, we find only a single material present in our list of $66$ candidates predicted by the four machine learning methods after training on the dataset resulting from the empirical approach. This material is the nontoxic compound MgSe (mp-$10760$) which crystallizes in the rock-salt structure. Its band gap is $2.0$~eV in the MP database with the experimental value being reported from $2.45$~eV (optical band gap) \cite{Ubale2014} up to as large as $5.6$~eV \cite{SaumGeorge1959}. MgSe is notable for its available spin-zero isotopes in accordance with the criteria set by  \citeauthor{Weber2010} \cite{Weber2010} and \citeauthor{Ferrenti2020} \cite{Ferrenti2020}. We believe that these criteria may favor defects acting as spin centers with qubit potential and MgSe is thus identified as another promising host material in this respect.

\subsubsection*{Comparing the approaches}
Comparing the three approaches (see Table~\ref{tab:probabilites}), we observe  that the extended Ferrenti approach is the least restricted also for the predicted materials and admits the largest number of entries through to the predictions regardless of the confidence cut-off value. The Ferrenti approach also admits a large amount of materials through from the training set as compared to the empirical approach, but the resulting data are otherwise not very different from the extended version, whereas the empirical approach predicts fewer suitable candidates where manually verification is achievable. Manual verification through a literature survey will often not be possible, however, and perfecting automatic data mining and analysis is therefore an important goal of material informatics \cite{rickman2019}.

Despite notable differences, as mentioned above, there is substantial overlap of predicted materials between the three approaches. 
We find that $119$ of the $214$ candidates predicted as suitable by the empirical approach (when the cut-off is set to $0.5$) were also selected by all machine learning methods in the extended Ferrenti approach. 
Similarly, $78$ of the $214$ are also predicted as suitable by all models in the Ferrenti approach. All approaches and their corresponding machine learning models agree on a total of $47$ potential candidates to $0.5$ confidence, where $8$ are elementary (unary), $29$ binary, and $10$ tertiary  
(see the Supplementary Material at \cite{supplementary} for the full list). 
Several interesting materials that were also discussed above for the predictions by the empirical approach are present among these $47$, including BP, CdSe, BP, GeC, InP and Ge. However, we also find certain materials that can be classified neither as semiconductors nor solid-state, such as P, I, N$_2$ and H$_2$. We note that none of these were admitted through for the empirical approach when the confidence cut-off was set to $0.75$ or higher.

Importantly, there are $6$ materials that all approaches and methods agree on to a $0.75$ confidence level. These are ZnGeAs$_2$ (tetragonal mp-$4008$), CdSnP$_2$ (tetragonal mp-$5213$), GeSe (cubic mp-$10759$), InP (cubic mp-$20351$), InP (hexagonal mp-$966800$) and SiSn (cubic mp-$1009813$). Here, we can distinguish three groupings in crystal structure and they agree with those discussed above; cubic, hexagonal and tetragonal. See Fig.~\ref{fig:crystalsystems} for illustration of the crystal systems and Table~\ref{tab:materialproperties} for an overview of the corresponding material properties.     
Note that hexagonal InP is listed in the MP database as decomposing into cubic InP while SiSn is stated to decompose into Si $+$ Sn. 
We can deduce a series of similar trends as for the $9$ materials from the empirical approach --- e.g., the same space groups are present, the materials take on only binary or tertiary compounds, and $a=b$ for all materials while $c$ is different for hexagonal InP (unstable). However, four-fold coordination and the bond angle trends are not maintained for all materials (GeSe is six-fold coordinated). 
We highlight these six compounds, along with the nine predicted by the empirical approach to $0.85$ confidence, as particularly interesting for future in-depth theoretical and experimental study.

\section*{Discussion}
%\subsection*{Important features for QT}

%histograms 
\begin{figure}[ht]
    \begin{subfigure}[b]{1\textwidth}
    \centering
    \input{figures/histograms/new/LegendHistPaper.tex}
  \end{subfigure}
  \par\bigskip
\begin{subfigure}[b]{0.45\textwidth}
    \input{figures/histograms/new/MP_Eg_count_075.tex}
    \subcaption{}
\end{subfigure}
\begin{subfigure}[b]{0.45\textwidth}
    \input{figures/histograms/new/IonProperty_max ionic char_count_075.tex}
    \subcaption{}
\end{subfigure}%

\begin{subfigure}[b]{0.45\textwidth}
    \input{figures/histograms/new/GeneralizedRDF_std_dev Gaussian center=1.0 width=1.0_count_075.tex}
    \subcaption{}
\end{subfigure}
\begin{subfigure}[b]{0.45\textwidth}
    \input{figures/histograms/new/AverageBondLength_std_dev Average bond length_count_075.tex}
    \subcaption{}
\end{subfigure}
\caption{Histograms of predicted suitable materials as a function of the (a) PBE-calculated band gap from Materials Project, (b) maximum ionic character, (c) standard deviation of the radial distribution function (RDF) and (d) standard deviation of the average bond length. The total number of predicted materials is  $6804$ for the Ferrenti approach, $9227$ for the extended Ferrenti approach, and $214$ for the empirical approach. The Ferrenti and extended Ferrenti approaches refer to the left y-axis and the empirical approach to the right.
    }
\label{fig:histogram_new}
\end{figure}

%Turning to why the specific materials were chosen ... 
Taking a closer look at the reasoning behind the choices made by the different ML methods during the classification process, we find important differences between the three labeling approaches, and can start to discern the driving forces for manifestation of quantum compatible properties in semiconductors. 
The analysis of the principal components extracted from the machine learning approaches revealed that the most important principal component for both the Ferrenti and extended Ferrenti approaches, includes features related to the band gap and chemical environment. This means that the band gap criterion imposed in the training set selection is at least somewhat satisfied. 
In other words, the Ferrenti approaches appear to reproduce the logic of the initial selection process and little new information is gained. 
On the other hand, the band gap related features were not important in the dominant principal component in the empirical approach.

Figure~\ref{fig:histogram_new}(a) contains histograms of the number of predicted suitable materials for all approaches as a function of  band gap (as listed in the MP database), and reveals that all approaches predicted a substantial amount of materials with a low band gap. Moreover, the distribution of materials that was identified as suitable is rather broad with the Ferrenti based approaches exhibiting a peak around $2.5$~eV. 
Coincidentally, the band gap of, e.g., 4$H$-SiC is usually computed in the range around $2.5$~eV at the PBE level. 
The empirical approach exhibits a more scattered data distribution with local maxima around $1$-$2$ and $4$~eV.

The secondmost important principal component in the Ferrenti approach encompasses properties such as the ionic character, covalent radius and maximum packing efficiency (see the Supplementary Material at \cite{supplementary} for the predicted material distributions of the latter two features).  
%The maximum ionic character is also deemed an important part of PC2. 
Intriguingly, as shown in Figure~\ref{fig:histogram_new}(b), the predicted candidates distribute over a broad range of ionic characters  for all three approaches - even peaking at a relatively high ionic character of $0.8$. 
Moreover, we find that the covalent radii of the materials (see the Supplementary Material) distribute with two peaks in the data in all approaches. 
%The maxima are slightly shifted towards the right for the empirical approach compared to the two others but otherwise the data distributions are similar. 
The trend of two data peaks is repeated for the maximum packing efficiency  but is much more prominent for the empirical approach. This indicates that the material density, or in other words the bond length, is an important parameter for QT suitability.

The machine learning methods in the empirical approach consistently identified the first principal component as the predominant one. Identifying the single most important feature in this principal component proved challenging as it is the combined impact of several features that is important.
Here, we find that the standard deviation of the radial distribution function (RDF) has a  particularly strong impact since it appears four times in different forms in the top ten list over dominating features. One configuration of the standard deviation of the RDF is demonstrated in Figure~\ref{fig:histogram_new}(c), with two others being included in the Supplementary Material \cite{supplementary}. Notably, substantial discrepancies between the Ferrenti approaches and the empirical approach are found for the standard deviation of the radial distribution function. This reflects the observations from the principal component analysis where the first principal component is dominated by the RDF and similarly symmetry related features. For the empirical approach, there is a sharp peak in the preferred value for the standard deviation for the RDF, while the Ferrenti and augmented Ferrenti approaches experience a more even distribution across a broader range.

We interpret the standard deviation  of the RDF such that zero std in the RDF means that there is no variation in the radial symmetry throughout the material. Similarly, zero standard deviation in the average bond length would mean that all bonds are identical throughout the crystal. Intriguingly, the peaks in the distributions of the predicted materials are not found for perfectly symmetric materials with identical bond lengths; instead, some variation in the bond and wavefunction distributions is found to be optimal for a material to be suitable for QT.  
%This is in keeping with no elemental semiconductors being present in the top lists of predicted materials above the $0.75$ confidence level. 
Note that the maxima  in the histograms in Fig.~\ref{fig:histogram_new} seem to appear for moderate standard deviations in the RDF and average bond length, indicating that a certain degree of symmetry is likely necessary for a material to act as a good QT host. The exact degree and type of symmetry is still open to debate and should be studied further. 
Similar symmetry related features such as the standard deviation of the average bond length (see Figure~\ref{fig:histogram_new}d), the site fingerprint of the chemical environment and the bond orientation (see the Supplementary Material) are also influential in guiding the machine learning algorithms upon classifying suitable and unsuitable materials for QT in the empirical approach.

Interestingly, none of the nine materials predicted by the empirical approach with $0.85$ confidence, nor any of the six materials predicted by all approaches to $0.75$ confidence, are elemental. At most three different elements are present, but the emphasis is clearly on binary compounds. This resonates strongly with the observations from the principal component analysis: an optimal degree of crystalline order likely exists for a material to manifest QT compatible properties, but retaining some variations throughout the crystal in e.g. bond length and symmetry.  This is  in  
%in contrast 
contrast to the expectations that guided the formulation of the test, training and validation sets in the Ferrenti and extended Ferrenti approaches. Where most previous works have highlighted features such as band gap, polarity and ionic character as vital for a semiconductor to be a good quantum host, our results show that local variations in the crystal structure related to symmetry and bond angles, length and orientation should be considered as well.

%MB 
%Also discuss max packing efficiencies. Now in Supplementary. Two clear peaks for empirical approach.

Considering the trends in machine learning selectivity in light of the specific defect centers we know to be quantum compatible reveals fascinating characteristics. None of the known quantum emitters or spin centers seem to appear in completely uniform systems. For example, in the symmetric materials of diamond and silicon, QT compatible defects are not intrinsic; neither the silicon vacancy in silicon nor the carbon vacancy in diamond, for example,  have exhibited single-photon emission or controllable spin coherence. Instead, we find that QT effects often appear after impurities are introduced, as evidenced by the phosphorous and carbon impurities in silicon [REF] and the nitrogen-vacancy, germanium-vacancy, tin-vacancy an lead-vacancy centers in diamond [REF]. 
For the binary system of silicon carbide in different polytypes, on the other hand, intrinsic defects like the silicon vacancy and the divacancy are appropriate for our goals. 
While these trends have yet to be verified on a grander scale, our findings provide strong indications that local variations in the crystal structure are paramount for QT  compatible properties to manifest in a semiconductor host. 
Our findings are corroborated by the fact that all four ML methods exhibit similar results for all datasets. The methods we have chosen are all well tested, with random forests and gradient boosting methods as those which tend to outperform the other, resulting normally in a small bias and variance \cite{Hastie2009,Mehta2019,Murphy2012}.

%\section*{Discussion} 
%MB I removed the initial discussion here, as it was a bit repetitive 
%We have developed data extraction tools for numerous material science databases, and constructed over $4800$ physics-informed features for a dataset consisting of more than $25.000$ materials. Furthermore, three approaches to data mining for QT applications were developed; (i) the Ferrenti approach, (ii) the extended Ferrenti approach, and (iii) the empirical approach. Three distinct sets of labeled data were defined for training, testing and validation of four different machine learning algorithms (logistic regression, decision trees, random forests and gradient boosting).

%NB 
%Morten, kan ta en titt her?  Se ovenfor.

To summarize, the data extraction tools developed in this work resulted in a dataset of over $25.000$ materials and more than $4800$ physics-informed features. Three different approaches (Ferrenti, extended Ferrenti and empirical) for data labeling were employed for training, testing and validation of four machine learning algorithms. 
We find a much clearer separation between suitable and unsuitable candidates after data labeling in the empirical approach than for the two Ferrenti based ones. 
Additionally, much fewer principal components are needed for obtaining optimal performance of the machine learning algorithms for the empirical approach.
The principal component analyses of the ML methods' performance imply that the Ferrenti and extended Ferrenti approaches, with their strong focus on band gap and bonding character when assessing a material's quantum compatibility, largely replicate the specifications of the data labeling process. 
Valuable insight is gained from the empirical approach which highlights the importance of symmetry related properties in the bond orientations and wavefunctions.% instead without being guided by assumptions. 
%that the formulation of the training sets for the Ferrenti and extended Ferrenti approaches was broad, and hence multiple dimensions would be required for distinguishing between materials labeled as suitable and unsuitable for QT in the two training sets. 
%The stringent restrictions set for the empirical approach, on the other hand, proved to give a distinct difference between suitable and unsuitable candidates by visualizing the two main principal components of the covariance matrix. Moreover, the smaller number of materials predicted for the empirical approach permits manual verification.

The findings presented here firmly establish that material informatics (from data mining via featurization to machine learning predictions) is a viable and important route to new discoveries in important fields.  Our focus has been on predicting new candidate materials to host single-photon emitters and spin centers for quantum technology applications, but the framework we have developed is suitable for other fields as well. 
Two main routes can be followed to exploit the results. One aspect is the pursuit of experimental verification of QT compatible effect in the materials predicted as suitable by the four machine learning methods. Another, perhaps even more important, is to use the features and trends identified during the data mining and prediction process to understand the distinct material characteristics that enable quantum effects to manifest, opening  thereby up for new discoveries in the field of quantum technologies.

\section*{Methods}

\subsection*{Databases}
The Materials Project \cite{Jain2013, Jain2018} is an open-source project containing ground state properties of materials calculated using Density functional theory (DFT) [REF] as implemented in the Vienna Ab initio Simulation Package (VASP) \cite{Kresse1996}. The Perdew-Burke-Ernzerhof \cite{Perdew1996} (PBE) functional is used to calculate band structures, while for transition metals, a $+U$ correction is applied to correct for correlation effects \cite{Wang2006}. The project is known as the initiator of materials genomics and offers a variety of calculated properties for over one hundred thousand inorganic crystalline materials, with frequent updates and extensions. Data extraction from Materials Project was performed in December of $2020$ for the Ferrenti and extended Ferrenti approach, and in March $2021$ for the empirical approach. Therefore, the two former approaches includes $77$ more materials due to deprecated entries in the Materials Project database.

The Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015} contains thermodynamic and structural properties of more than $600.000$  materials. The calculations are performed with the VASP software and the electron exchange and correlation are described with the PBE functional. The $+U$ extension is included for several calculations considering specific elements \cite{Stevanovic2012}. Data extraction from OQMD was done in February of $2021$.

JARVIS-DFT \cite{Choudhary2020} is an open-source database based on the VASP software and consists of roughly $40.000$ three-dimensional materials using the vdW-DF-OptB88 (OPT) functional \cite{Thonhauser2007, Klimes2011}. Structures included in the database are originally taken from the Materials 
%mhj: ref to materials project is missing
Project \cite{Jain2013, Jain2018}, and then re-optimized using the OPT functional. Finally, the combination of the OPT and modified Becke-Johnson (mBJ) functionals \cite{Tran2009} is used to obtain a representative band gap for each structure \cite{Choudhary2018a}. Data extraction from JARVIS-DFT was done in January of $2021$, were we utilized the version made available 30.04.2021 at \url{figshare.com}.

The AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} repository is an automatic software framework for the calculations of a wide range of inorganic material properties. They utilize the PBE functional within VASP 
%with projector-extended wave (PAW) potentials 
to relax and optimize the ICSD-structure.  Data-extraction was done in the period of January-February of $2021$.

AFLOW-ML \cite{Isayev2017} is an application programming interface (API) that uses machine learning to predict thermo-mechanical and electronic properties based on the chemical composition and atomic structure alone, which are denoted as \textit{fragment descriptors}. Initially, the API decides whether a given material is a metal or an insulator, where the latter is confirmed with an additional regression method to predict the band gap. The accuracy is validated by a five-fold cross-validation process for each machine learning method, where they report a $93 \ \%$ prediction success of their initial binary classification method. In this work we utilized the Property Labeled Material Fragments (PLMF) openly available at their website \url{aflowlib.org}. We extract the crystallographic information file (CIF) of the crystals from Materials Project, and then use this as input to AFLOW-ML, which then returns an anticipated band gap. The process was done during January of $2021$.

\subsection*{Material Informatics}  
Matminer \cite{Ward2018} is an open source toolkit for material analysis written in Python. Matminer provides modules to extract information from a wide variety of databases. Additionally, they provide the tools to construct possibly thousands of features from calculations based on a material's composition, structure and electronic properties from DFT calculations, and have frameworks for visualization and automatic machine learning. 
To apply Matminer's featurization tools, we extend an existing implementation by \citeauthor{Breuck2021} \cite{Breuck2021}, which was used to generate a supervised machine learning framework called the MODnet. The implementation by \citeauthor{Breuck2021} provides featurization for a material's composition, structure and atomic sites. However, Matminer also provides featurization tools for a material's density of states (DOS) and band structure. Therefore, we extend their implementation to facilitate such featurizations. The features selected for featurization herein are summarized in the Supplementary Material at \cite{supplementary}.

Pymatgen, a robust, open-source Python library for materials analysis \cite{pymatgen}, was also employed to extract and generate features for several of the databases mentioned above.

\subsection*{Machine Learning}

Machine learning represents the science of giving computers the ability to learn without being explicitly programmed. The idea is that there exist generic algorithms which can be used to find patterns in a broad class of datasets without having to write code specifically for each problem. The algorithm builds its own logic based on the data.

The approaches to machine learning are many, but are often split into two main categories, supervised and unsupervised. In supervised learning we know the answer to a problem, and let the computer deduce the logic behind it. On the other hand, unsupervised learning is a method for finding patterns and relationship in datasets without any prior knowledge of the system. Many researchers also operate with a third category, namely reinforcement learning. This is a paradigm of learning inspired by behavioral psychology, where learning is achieved by trial-and-error, solely from rewards and punishment. In this work our focus is on supervised learning only with labeled data for classification problems.

In this work we have applied four well-known and tested machine learning methods for classification problems, these are (see for example \cite{Hastie2009,Mehta2019} for discussions and applications):
\begin{enumerate}
    \item Logistic regression,
    \item Decision trees,
    \item Random forests,
    \item Gradient boosting.
\end{enumerate}
Logistic regression \cite{Hastie2009} is a simple and frequently used method for binary and multi-category classification problems. In addition to logistic regression, we have also applied and tested the predictions made by decision trees and ensemble methods like random forests and gradient boosting, the latter through the application of the computationally efficient XGBoost library \cite{xgboost2016}. Gradient boosting and random forests use decision trees as weak learners and improve their predictability. For random forests this is implemented through a collection of randomized decision trees where a  subset of the features in the datasets are selected randomly when building a decision tree. Boosting methods like gradient boosting use decision trees as  weak learners and improve upon these by an iterative process that involves the estimation of the gradients of the cost/loss function \cite{Hastie2009}. Pure decision trees can easily lead to overfitting of the data under study, leading to a machine learning method that exhibits a high variance. Ensemble methods like random forests and gradient boosting on the other hand tend to soften the overfitting problem, resulting in both a small bias and a reduced variance of the employed method, see for example Refs.~\cite{Hastie2009,Mehta2019} for an in-depth discussion of the bias-variance trade-off in machine learning. Gradient boosting implemented through the  XGBoost library \cite{xgboost2016} is widely used by data scientists to achieve state-of-the-art results on many machine learning challenges.

\section*{Data availability} 
The datasets that support the findings of this study are available online at \cite{Ohebbi2021}.

\section*{Code availability} 
The codes employed to develop the machine learning results are available online at 
%https://github.com/ohebbi/predicting-solid-state-qubit-material-hosts 
\cite{Ohebbi2021}.
%Update with new release later).

\bibliography{apssamp}

\begin{acknowledgments}

The work of LV and MEB was supported by the Research Council of Norway and the University of Oslo through the frontier research project FUNDAMeNT (no. 251131, FriPro ToppForsk-program). 
%Some of the computations were performed on resources provided by UNINETT Sigma2 - the National Infrastructure for High Performance Computing and Data Storage in Norway.  
The work of MEB was supported by an ETH Zurich Postdoctoral Fellowship. 
The work of MHJ was supported by the U.S. Department of Energy,
Office of Science, office of Nuclear Physics under grant
No. DE-SC0021152 and U.S. National Science Foundation Grants
No. PHY-1404159 and PHY-2013047.
The work of SGWL and ØSS was supported by the Norwegian Directorate for International Cooperation and Quality Enhancement in Higher Education (DIKU) which supports the Center for Computing in Science Education (CCSE).

\end{acknowledgments}

\section*{Author contributions}
Add! 
%MEB, LV and MHJ conceived the main theme of the work. OH developed the programs and performed the bulk of the work. All authors have contributed to the writing of the paper and the discussion and analysis of the data.

\section*{Competing interests}
The authors declare no competing interests.

\section*{Additional information}
Correspondence and requests for materials should be addressed to ?

\end{document}